### 2023-06-30 scRNAseq/paper_figures.Rmd
add scRNAseq figures

---
title: "20230627 paper figures"
output: 
  html_document:
    df_print: paged
    code_folding: hide
    toc: true
    toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, cache=TRUE, cache.lazy=FALSE, message=FALSE, warning=FALSE, dev=c('png','pdf'))
library(sctransform)
library(Seurat)
library(SeuratDisk)
library(ggplot2)
library(cowplot)
library(ggrepel)
library(dplyr)
library(scRepertoire)
library(tidyr)
library(cluster)
library(plyr)
library(pheatmap)
library(dendextend)
library(ggpubr)
library(ComplexHeatmap)
library(RColorBrewer)
library(Binarize)
options(future.globals.maxSize = 10 * 1024 * 1024^2)
path <- getwd()
```

# setup

load our own PBMC data after projecting on to the reference

- K456985_PBMC
- K783496_PBMC
- K693104_PBMC
- K955601_PBMC
- P1_1_PBMC (re-sequenced)
- P1_2_PBMC
- P3_1_PBMC
- P3_2_PBMC
- P8_1_PBMC
- P8_2_PBMC (re-sequenced)
- P22_1_PBMC
- P22_2_PBMC
- P23_1_PBMC
- P23_2_PBMC
- P25_1_PBMC
- P25_2_PBMC
- P27_1_PBMC
- P27_2_PBMC
- P28_1_PBMC
- P28_2_PBMC
- P22_P23_K783496_CD4
- P8_P3_K456985_CD4
- P27_P28_K693104_CD4
- P12_1_PBMC
- P12_2_PBMC
- P26_1_PBMC
- P26_2_PBMC
- K496398_PBMC
- P12_P26_K496398_CD4

```{r get_data}
pbmc <- readRDS(file.path(path,'seurat','PBMC_combined.rds'))
tcr_combined <- readRDS(file.path(path,'seurat','TCR_combined.rds'))
bcr_combined <- readRDS(file.path(path,'seurat','BCR_combined.rds'))
CD4.integrated <- readRDS(file.path(path,'seurat','CD4_reclustered.rds'))
CD4.integrated <- RenameIdents(CD4.integrated, '12'='10')
Idents(CD4.integrated) <- as.factor(as.numeric(Idents(CD4.integrated)))
```

```{r get_metadata}
study <- read.csv(file.path(path,'sample_sheet','s_hegazy_IBD.txt'),
                 sep='\t') %>%
  dplyr::mutate(Sample.Name=gsub('p','P',Sample.Name))
assay <- read.csv(file.path(path,'sample_sheet','a_hegazy_IBD_totalseq_nucleotide_sequencing.txt'),
                 sep='\t') %>%
  dplyr::mutate(Sample.Name=gsub('p','P',Sample.Name),
                Library.Name=gsub('p','P',Library.Name))

timepoint_colors <- c('ctrl'='gray','Vedo_0d'='#ff7700','Vedo_6wk'='#882e72')
condition_colors <- c('HC'='black','UC'='#1965b0')
responder_colors <- c('R'='#34a048','NR'='#db2F26')
```

```{r normalize}
for (nn in names(pbmc)) {
  DefaultAssay(pbmc[[nn]]) <- 'RNA'
  pbmc[[nn]] <- NormalizeData(pbmc[[nn]],normalization.method='LogNormalize',scale.factor=10000,verbose=FALSE)
}
```

# Fig 1 + supplements

show all data in the reference embedding; match cluster granularity to CyTOF data;

```{r Fig_1C,fig.width=7,fig.height=5}
ct_map <- c('CD8 TCM'='Mmry CD8+ T cells',
            'CD8 TEM'='Mmry CD8+ T cells',
            'CD8 Naive'='Naive CD8+ T cells', 
            'CD4 TCM'='Mmry CD4+ T cells',
            'CD4 TEM'='Mmry CD4+ T cells',
            'CD4 Naive'='Naive CD4+ T cells',
            'NK'='NK cells',
            'NK Proliferating'='NK cells',
            'NK_CD56bright'='NK cells',
            'gdT'='gd T cells',
            'MAIT'='MAIT cells',
            'B intermediate'='IgD+ B cells',
            'B naive'='IgD+ B cells',
            'B memory'='IgD- B cells',
            'CD14 Mono'='cMonocytes',
            'CD16 Mono'='ncMonocytes',
            'CD4 CTL'='Cytotoxic T cells',
            'CD4 Proliferating'='Mmry CD4+ T cells',
            'CD8 Proliferating'='Mmry CD8+ T cells',
            'cDC1'='cDCs',
            'cDC2'='cDCs',
            'pDC'='pDCs',
            'Doublet'='other',
            'Eryth'='other',
            'dnT'='other',
            'ASDC'='other',
            'HSPC'='other',
            'ILC'='ILCs',
            'Plasmablast'='Plasmablast',
            'Platelet'='other',
            'Treg'='Tregs')

celltype_colors <- c('IgD+ B cells'='#E57A3B',
                     'IgD- B cells'='#A57496',
                     'Cytotoxic T cells'='darkolivegreen1',
                     'Naive CD4+ T cells'='#399649',
                     'Mmry CD4+ T cells'='#D3317D',
                     'Naive CD8+ T cells'='#AFD087',
                     'Mmry CD8+ T cells'='#D685AC',
                     'cDCs'='#E47970',
                     'pDCs'='#9B743B',
                     'cMonocytes'='#295D98',
                     'ncMonocytes'='#5999A2',
                     'NK cells'='#8FCBBA',
                     'other'='gray30',
                     'Plasmablast'='#D9A53C',
                     'gd T cells'='#7A3163',
                     'MAIT cells'='#F0AD69',
                     'Tregs'='darkolivegreen3',
                     'ILCs'='plum1')

pbmc_meta <- do.call(rbind,
                     lapply(pbmc[grepl('PBMC',names(pbmc))], 
                            function(sobj) FetchData(sobj, c('refUMAP_1','refUMAP_2',
                                                             'orig.ident','sample',
                                                             'predicted.celltype.l1',
                                                             'predicted.celltype.l2',
                                                             'predicted.celltype.l3')) %>%
                              tibble::rownames_to_column('cell') %>%
                              dplyr::mutate(cell=paste0(orig.ident,'_',cell)))) %>%
  dplyr::left_join(study %>% 
                     dplyr::mutate(Patient.ID=gsub('[P]*([0-9]*)[\\/_][12]','P\\1',Characteristics.Patient.ID.),
                                   Responder=Characteristics.Responder.,
                                   Diagnosis=Characteristics.Diagnosis.,
                                   Batch=gsub(' ','',gsub('_scRNAseq','',Parameter.Value.Method.)),
                                   Timepoint=ifelse(grepl('_1_',Sample.Name),'Vedo_0d',
                                                    ifelse(grepl('_2_',Sample.Name),'Vedo_6wk','ctrl')),
                                   Material=ifelse(Characteristics.Material.=='PBMC','PBMC','CD4')) %>%
                     dplyr::select(Patient.ID,Responder,Diagnosis,Sample.Name,Material,Timepoint,Batch),
                   by=c('sample'='Sample.Name')) %>%
  dplyr::mutate(celltype=plyr::revalue(predicted.celltype.l2,ct_map))

ggplot(pbmc_meta, aes(x=refUMAP_1,y=refUMAP_2,color=celltype)) + 
  geom_point(size=.25) +
  theme_void() + 
  coord_fixed() +
  scale_color_manual(values=celltype_colors) +
  guides(color=guide_legend(override.aes = list(size=3))) +
  theme(legend.key.size=unit(3.5,'mm'))
```

```{r Fig_S3B,fig.width=4.5,fig.height=3.5}
ggplot(pbmc_meta %>% dplyr::sample_frac(1L), 
       aes(x=refUMAP_1,y=refUMAP_2,color=Patient.ID)) + 
  geom_point(size=.25) +
  theme_void() + 
  coord_fixed() +
  guides(color=guide_legend(override.aes = list(size=3))) +
  theme(legend.key.size=unit(3.5,'mm'))
```

create heatmap of ADT expression for these celltypes

```{r Fig_S3A,fig.width=7,fig.height=4.5, cache=FALSE}
ABs <- c("CD3","CD8a","CD40","CD19","CD20","CD45RA","CD38","CD69","HLA-DR",
         "CD11c","CD123","CD16","CD161","CD1c","CD14","CD25","CD127","CD56","CD4","CD45")

for (nn in grep('PBMC',names(pbmc),value=TRUE)) {
  DefaultAssay(pbmc[[nn]]) <- 'orig_ADT'
}

ADT <- do.call(rbind,
               lapply(pbmc[grepl('PBMC',names(pbmc))], 
                      function(sobj) FetchData(sobj, c(ABs,'orig.ident')) %>%
                        tibble::rownames_to_column('cell') %>%
                        dplyr::mutate(cell=paste0(orig.ident,'_',cell)))) %>%
  dplyr::left_join(pbmc_meta[,c('cell','celltype')],by='cell') %>%
  dplyr::mutate_if(is.numeric,  scales::rescale, to=c(0,1)) %>%
  dplyr::group_by(celltype) %>%
  dplyr::summarise_if(is.numeric, median)

pheatmap(ADT %>% tibble::column_to_rownames('celltype'),
         color=colorRampPalette(c("white","gray80","gray20","black"))(100),
         annotation_row = data.frame(celltype=as.factor(ADT$celltype),row.names=ADT$celltype),
         annotation_colors = list(celltype=celltype_colors[ADT$celltype]),
         annotation_legend=FALSE,
         annotation_names_row=FALSE,
         name='scaled median expression')
```

compare PBMC composition between the time points

```{r Fig_S3C, fig.width=7,fig.height=4}
#library(lme4)
library(gtools)

cols <- c('sample','orig.ident','predicted.celltype.l2')
all_frac <- do.call(rbind,lapply(pbmc, function(x) x[[cols]])) %>%
  dplyr::mutate(celltype=plyr::revalue(predicted.celltype.l2,ct_map)) %>%
  dplyr::group_by(sample,celltype) %>%
  dplyr::summarise(n=n()) %>%
  dplyr::mutate(frac=n/sum(n)) %>%
  dplyr::left_join(study %>% 
                     dplyr::mutate(Patient.ID=gsub('[P]*([0-9]*)[\\/_][12]','P\\1',Characteristics.Patient.ID.),
                                   Responder=Characteristics.Responder.,
                                   Diagnosis=Characteristics.Diagnosis.,
                                   Batch=gsub(' ','',gsub('_scRNAseq','',Parameter.Value.Method.)),
                                   Timepoint=ifelse(grepl('_1_',Sample.Name),'Vedo_0d',
                                                    ifelse(grepl('_2_',Sample.Name),'Vedo_6wk','ctrl')),
                                   Material=ifelse(Characteristics.Material.=='PBMC','PBMC','CD4')) %>%
                     dplyr::select(Patient.ID,Responder,Diagnosis,Sample.Name,Material,Timepoint,Batch),
                   by=c('sample'='Sample.Name'))

pbmc_frac <- all_frac[all_frac$Material=='PBMC',]
pbmc_frac$Timepoint <- factor(pbmc_frac$Timepoint, levels=c('Vedo_0d','Vedo_6wk','ctrl'))

pl1 <- pbmc_frac %>%
  dplyr::filter(Timepoint %in% c('ctrl','Vedo_0d')) %>%
  dplyr::mutate(condition=factor(ifelse(Diagnosis=='control','HC','UC'),levels=c('HC','UC'))) %>%
  ggplot(aes(x=condition,y=frac,fill=condition,color=condition)) + 
  geom_boxplot(outlier.shape=NA) +
  geom_point(size=.5,color='gray30') +
  stat_compare_means(aes(group=celltype),comparisons=list(c('UC','HC')),label = "p.format", 
                     method='wilcox.test',bracket.size=.25,tip.length=.01,size=2) +
  theme_classic() + 
  coord_cartesian(clip='off') +
  facet_grid(.~celltype,switch='both') +
  scale_fill_manual(values=condition_colors,limits=force) +
  scale_color_manual(values=condition_colors,limits=force) +
  theme(axis.text.x=element_blank(),
        legend.position='right',
        legend.key.size=unit(3.5,'mm'),
        strip.background=element_blank(),
        strip.placement='outside',
        strip.text=element_text(size=8,angle=45,hjust=1,vjust=1))  +
  labs(x='',y='cell fraction',fill='',color='')

pg <- ggplotGrob(pl1)
for(i in which(grepl("strip-b", pg$layout$name))){
  pg$grobs[[i]]$layout$clip <- "off"
}

grid::grid.draw(pg)
#pl1
```

```{r Fig_1E,fig.width=4.5,fig.height=4, cache=FALSE}
for (nn in grep('PBMC',names(pbmc),value=TRUE)) {
  DefaultAssay(pbmc[[nn]]) <- 'RNA'
}

RNA <- do.call(rbind,
               lapply(pbmc[grepl('PBMC',names(pbmc))], 
                      function(sobj) FetchData(sobj, c('ITGA4','ITGB7','orig.ident')) %>%
                        tibble::rownames_to_column('cell') %>%
                        dplyr::mutate(cell=paste0(orig.ident,'_',cell)) %>%
                        dplyr::select(-orig.ident))) %>%
  dplyr::left_join(pbmc_meta,by='cell') %>%
  dplyr::filter(Timepoint %in% c('Vedo_0d','ctrl')) %>%
  gather(gene,expression,c(ITGA4,ITGB7)) %>%
  dplyr::arrange(gene,expression)

ggplot(RNA, aes(x=refUMAP_1,y=refUMAP_2,color=expression)) +
  geom_point(size=.25) +
  theme_void() + 
  coord_fixed() +
  facet_grid(gene~Diagnosis,switch='y') +
  scale_color_gradient(low='lightgoldenrod1',high='darkred', 
                       na.value='gray80', 
                       #lim=c(1.e-6,5), 
                       #breaks=c(1.e-6,5),
                       #labels=c('low','high'),
                       oob=scales::squish,
                       guide=guide_colorbar(title.position='right',
                                            title.theme=element_text(angle=90,hjust=.5,vjust=.5))) + 
  theme(legend.key.size=unit(3.5,'mm')) +
  labs(color='')
```

```{r Fig_S8A,fig.width=7,fig.height=3}
RNA %>% 
  spread(gene,expression) %>% 
  dplyr::group_by(orig.ident,Patient.ID, celltype,Timepoint,Diagnosis) %>% 
  dplyr::summarise(pos=mean((ITGA4 > 0) & (ITGB7 > 0))) %>%
  dplyr::filter(Timepoint %in% c("Vedo_0d","ctrl")) %>%
  dplyr::mutate(condition=factor(ifelse(Diagnosis=='control','HC','UC'),levels=c('HC','UC'))) %>%
  ggplot(aes(x=condition,y=pos,fill=condition,color=condition)) + 
  geom_boxplot(outlier.shape=NA) +
  geom_point(size=.5,color='gray30') +
  stat_compare_means(aes(group=celltype),comparisons=list(c('UC','HC')),label = "p.format", 
                     method='wilcox.test',bracket.size=.25,tip.length=.01,size=2) +
  theme_classic() + 
  facet_grid(.~celltype,switch='both') +
  scale_fill_manual(values=condition_colors,limits=force) +
  scale_color_manual(values=condition_colors,limits=force) +
  theme(axis.text.x=element_blank(),
        legend.position='right',
        legend.key.size=unit(3.5,'mm'),
        strip.background=element_blank(),
        strip.placement='outside',
        strip.text=element_text(size=8,angle=45,hjust=1,vjust=1))  +
  labs(x='',y='fraction ITGA4/ITGB7\ndouble-positive cells',fill='',color='')
```

# Fig 2

differential expression of ITGA4 and ITGB7

```{r Fig_2A,fig.width=5,fig.height=4}
RNA <- do.call(rbind,
               lapply(pbmc[grepl('PBMC',names(pbmc))], 
                      function(sobj) FetchData(sobj, c('ITGA4','ITGB7','orig.ident')) %>%
                        tibble::rownames_to_column('cell') %>%
                        dplyr::mutate(cell=paste0(orig.ident,'_',cell)) %>%
                        dplyr::select(-orig.ident))) %>%
  dplyr::left_join(pbmc_meta,by='cell') %>%
  gather(gene,expression,c(ITGA4,ITGB7)) %>%
  dplyr::arrange(gene,expression)

ggplot(RNA %>%
         dplyr::group_by(gene,Timepoint,celltype) %>%
         dplyr::summarise(pct.expressed=100*mean(expression > 0),
                          mean_expression=mean(expression)), 
       aes(x=Timepoint,y=celltype,color=mean_expression,size=pct.expressed)) +
  geom_point() +
  facet_wrap(~gene,ncol=2) +
  theme_bw() +
  scale_color_gradient(low='lightgoldenrod1',high='darkred', 
                       na.value='gray80', 
                       #lim=c(1.e-6,1.3),
                       #breaks=c(1.e-6,1.3),
                       #labels=c('low','high'),
                       oob=scales::squish) +
  theme(legend.key.size=unit(3.5,'mm'),
        strip.background=element_blank(),
        axis.text.x=element_text(angle=90,hjust=1,vjust=0)) +
  labs(x='',y='',color='')
```

```{r Fig_2B,fig.width=3.75,fig.height=4, cache=FALSE}
for (nn in grep('PBMC',names(pbmc),value=TRUE)) {
  DefaultAssay(pbmc[[nn]]) <- 'orig_ADT'
}

ADT <- do.call(rbind,
               lapply(pbmc[grepl('PBMC',names(pbmc))], 
                      function(sobj) FetchData(sobj, c('Integrin-b7','orig.ident')) %>%
                        tibble::rownames_to_column('cell') %>%
                        dplyr::mutate(cell=paste0(orig.ident,'_',cell)) %>%
                        dplyr::select(-orig.ident))) %>%
  dplyr::left_join(pbmc_meta,by='cell') %>%
  gather(gene,expression,`Integrin-b7`) %>%
  dplyr::arrange(gene,expression)

threshold <- binarize.kMeans(ADT$expression)@threshold

ggplot(ADT %>%
         dplyr::group_by(gene,Timepoint,celltype) %>%
         dplyr::summarise(pct.expressed=100*mean(expression > threshold),
                          mean_expression=mean(expression)), 
       aes(x=Timepoint,y=celltype,color=mean_expression,size=pct.expressed)) +
  geom_point() +
  facet_wrap(~gene,ncol=2) +
  theme_bw() +
  scale_color_gradient(low='honeydew1',high='forestgreen', 
                       na.value='gray80', 
                       # lim=c(1.e-6,1.3),
                       # breaks=c(1.e-6,1.3),
                       # labels=c('low','high'),
                       oob=scales::squish) +
  theme(legend.key.size=unit(3.5,'mm'),
        strip.background=element_blank(),
        axis.text.x=element_text(angle=90,hjust=1,vjust=0)) +
  labs(x='',y='',color='')
```

# Fig. 3 + supplements

show where we detect clonotypes (not yet filtered for CD4/CD8 T cells!)

```{r Fig_3B,fig.width=4.5,fig.height=3.5}
pbmc.T <- lapply(names(pbmc), function(sample) {
  bcs <- tcr_combined[[sample]]$barcode
  cells <- paste0(sample,'_',sample,'_',Cells(pbmc[[sample]]))
  tmp <- RenameCells(pbmc[[sample]], new.names=cells)
  combineExpression(tcr_combined[[sample]][bcs %in% cells,],tmp, cloneCall="aa")
  })
names(pbmc.T) <- names(pbmc)

pbmc_T_meta <- do.call(rbind,
                       lapply(pbmc.T,#[grepl('PBMC',names(pbmc))], 
                              function(sobj) FetchData(sobj, c('refUMAP_1','refUMAP_2',
                                                               'orig.ident','sample',
                                                               'predicted.celltype.l2',
                                                               'CTaa',
                                                               'cloneType')) %>%
                                tibble::rownames_to_column('cell'))) %>%
                                #dplyr::mutate(cell=paste0(orig.ident,'_',cell)))) %>%
  dplyr::left_join(study %>% 
                     dplyr::mutate(Patient.ID=gsub('[P]*([0-9]*)[\\/_][12]','P\\1',Characteristics.Patient.ID.),
                                   Responder=Characteristics.Responder.,
                                   Diagnosis=Characteristics.Diagnosis.,
                                   Batch=gsub(' ','',gsub('_scRNAseq','',Parameter.Value.Method.)),
                                   Timepoint=ifelse(grepl('_1_',Sample.Name),'Vedo_0d',
                                                    ifelse(grepl('_2_',Sample.Name),'Vedo_6wk','ctrl')),
                                   Material=ifelse(Characteristics.Material.=='PBMC','PBMC','CD4')) %>%
                     dplyr::select(Patient.ID,Responder,Diagnosis,Sample.Name,Material,Timepoint,Batch),
                   by=c('sample'='Sample.Name')) %>%
  dplyr::mutate(celltype=plyr::revalue(predicted.celltype.l2,ct_map))

ggplot(pbmc_T_meta %>% dplyr::mutate(TCR_detection=ifelse(is.na(CTaa),'no TCR','TCR')) %>% dplyr::sample_frac(1L), 
       aes(x=refUMAP_1,y=refUMAP_2,color=TCR_detection)) + 
  geom_point(size=.25) +
  theme_void() + 
  coord_fixed() +
  scale_color_manual(values=c('TCR'='lightblue','no TCR'='grey'),limits=force) +
  guides(color=guide_legend(override.aes = list(size=3))) +
  theme(legend.key.size=unit(3.5,'mm'))

cloneType_colors <- c('hyperexpanded'='plum4',
                      'large'='plum3',
                      'medium'='plum2',
                      'small'='plum1')

ggplot(pbmc_T_meta %>% dplyr::mutate(cloneType=factor(tolower(gsub('Single','Small',gsub(' .*','',cloneType))),
                                                      levels=c('small','medium','large','hyperexpanded'))) %>% dplyr::sample_frac(1L),
       aes(x=refUMAP_1,y=refUMAP_2,color=cloneType)) + 
  geom_point(size=.25) +
  theme_void() + 
  coord_fixed() +
  scale_color_manual(values=cloneType_colors,na.value='gray80') +
  guides(color=guide_legend(override.aes = list(size=3))) +
  theme(legend.key.size=unit(3.5,'mm'))
```

```{r clonotype_celltypes_TCR,fig.width=6.5,fig.height=10}
tmp <- list()
for (sample in grep('PBMC',names(pbmc.T),value=TRUE)) {
  sobj <- pbmc.T[[sample]]
  sobj$celltype <- plyr::revalue(sobj$predicted.celltype.l2,ct_map)
  sobj$cloneType <- factor(tolower(gsub('Single','Small',gsub(' .*','',sobj$cloneType))),
                           levels=c('small','medium','large','hyperexpanded'))
  tmp[[sample]] <- occupiedscRepertoire(sobj,
                                        x.axis = "celltype", 
                                        exportTable=TRUE) %>%
    dplyr::mutate(sample=gsub('_PBMC','',sample))
}

ggplot(do.call(rbind,tmp),aes(x=celltype,y=value,fill=cloneType)) + 
  geom_col() + 
  theme_classic() +
  facet_wrap(~sample,scales='free_y',ncol=2,strip.position='left') +
  scale_fill_manual(values=cloneType_colors) +
  theme(axis.text.x=element_text(angle=90,hjust=1,vjust=.5),
        strip.background=element_blank(),
        strip.placement='outside',
        legend.key.size=unit(3,'mm')) +
  labs(x='',y='')
```

look at clonal diversity (Inv. Simpson) for different CD4 and CD8 subsets in PBMC separately, using the more fine-grained annotation

```{r Fig_3C,fig.width=6,fig.height=2}
cols <- c('sample','orig.ident','predicted.celltype.l2')
all_meta <- do.call(rbind,lapply(pbmc[grepl('PBMC',names(pbmc))], function(x) x@meta.data[,cols])) %>%
  tibble::rownames_to_column('cell') %>%
  dplyr::mutate(celltype=plyr::revalue(predicted.celltype.l2,ct_map),
                cell=paste0(sample,'_',gsub('\\.','_',cell))) %>%
  dplyr::left_join(study %>% 
                     dplyr::mutate(Patient.ID=gsub('[P]*([0-9]*)[\\/_][12]','P\\1',Characteristics.Patient.ID.),
                                   Responder=ifelse(Characteristics.Responder.=='yes','R','NR'),
                                   Diagnosis=Characteristics.Diagnosis.,
                                   Batch=gsub(' ','',gsub('_scRNAseq','',Parameter.Value.Method.)),
                                   Timepoint=ifelse(grepl('_1_',Sample.Name),'Vedo_0d',
                                                    ifelse(grepl('_2_',Sample.Name),'Vedo_6wk','ctrl')),
                                   Material=ifelse(Characteristics.Material.=='PBMC','PBMC','CD4')) %>%
                     dplyr::select(Patient.ID,Responder,Diagnosis,Sample.Name,Material,Timepoint,Batch),
                   by=c('sample'='Sample.Name'))

dfs <- list()
for (ct in grep('CD[48]|Treg',unique(all_meta$celltype),value=TRUE)) {
  bcs <- all_meta$cell[all_meta$celltype==ct]
  tcrcomb <- lapply(tcr_combined,
                    function(x) x[(x$barcode %in% bcs) & !is.na(x$TCR1) & !is.na(x$TCR2),])
  dfs[[ct]] <- clonalDiversity(tcrcomb[lapply(tcrcomb,function(x) dim(x)[1]) > 0],
                               cloneCall='gene',group='sample',exportTable=TRUE, skip.boots=TRUE) %>%
    dplyr::mutate(celltype=ct)
}
df <- do.call(rbind,dfs) %>%
  dplyr::left_join(study %>% 
                     dplyr::mutate(Patient.ID=gsub('[P]*([0-9]*)[\\/_][12]','P\\1',Characteristics.Patient.ID.),
                                   Responder=ifelse(Characteristics.Responder.=='yes','R','NR'),
                                   Diagnosis=Characteristics.Diagnosis.,
                                   Batch=gsub(' ','',gsub('_scRNAseq','',Parameter.Value.Method.)),
                                   Timepoint=ifelse(grepl('_1_',Sample.Name),'Vedo_0d',
                                                    ifelse(grepl('_2_',Sample.Name),'Vedo_6wk','ctrl'))) %>%
                     dplyr::select(Patient.ID,Responder,Diagnosis,Sample.Name,Timepoint,Batch),
                   by=c('sample'='Sample.Name')) %>%
  dplyr::arrange(Patient.ID)

pl <- ggplot(df %>% 
         dplyr::filter(Diagnosis=='UC') %>%
         gather(metric,value,-c(sample,Patient.ID,Responder,Diagnosis,Timepoint,Batch,celltype)) %>%
         dplyr::filter(metric=='Inv.Simpson') %>%
         dplyr::arrange(Responder,celltype,Timepoint,Patient.ID),
       aes(x=Timepoint,y=value)) +
  geom_boxplot(aes(fill=Timepoint,color=Timepoint),outlier.shape=NA) +
  geom_line(aes(group=Patient.ID),color='gray',size=.25) +
  geom_point(size=1,color='gray30') +
  theme_classic() +
  coord_cartesian(clip='off') +
  scale_y_log10() +
  facet_grid(. ~ celltype,scales='free_y',switch='y') +
  stat_compare_means(aes(group=Patient.ID),comparisons=list(c('Vedo_6wk','Vedo_0d')),label = "p.format",paired=TRUE,size=2,bracket.size=.25,tip.length=.01) +
  scale_fill_manual(values=timepoint_colors,limits=force) +
  scale_color_manual(values=timepoint_colors,limits=force) +
  labs(x='',y='Inv. Simpson',title='') +
  theme(legend.key.size=unit(4,'mm'),
        legend.box='horizontal',
        axis.text.x=element_blank(),
        strip.background=element_blank(),
        strip.text=element_text(size=8),
        strip.placement='outside')

pg <- ggplotGrob(pl)
for(i in which(grepl("strip-t", pg$layout$name))){
  pg$grobs[[i]]$layout$clip <- "off"
}

grid::grid.draw(pg)
#pl
```

```{r Fig_S11F,fig.width=6,fig.height=3}
pl <- ggplot(df %>% 
               dplyr::filter(Timepoint!='ctrl',celltype!='Tregs') %>%
               gather(metric,value,-c(sample,Patient.ID,Responder,Diagnosis,Timepoint,Batch,celltype)) %>%
               dplyr::filter(metric=='Inv.Simpson') %>%
               dplyr::arrange(Responder,celltype,Timepoint,Patient.ID),
             aes(x=Responder,y=value)) +
  geom_boxplot(aes(fill=Responder,color=Responder),outlier.shape=NA) +
  geom_point(size=1,color='gray30') +
  theme_classic() +
  coord_cartesian(clip='off') +
  scale_y_log10() +
  facet_grid(Timepoint ~ celltype,scales='free_y',switch='y') +
  stat_compare_means(comparisons=list(c('NR','R')),label = "p.format",size=2,bracket.size=.25,tip.length=.01) +
  scale_fill_manual(values=responder_colors,limits=force) +
  scale_color_manual(values=responder_colors,limits=force) +
  labs(x='',y='Inv. Simpson',title='') +
  theme(legend.key.size=unit(4,'mm'),
        legend.box='horizontal',
        axis.text.x=element_blank(),
        strip.background=element_blank(),
        strip.text=element_text(size=8),
        strip.placement='outside')

pg <- ggplotGrob(pl)
for(i in which(grepl("strip-t", pg$layout$name))){
  pg$grobs[[i]]$layout$clip <- "off"
}

grid::grid.draw(pg)
#pl
```

check TCR diversity for CD4 / CD8 subsets that are

- ITGA4 or ITGB7 positive
- Integrin-b7 positive (after determining a threshold)
- GPR15,  CCR9 or ITGAE positive

```{r Figs_3E_3F_S11B_S11C_S11D,fig.width=2.5,fig.height=2.5, cache=FALSE}
library(Binarize)
for (nn in grep('PBMC',names(pbmc),value=TRUE)) {
  DefaultAssay(pbmc[[nn]]) <- 'RNA'
}

cols <- c('sample','orig.ident','predicted.celltype.l2','ITGA4','ITGB7','Integrin-b7','GPR15','CCR9','ITGAE','MKI67')
pbmc_md <- do.call(rbind,lapply(pbmc[grepl('PBMC',names(pbmc))], function(x) FetchData(x,cols))) %>%
  tibble::rownames_to_column('cell') %>%
  dplyr::mutate(celltype=plyr::revalue(predicted.celltype.l2,ct_map),
                cell=paste0(sample,'_',gsub('\\.','_',cell))) %>%
  dplyr::left_join(study %>% 
                     dplyr::mutate(Patient.ID=gsub('[P]*([0-9]*)[\\/_][12]','P\\1',Characteristics.Patient.ID.),
                                   Responder=Characteristics.Responder.,
                                   Diagnosis=Characteristics.Diagnosis.,
                                   Batch=gsub(' ','',gsub('_scRNAseq','',Parameter.Value.Method.)),
                                   Timepoint=ifelse(grepl('_1_',Sample.Name),'Vedo_0d',
                                                    ifelse(grepl('_2_',Sample.Name),'Vedo_6wk','ctrl')),
                                   Material=ifelse(Characteristics.Material.=='PBMC','PBMC','CD4')) %>%
                     dplyr::select(Patient.ID,Responder,Diagnosis,Sample.Name,Material,Timepoint,Batch),
                   by=c('sample'='Sample.Name'))

threshold <- binarize.kMeans(pbmc_md$`origadt_Integrin-b7`)@threshold
pbmc_md$Integrinb7 <- ifelse(pbmc_md$`origadt_Integrin-b7` > threshold,'pos','neg')

dfs <- list()
for (ct in c('Mmry CD4','Mmry CD8')) {
  bcs <- list(ITGA4=pbmc_md$cell[grepl(ct,pbmc_md$celltype) & (pbmc_md$ITGA4 > 0)],
              ITGB7=pbmc_md$cell[grepl(ct,pbmc_md$celltype) & (pbmc_md$ITGB7 > 0)],
              Integrin_b7=pbmc_md$cell[grepl(ct,pbmc_md$celltype) & (pbmc_md$Integrinb7=='pos')],
              GPR15=pbmc_md$cell[grepl(ct,pbmc_md$celltype) & (pbmc_md$GPR15 > 0)],
              CCR9=pbmc_md$cell[grepl(ct,pbmc_md$celltype) & (pbmc_md$CCR9 > 0)],
              ITGAE=pbmc_md$cell[grepl(ct,pbmc_md$celltype) & (pbmc_md$ITGAE > 0)],
              MKI67=pbmc_md$cell[grepl(ct,pbmc_md$celltype) & (pbmc_md$MKI67 > 0)])
  for (nn in names(bcs)) {
    if (length(bcs[[nn]]) > 0) {
      tcrcomb <- lapply(tcr_combined,
                        function(x) x[(x$barcode %in% bcs[[nn]]) & !is.na(x$TCR1) & !is.na(x$TCR2),])
      dfs[[paste0(ct,'_',nn)]] <- clonalDiversity(tcrcomb,cloneCall='gene',group='sample',exportTable=TRUE,skip.boots=TRUE) %>%
        dplyr::mutate(celltype=ct,
                      subset=nn)
    }
  }
}
df <- do.call(rbind,dfs) %>%
  dplyr::left_join(study %>% 
                     dplyr::mutate(Patient.ID=gsub('[P]*([0-9]*)[\\/_][12]','P\\1',Characteristics.Patient.ID.),
                                   Responder=Characteristics.Responder.,
                                   Diagnosis=Characteristics.Diagnosis.,
                                   Batch=gsub(' ','',gsub('_scRNAseq','',Parameter.Value.Method.)),
                                   Timepoint=ifelse(grepl('_1_',Sample.Name),'Vedo_0d',
                                                    ifelse(grepl('_2_',Sample.Name),'Vedo_6wk','ctrl'))) %>%
                     dplyr::select(Patient.ID,Responder,Diagnosis,Sample.Name,Timepoint,Batch),
                   by=c('sample'='Sample.Name')) %>%
  dplyr::arrange(Patient.ID)

for (ss in c('CCR9','GPR15','Integrin_b7','ITGA4','ITGAE','ITGB7')) {
  for (ct in c('Mmry CD4','Mmry CD8')) {
    pl <- ggplot(df %>% 
                   dplyr::filter(Diagnosis=='UC') %>%
                   gather(metric,value,-c(sample,Patient.ID,Responder,Diagnosis,Timepoint,Batch,celltype,subset)) %>%
                   dplyr::filter(metric=='Inv.Simpson') %>%
                   dplyr::group_by(celltype,subset,Patient.ID) %>%
                   dplyr::mutate(ns=n()) %>%
                   dplyr::filter(ns==2,subset==ss,celltype==ct), 
                 aes(x=Timepoint,y=value)) +
      geom_boxplot(aes(fill=Timepoint,color=Timepoint),outlier.shape=NA) +
      geom_line(aes(group=Patient.ID),color='gray',size=.25) +
      geom_point(size=1,color='gray30') +
      theme_classic() +
      coord_cartesian(clip='off') +
      scale_y_log10() +
      stat_compare_means(comparisons=list(c('Vedo_6wk','Vedo_0d')),paired=TRUE,label = "p.format",size=2,bracket.size=.25,tip.length=.01) +
      scale_fill_manual(values=timepoint_colors) +
      scale_color_manual(values=timepoint_colors) +
      labs(x='',y='Inv. Simpson',title=paste0(ct,'\n',ss)) +
      theme(legend.key.size=unit(4,'mm'),
            legend.box='horizontal',
            axis.text.x=element_text(angle=90,hjust=1,vjust=0),
            strip.background=element_blank(),
            strip.placement='outside')
    print(pl)
  }
}
```

```{r Fig_3H,fig.width=2.5,fig.height=2}
ADT %>% 
  dplyr::group_by(orig.ident,Patient.ID, celltype,Timepoint) %>% 
  dplyr::summarise(pos=mean((expression > threshold))) %>%
  dplyr::filter(Timepoint %in% c("Vedo_0d","Vedo_6wk")) %>%
  dplyr::group_by(celltype,Patient.ID) %>%
  dplyr::mutate(ns=n()) %>%
  dplyr::filter(ns==2, celltype=='Mmry CD4+ T cells') %>%
  ggplot(aes(x=Timepoint,y=pos,fill=Timepoint,color=Timepoint)) + 
  geom_boxplot(outlier.shape=NA) +
  geom_point(size=.5,color='gray30') +
  geom_line(aes(group=Patient.ID),size=.25,color='gray30') +
  stat_compare_means(aes(group=celltype),comparisons=list(c('Vedo_6wk','Vedo_0d')),label = "p.format", 
                     paired=TRUE,method='wilcox.test',bracket.size=.25,tip.length=.01,size=2) +
  theme_classic() + 
  coord_cartesian(clip='off') +
  #facet_grid(.~celltype,switch='both') +
  scale_fill_manual(values=timepoint_colors,limits=force) +
  scale_color_manual(values=timepoint_colors,limits=force) +
  theme(axis.text.x=element_blank(),
        legend.position='right',
        legend.key.size=unit(3.5,'mm'),
        strip.background=element_blank(),
        strip.placement='outside',
        strip.text=element_text(size=8,angle=90,hjust=1,vjust=.5))  +
  labs(x='',y='fraction Integrin-b7-positive cells',fill='',color='')
```

```{r Fig_3G, fig.width=2.5,fig.height=2.}
pbmc_frac %>%
  dplyr::filter(Timepoint %in% c('Vedo_0d','Vedo_6wk')) %>%
  dplyr::group_by(celltype,Patient.ID) %>%
  dplyr::mutate(ns=n()) %>%
  dplyr::filter(ns==2,celltype=='Mmry CD4+ T cells') %>%
  ggplot(aes(x=Timepoint,y=frac,fill=Timepoint,color=Timepoint)) + 
  geom_boxplot(outlier.shape=NA) +
  geom_point(size=.5,color='gray30') +
  geom_line(aes(group=Patient.ID),size=.25,color='gray30') +
  stat_compare_means(aes(group=celltype),comparisons=list(c('Vedo_6wk','Vedo_0d')),label = "p.format", 
                     paired=TRUE,method='wilcox.test',bracket.size=.25,tip.length=.01,size=2) +
  theme_classic() + 
  #facet_grid(.~celltype) +
  coord_cartesian(clip='off') +
  scale_fill_manual(values=timepoint_colors,limits=force) +
  scale_color_manual(values=timepoint_colors,limits=force) +
  theme(axis.text.x=element_blank(),
        legend.position='right',
        legend.key.size=unit(3.5,'mm'),
        strip.background=element_blank(),
        strip.placement='outside',
        strip.text=element_text(size=8,angle=45,hjust=1,vjust=1))  +
  labs(x='',y='cell fraction',fill='',color='',title='Mmry CD4+ T cells')
```

```{r Fig_S11E, fig.width=2.5,fig.height=2.}
pbmc_frac %>%
  dplyr::filter(Timepoint %in% c('Vedo_0d','Vedo_6wk')) %>%
  dplyr::group_by(celltype,Patient.ID) %>%
  dplyr::mutate(ns=n()) %>%
  dplyr::filter(ns==2,celltype=='Mmry CD8+ T cells') %>%
  ggplot(aes(x=Timepoint,y=frac,fill=Timepoint,color=Timepoint)) + 
  geom_boxplot(outlier.shape=NA) +
  geom_point(size=.5,color='gray30') +
  geom_line(aes(group=Patient.ID),size=.25,color='gray30') +
  stat_compare_means(aes(group=celltype),comparisons=list(c('Vedo_6wk','Vedo_0d')),label = "p.format", 
                     paired=TRUE,method='wilcox.test',bracket.size=.25,tip.length=.01,size=2) +
  theme_classic() + 
  coord_cartesian(clip='off') +
  scale_fill_manual(values=timepoint_colors,limits=force) +
  scale_color_manual(values=timepoint_colors,limits=force) +
  theme(axis.text.x=element_blank(),
        legend.position='right',
        legend.key.size=unit(3.5,'mm'),
        strip.background=element_blank(),
        strip.placement='outside',
        strip.text=element_text(size=8,angle=45,hjust=1,vjust=1))  +
  labs(x='',y='cell fraction',fill='',color='',title='Mmry CD8+ T cells')
```

```{r Fig_3D,fig.width=3,fig.height=2.5}
cols <- c('sample','orig.ident','predicted.celltype.l2')
pbmc_md <- do.call(rbind,lapply(pbmc[grepl('PBMC',names(pbmc))], function(x) FetchData(x,cols))) %>%
  tibble::rownames_to_column('cell') %>%
  dplyr::mutate(celltype=predicted.celltype.l2,
                cell=paste0(sample,'_',gsub('\\.','_',cell))) %>%
  dplyr::left_join(study %>% 
                     dplyr::mutate(Patient.ID=gsub('[P]*([0-9]*)[\\/_][12]','P\\1',Characteristics.Patient.ID.),
                                   Responder=Characteristics.Responder.,
                                   Diagnosis=Characteristics.Diagnosis.,
                                   Batch=gsub(' ','',gsub('_scRNAseq','',Parameter.Value.Method.)),
                                   Timepoint=ifelse(grepl('_1_',Sample.Name),'Vedo_0d',
                                                    ifelse(grepl('_2_',Sample.Name),'Vedo_6wk','ctrl')),
                                   Material=ifelse(Characteristics.Material.=='PBMC','PBMC','CD4')) %>%
                     dplyr::select(Patient.ID,Responder,Diagnosis,Sample.Name,Material,Timepoint,Batch),
                   by=c('sample'='Sample.Name'))

dfs <- list() 
#smps <- unique(gsub('_[CD4PMBC]*$','',names(tcr_combined)))
smps <- unique(gsub('_[CD4PMBC]*$','',grep('PBMC',names(tcr_combined),value=TRUE)))
for (ct in c('CD4 Naive','CD4 Proliferating','CD4 TCM','CD4 TEM')) {
  bcs <- pbmc_md$cell[(pbmc_md$celltype==ct)]
  tcrcomb <- lapply(smps,
                    function(x) tcr_combined[[paste0(x,'_PBMC')]] %>%
                      dplyr::filter(barcode %in% bcs,!is.na(TCR1),!is.na(TCR2)) %>%
                      dplyr::mutate(sample=gsub('_PBMC$','',sample)))
  #                          tcr_combined[[paste0(x,'_CD4')]] %>%
  #                            dplyr::filter(barcode %in% bcs,!is.na(TCR1),!is.na(TCR2)) %>%
  #                            dplyr::mutate(sample=gsub('_CD4$tcr_','',sample)))),
  # lapply(setdiff(smps,smps_2),
  #        function(x) tcr_combined[[paste0(x,'_PBMC')]] %>%
  #          dplyr::filter(barcode %in% bcs,!is.na(TCR1),!is.na(TCR2)) %>%
  #          dplyr::mutate(sample=gsub('_PBMC$','',sample))))
  names(tcrcomb) <- smps #c(smps_2,setdiff(smps,smps_2))

  dfs[[ct]] <- clonalDiversity(tcrcomb,cloneCall='gene',group='sample',exportTable=TRUE,skip.boots=TRUE) %>%
    dplyr::mutate(celltype=ct)
}
df <- do.call(rbind,dfs) %>%
  #dplyr::mutate(sample=ifelse(grepl('_CD4$',sample),sample,paste0(sample,'_PBMC'))) %>%
  dplyr::mutate(sample=paste0(sample,'_PBMC')) %>%
  dplyr::left_join(study %>% 
                     dplyr::mutate(Patient.ID=gsub('[P]*([0-9]*)[\\/_][12]','P\\1',Characteristics.Patient.ID.),
                                   Responder=ifelse(Characteristics.Responder.=='yes','R','NR'),
                                   Diagnosis=Characteristics.Diagnosis.,
                                   Batch=gsub(' ','',gsub('_scRNAseq','',Parameter.Value.Method.)),
                                   Timepoint=ifelse(grepl('_1_',Sample.Name),'Vedo_0d',
                                                    ifelse(grepl('_2_',Sample.Name),'Vedo_6wk','ctrl'))) %>%
                     dplyr::select(Patient.ID,Responder,Diagnosis,Sample.Name,Timepoint,Batch),
                   by=c('sample'='Sample.Name')) %>%
  dplyr::arrange(Patient.ID)

pl <- ggplot(df %>%
               dplyr::filter(celltype %in% c('CD4 TCM','CD4 TEM'),
                             Timepoint!='ctrl')%>%
               dplyr::group_by(celltype,Diagnosis,Patient.ID) %>%
               dplyr::mutate(ns=n()) %>%
               dplyr::filter((ns > 1) | (Diagnosis=='control')),
             aes(x=Timepoint,y=`Inv.Simpson`)) +
  geom_boxplot(aes(fill=Timepoint,color=Timepoint),outlier.shape=NA) +
  geom_line(aes(group=Patient.ID),color='gray',size=.25) +
  geom_point(size=1,color='gray30') +
  theme_classic() +
  coord_cartesian(clip='off') +
  scale_y_log10() +
  facet_grid(. ~ celltype,scales='free_y',switch='y') +
  stat_compare_means(comparisons=list(c('Vedo_6wk','Vedo_0d')),paired=TRUE,label = "p.format",size=2,bracket.size=.25,tip.length=.01) +
  stat_compare_means(comparisons=list(c('ctrl','Vedo_0d')),label = "p.format",size=2,bracket.size=.25,tip.length=.01,step.increase=1) +
  scale_fill_manual(values=timepoint_colors) +
  scale_color_manual(values=timepoint_colors) +
  labs(x='',y='Inv. Simpson') +
  theme(legend.key.size=unit(4,'mm'),
        legend.box='horizontal',
        axis.text.x=element_text(angle=90,hjust=1,vjust=0),
        strip.background=element_blank(),
        strip.placement='outside')

pg <- ggplotGrob(pl)
for(i in which(grepl("strip|panel|axis", pg$layout$name))){
  pg$grobs[[i]]$layout$clip <- "off"
}

grid::grid.draw(pg)
```

```{r Fig_S11A,fig.width=6,fig.height=4}
dfs <- list()
for (ct in c('CD4 Naive','CD4 Proliferating','CD4 TCM','CD4 TEM',
             'CD8 Naive','CD8 Proliferating','CD8 TCM','CD8 TEM')) {
  bcs <- pbmc_md$cell[(pbmc_md$celltype==ct) & (pbmc_md$Material=='PBMC')]
  if (length(bcs) > 0) {
    tcrcomb <- lapply(tcr_combined,
                      function(x) x[(x$barcode %in% bcs) & !is.na(x$TCR1) & !is.na(x$TCR2),])
    dfs[[ct]] <- clonalDiversity(tcrcomb,cloneCall='gene',group='sample',exportTable=TRUE,skip.boots=TRUE) %>%
      dplyr::mutate(celltype=gsub('CD[48] ','',ct),
                    Material=ifelse(grepl('CD4',ct),'CD4','CD8'))
  }
}
df <- do.call(rbind,dfs) %>%
  dplyr::left_join(study %>% 
                     dplyr::mutate(Patient.ID=gsub('[P]*([0-9]*)[\\/_][12]','P\\1',Characteristics.Patient.ID.),
                                   Responder=Characteristics.Responder.,
                                   Diagnosis=Characteristics.Diagnosis.,
                                   Batch=gsub(' ','',gsub('_scRNAseq','',Parameter.Value.Method.)),
                                   Timepoint=ifelse(grepl('_1_',Sample.Name),'Vedo_0d',
                                                    ifelse(grepl('_2_',Sample.Name),'Vedo_6wk','ctrl'))) %>%
                     dplyr::select(Patient.ID,Responder,Diagnosis,Sample.Name,Timepoint,Batch),
                   by=c('sample'='Sample.Name')) %>%
  dplyr::arrange(Patient.ID)

pl <- ggplot(df %>%
               dplyr::group_by(celltype,Diagnosis,Material,Patient.ID) %>%
               dplyr::mutate(ns=n()) %>%
               dplyr::filter((ns==2) | (Diagnosis=='control')),
             aes(x=Timepoint,y=`Inv.Simpson`)) +
  geom_boxplot(aes(fill=Timepoint,color=Timepoint),outlier.shape=NA) +
  geom_line(aes(group=Patient.ID),color='gray',size=.25) +
  geom_point(width=.25,size=1,color='gray30') +
  theme_classic() +
  coord_cartesian(clip='off') +
  scale_y_log10() +
  facet_grid(Material ~ celltype,scales='free_y',switch='y') +
  stat_compare_means(comparisons=list(c('Vedo_6wk','Vedo_0d')),paired=TRUE,label = "p.format",size=2,bracket.size=.25,tip.length=.01) +
  stat_compare_means(comparisons=list(c('ctrl','Vedo_0d')),label = "p.format",size=2,bracket.size=.25,tip.length=.01,step.increase=1) +
  scale_fill_manual(values=timepoint_colors) +
  scale_color_manual(values=timepoint_colors) +
  labs(x='',y='Inv. Simpson') +
  theme(legend.key.size=unit(4,'mm'),
        legend.box='horizontal',
        axis.text.x=element_text(angle=90,hjust=1,vjust=0),
        strip.background=element_blank(),
        strip.placement='outside')

pg <- ggplotGrob(pl)
for(i in which(grepl("strip|panel|axis", pg$layout$name))){
  pg$grobs[[i]]$layout$clip <- "off"
}

grid::grid.draw(pg)

#pl
```

```{r Fig_S10B,fig.width=4.5,fig.height=3.5,cache=FALSE}
pbmc.B <- lapply(names(bcr_combined), function(sample) {
  bcs <- bcr_combined[[sample]]$barcode
  cells <- paste0(sample,'_',sample,'_',Cells(pbmc[[sample]]))
  tmp <- RenameCells(pbmc[[sample]], new.names=cells)
  combineExpression(bcr_combined[[sample]][bcs %in% cells,],tmp, cloneCall="aa")
  })
names(pbmc.B) <- names(bcr_combined)

pbmc_B_meta <- do.call(rbind,
                       lapply(pbmc.B,#[grepl('PBMC',names(pbmc))], 
                              function(sobj) FetchData(sobj, c('refUMAP_1','refUMAP_2',
                                                               'orig.ident','sample',
                                                               'predicted.celltype.l2',
                                                               'CTaa',
                                                               'cloneType')) %>%
                                tibble::rownames_to_column('cell'))) %>%
  #dplyr::mutate(cell=paste0(orig.ident,'_',cell)))) %>%
  dplyr::left_join(study %>% 
                     dplyr::mutate(Patient.ID=gsub('[P]*([0-9]*)[\\/_][12]','P\\1',Characteristics.Patient.ID.),
                                   Responder=Characteristics.Responder.,
                                   Diagnosis=Characteristics.Diagnosis.,
                                   Batch=gsub(' ','',gsub('_scRNAseq','',Parameter.Value.Method.)),
                                   Timepoint=ifelse(grepl('_1_',Sample.Name),'Vedo_0d',
                                                    ifelse(grepl('_2_',Sample.Name),'Vedo_6wk','ctrl')),
                                   Material=ifelse(Characteristics.Material.=='PBMC','PBMC','CD4')) %>%
                     dplyr::select(Patient.ID,Responder,Diagnosis,Sample.Name,Material,Timepoint,Batch),
                   by=c('sample'='Sample.Name')) %>%
  dplyr::mutate(celltype=plyr::revalue(predicted.celltype.l2,ct_map))

ggplot(pbmc_B_meta %>% dplyr::mutate(BCR_detection=ifelse(is.na(CTaa),'no BCR','BCR')) %>% dplyr::sample_frac(1L), 
       aes(x=refUMAP_1,y=refUMAP_2,color=BCR_detection)) + 
  geom_point(size=.25) +
  theme_void() + 
  coord_fixed() +
  scale_color_manual(values=c('BCR'='lightgreen','no BCR'='grey'),limits=force) +
  guides(color=guide_legend(override.aes = list(size=3))) +
  theme(legend.key.size=unit(3.5,'mm'))

ggplot(pbmc_B_meta %>% dplyr::mutate(cloneType=factor(tolower(gsub('Single','Small',gsub(' .*','',cloneType))),
                                                      levels=c('small','medium','large','hyperexpanded'))) %>% dplyr::sample_frac(1L),
       aes(x=refUMAP_1,y=refUMAP_2,color=cloneType)) + 
  geom_point(size=.25) +
  theme_void() + 
  coord_fixed() +
  scale_color_manual(values=cloneType_colors,na.value='gray80') +
  guides(color=guide_legend(override.aes = list(size=3))) +
  theme(legend.key.size=unit(3.5,'mm'))
```

look at clonal diversity for B cells

```{r Fig_S10C,fig.width=4.5,fig.height=3}
dfs <- list()
for (ct in grep('B|Plasmablast',unique(all_meta$celltype),value=TRUE)) {
  bcs <- all_meta$cell[all_meta$celltype==ct]
  bcrcomb <- lapply(bcr_combined,
                    function(x) x[(x$barcode %in% bcs) & !grepl('_NA|NA_',x$CTaa),])
  dfs[[ct]] <- clonalDiversity(bcrcomb,cloneCall='gene',group='sample',exportTable=TRUE,skip.boots=TRUE) %>%
    dplyr::mutate(celltype=ct,
                  Material='PBMC')
}
df <- do.call(rbind,dfs) %>%
  dplyr::left_join(study %>% 
                     dplyr::mutate(Patient.ID=gsub('[P]*([0-9]*)[\\/_][12]','P\\1',Characteristics.Patient.ID.),
                                   Responder=Characteristics.Responder.,
                                   Diagnosis=Characteristics.Diagnosis.,
                                   Batch=gsub(' ','',gsub('_scRNAseq','',Parameter.Value.Method.)),
                                   Timepoint=ifelse(grepl('_1_',Sample.Name),'Vedo_0d',
                                                    ifelse(grepl('_2_',Sample.Name),'Vedo_6wk','ctrl'))) %>%
                     dplyr::select(Patient.ID,Responder,Diagnosis,Sample.Name,Timepoint,Batch),
                   by=c('sample'='Sample.Name')) %>%
  dplyr::arrange(Patient.ID)

ggplot(df %>% 
         dplyr::filter(Diagnosis=='UC') %>%
         gather(metric,value,-c(sample,Patient.ID,Responder,Diagnosis,Timepoint,Batch,celltype,Material)) %>%
         dplyr::filter(metric=='Inv.Simpson') %>%
         dplyr::group_by(Patient.ID,celltype) %>%
         dplyr::mutate(ns=n()) %>%
         dplyr::filter(ns==2), 
       aes(x=Timepoint,y=value)) +
  geom_boxplot(aes(fill=Timepoint,color=Timepoint),outlier.shape=NA) +
  geom_line(aes(group=Patient.ID),color='gray',size=.25) +
  geom_point(size=1,color='gray30') +
  theme_classic() +
  coord_cartesian(clip='off') +
  scale_y_log10() +
  facet_grid(~celltype,scales='free_y',switch='y') +
  stat_compare_means(comparisons=list(c('Vedo_6wk','Vedo_0d')),paired=TRUE,label = "p.format",size=2,bracket.size=.25,tip.length=.01) +
  scale_fill_manual(values=timepoint_colors) +
  scale_color_manual(values=timepoint_colors) +
  labs(x='',y='Inv. Simpson',title='') +
  theme(legend.key.size=unit(4,'mm'),
        legend.box='horizontal',
        axis.text.x=element_text(angle=90,hjust=1,vjust=0),
        strip.background=element_blank(),
        strip.placement='outside')
```

```{r Fig_S10F,fig.width=4.5,fig.height=3}
ggplot(df %>% 
         dplyr::filter(Diagnosis=='UC') %>%
         gather(metric,value,-c(sample,Patient.ID,Responder,Diagnosis,Timepoint,Batch,celltype,Material)) %>%
         dplyr::filter(metric=='Inv.Simpson',Responder!='control') %>%
         dplyr::mutate(Responder=ifelse(Responder=='yes','R','NR')), 
       aes(x=Responder,y=value)) +
  geom_boxplot(aes(fill=Responder,color=Responder),outlier.shape=NA) +
  geom_point(size=1,color='gray30') +
  theme_classic() +
  coord_cartesian(clip='off') +
  scale_y_log10() +
  facet_grid(Timepoint~celltype,scales='free_y',switch='y') +
  stat_compare_means(comparisons=list(c('NR','R')),label = "p.format",size=2,bracket.size=.25,tip.length=.01) +
  scale_fill_manual(values=responder_colors) +
  scale_color_manual(values=responder_colors) +
  labs(x='',y='Inv. Simpson',title='') +
  theme(legend.key.size=unit(4,'mm'),
        legend.box='horizontal',
        axis.text.x=element_text(angle=90,hjust=1,vjust=0),
        strip.background=element_blank(),
        strip.placement='outside')
```

```{r clonotype_celltypes_BCR,fig.width=6.5,fig.height=10}
tmp <- list()
for (sample in grep('PBMC',names(pbmc.B),value=TRUE)) {
  sobj <- pbmc.B[[sample]]
  sobj$celltype <- plyr::revalue(sobj$predicted.celltype.l2,ct_map)
  sobj$cloneType <- factor(tolower(gsub('Single','Small',gsub(' .*','',sobj$cloneType))),
                           levels=c('small','medium','large','hyperexpanded'))
  tmp[[sample]] <- occupiedscRepertoire(sobj,
                                        x.axis = "celltype",
                                        exportTable=TRUE) %>%
    dplyr::mutate(sample=gsub('_PBMC','',sample))
}

ggplot(do.call(rbind,tmp),aes(x=celltype,y=value,fill=cloneType)) +
  geom_col() +
  theme_classic() +
  facet_wrap(~sample,scales='free_y',ncol=2,strip.position='left') +
  scale_fill_manual(values=cloneType_colors) +
  theme(axis.text.x=element_text(angle=90,hjust=1,vjust=.5),
        strip.background=element_blank(),
        strip.placement='outside',
        legend.key.size=unit(3,'mm')) +
  labs(x='',y='')
```

```{r Fig_S10D, fig.width=4.5,fig.height=2.}
pbmc_frac %>%
  dplyr::filter(Timepoint %in% c('Vedo_0d','Vedo_6wk')) %>%
  dplyr::group_by(celltype,Patient.ID) %>%
  dplyr::mutate(ns=n()) %>%
  dplyr::filter(ns==2,celltype %in% c('IgD- B cells','IgD+ B cells','Plasmablast')) %>%
  ggplot(aes(x=Timepoint,y=frac,fill=Timepoint,color=Timepoint)) + 
  geom_boxplot(outlier.shape=NA) +
  geom_point(size=.5,color='gray30') +
  geom_line(aes(group=Patient.ID),size=.25,color='gray30') +
  stat_compare_means(aes(group=celltype),comparisons=list(c('Vedo_6wk','Vedo_0d')),label = "p.format", 
                     paired=TRUE,method='wilcox.test',bracket.size=.25,tip.length=.01,size=2) +
  theme_classic() + 
  facet_grid(.~celltype) +
  coord_cartesian(clip='off') +
  scale_fill_manual(values=timepoint_colors,limits=force) +
  scale_color_manual(values=timepoint_colors,limits=force) +
  theme(axis.text.x=element_blank(),
        legend.position='right',
        legend.key.size=unit(3.5,'mm'),
        strip.background=element_blank(),
        strip.placement='outside',
        strip.text=element_text(size=8,hjust=.5,vjust=1))  +
  labs(x='',y='cell fraction',fill='',color='')
```

# Fig. 4 supplements

```{r get_GSE73661}
olink_genes <- c('CXCL8','VEGFA','CD8A','CCL7','GDNF','CDCP1','CD244','IL7','TNFRSF11B','TGFB1','PLAU','IL6','IL17C','CCL2','IL17A','CXCL11',
                 'AXIN1','TNFSF10','CXCL9','CST5','OSM','CXCL1','CCL4','CD6','SCF','IL18','SLAMF1','TGFA','CCL13','CCL11','TNFSF14','IL10RA',
                 'FGF5','MMP1','LIFR','FGF21','CCL19','IL15RA','IL10RB','IL18R1','CD274','CXCL5','TNFSF11','HGF','IL12B','MMP10','IL10','TNF',
                 'CCL23','CD5','CCL3','FLT3LG','CXCL6','CXCL10','EIF4EBP1','SIRT2','CCL28','DNER','S100A12','CD40','IFNG','FGF19','CCL8','CASP8',
                 'CCL25','CX3CL1','TNFRSF9','NTF3','TNFSF12','CCL20','SULT1A1','STAMBP','ADA','LTA','CSF1')

meta <- read.table(file.path(path,'GSE73661','GSE73661_Vedo.txt'),header=1,sep='\t') %>%
  dplyr::mutate(sample=make.names(geo_accession)) %>%
  tibble::rownames_to_column("accession")  %>%
  dplyr::mutate(diagnosis=factor(ifelse(grepl('UC',sample),'UC','control'),levels=c('control','UC')),
                treatment=factor(ifelse(diagnosis=='control','control',
                                        ifelse(grepl('plac',sample),'placebo',
                                               ifelse(grepl('IFX',sample),'IFX',
                                                      ifelse(grepl('vdz4w',sample),'vdz4w','vdz8w')))),
                                 levels=c('control','placebo','IFX','vdz4w','vdz8w')),
                timepoint=factor(ifelse(diagnosis=='control','control',
                                        ifelse(grepl('W0',sample),'w0',
                                               ifelse(grepl('W4|W6',sample),'w6',
                                                      ifelse(grepl('W12',sample),'w12',
                                                             ifelse(grepl('W52',sample),'w52','other'))))),
                                 levels=c('control','other','w0','w6','w12','w52')),
                responder=ifelse(diagnosis=='control','control',ifelse(Response %in% c(0,'Placebo'),'other',Response))) %>%
  dplyr::filter(!grepl('earlier.stop',sample),
                !grepl('_47$',sample),
                !grepl('_73$',sample),
                (diagnosis=='control') | ((timepoint %in% c('w0','w6')) & (treatment=='vdz4w') & (responder!='other'))) %>%
  droplevels()

row.names(meta) <- meta$sample
colnames(meta) <- gsub('.ch1','',colnames(meta))
data <- read.table(file.path(path,'GSE73661','GSE73661.txt'),header=1,sep='\t')
data <- data[,colnames(data) %in% row.names(meta)]

GSE73661 <- SummarizedExperiment(assays=list(data),colData=meta[colnames(data),])
```

show a heatmap of gene expression

```{r Fig_S17C,fig.width=10,fig.height=10}
olink_genes_here <- setdiff(olink_genes, c('SCF','TNFSF12'))

ha <- HeatmapAnnotation(df=colData(GSE73661)[,c('timepoint','responder','diagnosis')],
                        show_legend=TRUE,
                        col=list('timepoint'=c('control'='#000000',
                                               'w0'='#ff7700',
                                               'w6'='#882e72'),
                                 'responder'=c('control'='#000000',
                                               'R'='#db2F26',
                                               'NR'='#34a048'),
                                 'diagnosis'=c('UC'='#1965b0',
                                               'control'='#000000')))

mat <- t(scale(t(assay(GSE73661)[olink_genes_here,])))
Heatmap(mat,
        cluster_columns=TRUE,
        show_row_names=TRUE,
        show_column_names=FALSE,
        top_annotation=ha,
        column_names_gp=gpar(fontsize=6),
        row_names_gp=gpar(fontsize=6),
        name='z-score') 
```

run a few contrasts:

- w6vsw0 @ UC
- UCvscontrol

```{r GSE73661_DE,fig.width=6,fig.height=5}
library(limma)
res <- list()

take <- colData(GSE73661)$diagnosis=='UC'
design <- model.matrix(~0+timepoint,data=colData(GSE73661)[take,])
fit <- eBayes(lmFit(assay(GSE73661)[,take], design))
contrast.matrix <- makeContrasts(w6vsw0=timepointw6-timepointw0,levels=design)
contrast_fit <- eBayes(contrasts.fit(fit, contrast.matrix))
res[['w6vsw0@UC']] <- topTable(contrast_fit, coef='w6vsw0', n=Inf)  %>%
  tibble::rownames_to_column('gene') %>%
  dplyr::filter(gene %in% olink_genes)  %>%
  dplyr::mutate(contrast='w6vsw0@UC')

design <- model.matrix(~0+diagnosis,data=colData(GSE73661))
fit <- eBayes(lmFit(assay(GSE73661), design))
contrast.matrix <- makeContrasts(UCvscontrol=diagnosisUC-diagnosiscontrol,levels=design)
contrast_fit <- eBayes(contrasts.fit(fit, contrast.matrix))
res[['UCvscontrol']] <- topTable(contrast_fit, coef='UCvscontrol', n=Inf)  %>%
  tibble::rownames_to_column('gene') %>%
  dplyr::filter(gene %in% olink_genes)  %>%
  dplyr::mutate(contrast='UCvscontrol')

GSE73661_DE <- do.call(rbind,res) %>%
    dplyr::select(gene,AveExpr,logFC,P.Value,adj.P.Val,contrast)
```

show volcano plots (only for Olink genes!)

```{r Fig_S17A_B,fig.width=4.5,fig.height=4}
for (ct in c('w6vsw0@UC','UCvscontrol')) {
  res <- GSE73661_DE %>% dplyr::filter(contrast==ct)
  pl <-  res %>%
    ggplot(aes(x=logFC,y=-log10(P.Value))) + 
    geom_point(aes(color=(P.Value < .05)),size=.5) + 
    geom_text_repel(data=res %>% dplyr::filter(P.Value < .05),
                    aes(label=gene),segment.size=.25,segment.alpha=.25,color='red') +
    theme_classic() +
    geom_hline(yintercept=-log10(.05),linetype='dashed',size=.25) +
    geom_vline(xintercept=0,linetype='dashed',size=.25) +
    scale_color_manual(values=c('black','red'),guide=FALSE) +
    labs(title=ct)
  print(pl)
}
```

# Fig. 6 + supplements

```{r Fig_6B,fig.width=4.5,fig.height=3.5}
DimPlot(CD4.integrated,label=TRUE) +
  scale_color_brewer(palette='Set3') +
  theme_void() +
  theme(legend.key.size=unit(3,'mm'))
```

```{r CD4_material,fig.width=4.5,fig.height=3.5}
DimPlot(CD4.integrated,group.by='Material') +
  scale_color_brewer(palette='Set1') +
  theme_void() +
  theme(legend.key.size=unit(3,'mm')) +
  labs(title='')
```

```{r Fig_S19D,fig.width=6,fig.height=20,cache=FALSE}
DefaultAssay(CD4.integrated) <- 'RNA'
if (!file.exists(file.path(path,'seurat','CD4_RNA_markers.csv'))) {
  markers <- FindAllMarkers(CD4.integrated, only.pos = TRUE, min.pct = 0.25,
                            thresh.use = 0.25,
                            print.bar=FALSE, verbose=FALSE)
  write.csv(markers,file.path(path,'seurat','CD4_RNA_markers.csv'))
}

markers <- read.csv(file.path(path,'seurat','CD4_RNA_markers.csv'))

top10 <- markers %>% 
  group_by(cluster) %>% 
  top_n(10, avg_log2FC)

DotPlot(CD4.integrated, features = rev(unique(top10$gene))) + 
  theme(axis.text.x=element_text(angle=90,hjust=1,vjust=.5)) +
  labs(x='',y='cluster') + 
  coord_flip()
```

```{r Fig_6F,fig.width=12,fig.height=2,cache=FALSE}
DefaultAssay(CD4.integrated) <- 'RNA'

CD4.integrated$cluster1 <- factor(ifelse(Idents(CD4.integrated)=='1','1','other'),levels=c('1','other'))

# c('CXCR3','CXCR5','CCR2','CCR4','CCR5','CCR6','CCR7','CCR9',
#                 'IL1R1','IL1RAP','IL1RL1','IL2RA','IL4R','IL7R','IL12RB2','IL18RAP','IL18R1',
#                 'IL23R','IFNGR1','IFNGR2','TNFRSF1A','TNFRSF1B','TNFRSF4',
tcell_genes <- c('GPR15','ITGA4','ITGB1','ITGB7',
                 'AREG','CD40LG','CD38','CD69','CTLA4','GZMA','GZMB','HLA-DRA','ICOS','LAG3',
                 'MKI67','PDCD1','SELL','SLAMF6','HAVCR2','TIGIT',
                 'AHR','BATF','BCL6','CREM','EOMES','FOSB','GATA3','IKZF2','JUN','JUNB','MAF',
                 'NFATC1','PRDM1','RORA','RORC','TBX21','TCF7','TOX','ZBTB7B',
                 'CSF2','IFNG','IL2','IL4','IL5','IL9','IL10','IL17A','IL22','TNF')

DotPlot(CD4.integrated, features=tcell_genes,
        group.by='cluster1') +
  scale_color_viridis_c() +
  scale_y_discrete(limits=rev) +
  theme(axis.text.x=element_text(angle=90,hjust=1,vjust=.5)) +
  labs(x='',y='')  
```

```{r Fig_6G,fig.width=6,fig.height=2,cache=FALSE}
DefaultAssay(CD4.integrated) <- 'orig_ADT'
DotPlot(CD4.integrated, features = c('CD25','CD127','PD-1','CD38','HLA-DR'),
        group.by='cluster1') + 
  scale_color_viridis_c() +
  scale_y_discrete(limits=rev) +
  theme(axis.text.x=element_text(angle=90,hjust=1,vjust=.5),
        legend.key.size=unit(3,'mm'),
        legend.box='horizontal') +
  labs(x='',y='')
```

```{r Fig_6C,fig.width=4,fig.height=3,cache=FALSE}
DefaultAssay(CD4.integrated) <- 'RNA'
FetchData(CD4.integrated, c('UMAP_1','UMAP_2','MKI67')) %>%
  dplyr::arrange(MKI67) %>%
  ggplot(aes(x=UMAP_1,y=UMAP_2,color=MKI67)) +
  geom_point(size=.25) +
  theme_void() + 
  coord_fixed() +
  scale_color_gradient(low='lightgoldenrod1',high='darkred', 
                       na.value='gray80', 
                       oob=scales::squish,
                       guide=guide_colorbar(title.position='right',
                                            title.theme=element_text(angle=90,hjust=.5,vjust=.5))) + 
  theme(legend.key.size=unit(3.5,'mm')) +
  labs(color='MKI67')
```

```{r Fig_6E, fig.width=2,fig.height=2,cache=FALSE}
library(gtools)
CD4.integrated[['cluster']] <- Idents(CD4.integrated)

cols <- c('sample','orig.ident','cluster','Patient.ID','Responder','Diagnosis','Material','Timepoint','Batch')
CD4_frac <- CD4.integrated@meta.data[,cols] %>%
  dplyr::mutate(celltype=cluster) %>%
  dplyr::group_by(sample,Patient.ID,Responder,Diagnosis,Material,Timepoint,Batch,celltype) %>%
  dplyr::summarise(n=n()) %>%
  dplyr::mutate(frac=n/sum(n))

CD4_frac$Timepoint <- factor(CD4_frac$Timepoint, levels=c('ctrl','Vedo_0d','Vedo_6wk'))
CD4_frac$Responder <- revalue(CD4_frac$Responder, c('yes'='R','no'='NR'))

CD4_frac %>%
  dplyr::filter(Timepoint!='ctrl',celltype=='1',Timepoint=='Vedo_0d') %>%
  ggplot(aes(x=Responder,y=frac,fill=Responder,color=Responder)) + 
  geom_boxplot(outlier.shape=NA) +
  geom_point(size=1,aes(shape=Material),color='gray30') +
  stat_compare_means(aes(group=Patient.ID),comparisons=list(c('NR','R')),label = "p.format", 
                     method='wilcox.test',bracket.size=.25,tip.length=.01,size=2) +
  theme_classic() + 
  coord_cartesian(clip='off') +
  scale_fill_manual(values=responder_colors,limits=force) +
  scale_color_manual(values=responder_colors,limits=force) +
  scale_shape_manual(values=c(1,4)) +
  theme(axis.text.x=element_blank(),
        legend.position='right',
        legend.key.size=unit(3.5,'mm'),
        strip.background=element_blank(),
        strip.placement='outside')  +
  labs(x='',y='cell fraction',fill='',color='',shape='')
```

```{r Fig_S19E, fig.width=5,fig.height=2.5}
CD4_frac %>%
    dplyr::filter(Timepoint!='ctrl', celltype %in% seq(1,6)) %>%
    ggplot(aes(x=Responder,y=frac,fill=Responder,color=Responder)) + 
    geom_boxplot(outlier.shape=NA) +
    geom_point(size=1,aes(shape=Material),color='gray30') +
    stat_compare_means(aes(group=Patient.ID),comparisons=list(c('NR','R')),label = "p.format", 
                       method='wilcox.test',bracket.size=.25,tip.length=.01,size=2) +
    theme_classic() + 
    coord_cartesian(clip='off') +
    facet_grid(Timepoint~celltype,scales='free_y',switch='y') +
    scale_fill_manual(values=responder_colors,limits=force) +
    scale_color_manual(values=responder_colors,limits=force) +
    scale_shape_manual(values=c(1,4)) +
    theme(axis.text.x=element_blank(),
          legend.position='right',
          legend.key.size=unit(3.5,'mm'),
          strip.background=element_blank(),
          strip.placement='outside')  +
    labs(x='',y='cell fraction',fill='',color='',shape='')

CD4_frac %>%
    dplyr::filter(Timepoint!='ctrl', celltype %in% seq(7,12)) %>%
    ggplot(aes(x=Responder,y=frac,fill=Responder,color=Responder)) + 
    geom_boxplot(outlier.shape=NA) +
    geom_point(size=1,aes(shape=Material),color='gray30') +
    stat_compare_means(aes(group=Patient.ID),comparisons=list(c('NR','R')),label = "p.format", 
                       method='wilcox.test',bracket.size=.25,tip.length=.01,size=2) +
    theme_classic() + 
    coord_cartesian(clip='off') +
    facet_grid(Timepoint~celltype,scales='free_y',switch='y') +
    scale_fill_manual(values=responder_colors,limits=force) +
    scale_color_manual(values=responder_colors,limits=force) +
    scale_shape_manual(values=c(1,4)) +
    theme(axis.text.x=element_blank(),
          legend.position='right',
          legend.key.size=unit(3.5,'mm'),
          strip.background=element_blank(),
          strip.placement='outside')  +
    labs(x='',y='cell fraction',fill='',color='',shape='')
```

```{r Fig_6D, fig.width=3,fig.height=2,cache=FALSE}
library(gtools)
CD4.integrated[['frac_MKI67']] <- ifelse(CD4.integrated@assays$RNA@counts['MKI67',] > 0,'MKI67+','MKI67-')

cols <- c('sample','orig.ident','cluster','Patient.ID','Responder','Diagnosis','Material','Timepoint','Batch','frac_MKI67')
CD4_frac <- CD4.integrated@meta.data[,cols] %>%
  dplyr::mutate(celltype=cluster) %>%
  dplyr::group_by(sample,Patient.ID,Responder,Diagnosis,Material,Timepoint,Batch,celltype) %>%
  dplyr::summarise(frac=mean(frac_MKI67=='MKI67+'))

CD4_frac %>%
  dplyr::filter(Timepoint!='ctrl') %>%
  ggplot(aes(x=celltype,y=frac,fill=celltype,color=celltype)) + 
  geom_boxplot(outlier.shape=NA) +
  geom_point(size=.5,color='gray30') +
  theme_classic() + 
  coord_cartesian(clip='off') +
  scale_fill_brewer(palette='Set3') +
  scale_color_brewer(palette='Set3') +
  scale_shape_manual(values=c(1,4)) +
  theme(axis.text.x=element_text(angle=90,hjust=1,vjust=.5),
        legend.position='none',
        legend.key.size=unit(3.5,'mm'),
        strip.background=element_blank(),
        strip.placement='outside')  +
  labs(x='',y='fraction MKI67+',fill='',color='')
```

```{r sessionInfo}
sessionInfo()
```

### 2023-06-30 scRNAseq/processing.Rmd
add scRNAseq figures

---
title: "data processing"
output: 
  html_document:
    df_print: paged
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, cache=TRUE, cache.lazy=FALSE, message=FALSE, warning=FALSE)
library(sctransform)
library(Seurat)
library(SeuratDisk)
library(ggplot2)
library(cowplot)
library(ggrepel)
library(dplyr)
library(scRepertoire)
library(tidyr)
library(cluster)
library(plyr)
library(ComplexHeatmap)
library(dendextend)
options(future.globals.maxSize = 10 * 1024 * 1024^2)
path <- getwd()
```

we follow the [Seurat vignette](https://satijalab.org/seurat/v4.0/reference_mapping.html) and use their PBMC data as reference

```{r get_reference, fig.width=6,fig.height=5}
reference <- LoadH5Seurat(file.path(path, 'seurat', "pbmc_multimodal.h5seurat"))
```

load PBMC data; take all cells with less than 10% mitochondrial RNA and between 250 and 5000 genes

- K456985_PBMC
- K783496_PBMC
- K693104_PBMC
- K955601_PBMC
- P1_1_PBMC 
- P1_2_PBMC
- P3_1_PBMC
- P3_2_PBMC
- P8_1_PBMC
- P8_2_PBMC 
- P22_1_PBMC
- P22_2_PBMC
- P23_1_PBMC
- P23_2_PBMC
- P25_1_PBMC
- P25_2_PBMC
- P27_1_PBMC
- P27_2_PBMC
- P28_1_PBMC
- P28_2_PBMC
- P22_P23_K783496_CD4
- P8_P3_K456985_CD4
- P27_P28_K693104_CD4
- P12_1_PBMC
- P12_2_PBMC
- P26_1_PBMC
- P26_2_PBMC
- K496398_PBMC
- P12_P26_K496398_CD4

```{r get_data}
samples <- c('K456985_PBMC','K783496_PBMC','P1_1_PBMC','P22_1_PBMC','P22_P23_K783496_CD4','P23_2_PBMC','P25_2_PBMC','P27_2_PBMC',
             'P28_1_PBMC','P3_1_PBMC','P8_1_PBMC','P8_P3_K456985_CD4','K693104_PBMC','K955601_PBMC','P1_2_PBMC','P22_2_PBMC',
             'P23_1_PBMC','P25_1_PBMC','P27_1_PBMC','P27_P28_K693104_CD4','P28_2_PBMC','P3_2_PBMC','P8_2_PBMC',
             'P12_1_PBMC','P12_2_PBMC','P26_1_PBMC','P26_2_PBMC','K496398_PBMC','P12_P26_K496398_CD4')

pbmc <- list()
for (sample in samples) {
  dge <- Read10X_h5(file.path(path,'cellranger',paste0(sample,'_filtered_feature_bc_matrix.h5')))
  sobj <- CreateSeuratObject(counts = dge[['Gene Expression']], project=sample)
  sobj[['pct.mito']] <- PercentageFeatureSet(sobj, pattern='^MT-')
  sobj <- subset(sobj, (pct.mito < 10) & (nFeature_RNA <= 5000) & (nFeature_RNA >= 250))
  capture.output(sobj <- SCTransform(sobj, verbose = FALSE),file='/dev/null')
  cells <- colnames(sobj)
  row.names(dge[['Antibody Capture']]) <- gsub('_TotalSeqC','',row.names(dge[['Antibody Capture']]))
  hashtags <- grepl('^Hashtag_[1234]',row.names(dge[['Antibody Capture']]))
  sobj[["orig_ADT"]] <- CreateAssayObject(counts = dge[['Antibody Capture']][!hashtags,cells,drop=FALSE])
  sobj <- NormalizeData(sobj, normalization.method='CLR', assay='orig_ADT', verbose=FALSE, margin=2)
  if (grepl('CD4',sample)) {
    hts <- c("Hashtag_1","Hashtag_2","Hashtag_3","Hashtag_4")
    sobj[["HTO"]] <- CreateAssayObject(counts = dge[['Antibody Capture']][hts,cells,drop=FALSE])
    sobj <- NormalizeData(sobj, normalization.method='CLR', assay='HTO', verbose=FALSE)
    sobj <- HTODemux(sobj, assay='HTO', verbose=FALSE)
  }
  pbmc[[sample]] <- sobj
}
```

get some QC metrics

```{r QC,fig.width=10,fig.height=5}
md <- do.call(rbind, lapply(names(pbmc), function(x) pbmc[[x]]@meta.data[,c('nCount_RNA','nFeature_RNA','pct.mito')] %>%
                              dplyr::mutate(orig.ident=x))) %>%
  dplyr::group_by(orig.ident) %>%
  dplyr::mutate(orig.ident=paste0(orig.ident,' (n=',n(),')')) %>%
  gather(metric,value,-orig.ident)

ggplot(md,aes(x=orig.ident,y=value,fill=orig.ident)) + 
  #geom_boxplot(outlier.shape=NA) + 
  geom_violin(scale='width') +
  facet_wrap(~metric,scales='free_y') + 
  theme_classic() + 
  theme(axis.text.x=element_text(angle=90,hjust=1,vjust=.5),
        legend.position='none') + 
  labs(x='',y='')

```

project our cells onto this reference embedding using RNA data; predict cell type labels for the level 1, level 2 and level 3 clustering of the reference, and also antibody data ("what ADT data would be expect if our samples are like the reference?")

```{r project}
for (sample in samples) {
  anchors <- FindTransferAnchors(
    reference = reference,
    query = pbmc[[sample]],
    normalization.method = "SCT",
    reference.reduction = "spca",
    dims = 1:50,
    verbose=FALSE)
  pbmc[[sample]] <- MapQuery(
    anchorset = anchors,
    query = pbmc[[sample]],
    reference = reference,
    refdata = list(
      celltype.l1 = "celltype.l1",
      celltype.l2 = "celltype.l2",
      celltype.l3 = "celltype.l3",
      predicted_ADT = "ADT"
    ),
    reference.reduction = "spca", 
    reduction.model = "wnn.umap",
    verbose=FALSE)
}
```

identify and remove putative doublets (shown in pink), and cells with prediction score < .75

```{r doublets,fig.width=10,fig.height=12}
source(file.path(path,'my_doubletFinder.R'))
pbmc <- lapply(pbmc, function(sobj) {
  DefaultAssay(sobj) <- 'RNA'
  Idents(sobj) <- 'predicted.celltype.l1'
  homotypic.prop <- modelHomotypic(Idents(sobj))
  nExp_poi <- round(0.05*nrow(sobj@meta.data)) 
  nExp_poi.adj <- round(nExp_poi*(1-homotypic.prop))
  tmp <- my_doubletFinder(sobj, PCs = 1:10, pN = 0.25, pK = 0.09, nExp = nExp_poi, reuse.pANN = FALSE)
  tmp <- my_doubletFinder(tmp, PCs = 1:10, pN = 0.25, pK = 0.09, nExp = nExp_poi.adj, reuse.pANN = "pANN")
  AddMetaData(sobj, tmp@meta.data[,'DF.classifications',drop=FALSE])
})

do.call(rbind,lapply(pbmc, function(sobj) table(sobj$DF.classifications))) %>%
  DT::datatable()

plots <- lapply(names(pbmc), function(sample)
  DimPlot(pbmc[[sample]], group.by='DF.classifications',
          cols=c('Singlet'='gray',
                 'Doublet'='deeppink')) +
    theme_void() + 
    labs(title=sample) + 
    theme(legend.position='none'))

plot_grid(plotlist=plots,ncol=5,align='vh')

pbmc.sub <- lapply(pbmc, function(x) subset(x, subset=(DF.classifications=='Singlet') & (predicted.celltype.l1.score >= .75)))
```

```{r save_samples}
for (sample in samples) {
  saveRDS(pbmc.sub[[sample]], file.path(path,'seurat',paste0(sample,'.rds')))
}
```

aggregate and demultiplex hashtags

```{r aggregate}
pbmc <- list()
for (lib in grep('PBMC',samples,value=TRUE)) {
  pbmc[[lib]] <- readRDS(file.path(path,'seurat',paste0(lib,'.rds')))
  pbmc[[lib]]$sample <- lib
}

study <- read.csv(file.path(path,'sample_sheet','s_hegazy_IBD.txt'),
                 sep='\t') %>%
  dplyr::mutate(Sample.Name=gsub('p','P',Sample.Name))

assay <- read.csv(file.path(path,'sample_sheet','a_hegazy_IBD_totalseq_nucleotide_sequencing.txt'),
                 sep='\t') %>%
  dplyr::mutate(Sample.Name=gsub('p','P',Sample.Name),
                Library.Name=gsub('p','P',Library.Name))

for (lib in grep('CD4',samples,value=TRUE)) {
  sobj <- readRDS(file.path(path,'seurat',paste0(lib,'.rds')))
  vireo <- read.table(file.path(path,'vireo',paste0(lib,'_donor_ids.tsv')),
                      header=1,row.names=1,as.is=TRUE)
  colnames(vireo) <- paste0('vireo_',colnames(vireo))
  sobj <- AddMetaData(sobj, vireo)  
  sobj$vireo_donor_id[is.na(sobj$vireo_donor_id)] <- 'none'
  samples <- assay %>% 
    dplyr::filter(Library.Name==lib) %>% 
    dplyr::select(Parameter.Value.Multiplex.sample.tag.,Sample.Name) %>%
    dplyr::mutate(ht=Parameter.Value.Multiplex.sample.tag.) %>%
    dplyr::mutate(ht=ifelse(ht=='','Negative',gsub('_','-',ht))) 
  samples <- setNames(as.character(samples$Sample.Name), samples$ht)
  sobj$sample <- revalue(sobj$hash.ID, samples)
  print(table(sobj$sample,sobj$vireo_donor_id))
  sobjs <- SplitObject(sobj,split.by='sample')
  for (sample in samples) {
    pbmc[[sample]] <- sobjs[[sample]]
  }
}
```

add clonotype information

add clonotype information (TCR and BCR where available; now with `filterMulti=TRUE`)

```{r get_clonotypes}
cols <- c('sample','orig.ident')
samples <- assay %>% 
  dplyr::filter(Library.Name %in% libraries) %>%
  dplyr::select(Library.Name,Sample.Name)

bcs <- lapply(pbmc, function(x) gsub('.*\\.','',Cells(x)))

tcr_list <- list()
bcr_list <- list()
for (lib in libraries) {
  infile <- file.path(path,'cellranger',paste0(lib,'vdj_t_filtered_contig_annotations.csv'))
  if (file.exists(infile)) {
    tcr <- read.csv(infile,
                    stringsAsFactors=FALSE) %>%
      dplyr::mutate(is_cell=is_cell=='true',
                    high_confidence=high_confidence=='true',
                    full_length=full_length=='true',
                    productive=productive=='true')
    for (sample in samples %>% dplyr::filter(Library.Name==lib) %>% dplyr::pull(Sample.Name)) {
      tcr_list[[sample]] <- tcr %>% dplyr::filter(barcode %in% bcs[[sample]])
    }
  }
  infile <- file.path(path,'cellranger',paste0(lib,'vdj_b_filtered_contig_annotations.csv'))
  if (file.exists(infile)) {
    bcr <- read.csv(infile,
                    stringsAsFactors=FALSE) %>%
      dplyr::mutate(is_cell=is_cell=='true',
                    high_confidence=high_confidence=='true',
                    full_length=full_length=='true',
                    productive=productive=='true')
    for (sample in samples %>% dplyr::filter(Library.Name==lib) %>% dplyr::pull(Sample.Name)) {
      bcr_list[[sample]] <- bcr %>% dplyr::filter(barcode %in% bcs[[sample]])
    }
  }
}

tcr_combined <- sapply(which(sapply(tcr_list, function(x) !is.null(x))),
                       function(i) combineTCR(tcr_list[[i]],
                                              samples=c(names(tcr_list)[[i]]),
                                              ID=c(names(tcr_list)[[i]]),
                                              cells='T-AB',
                                              filterMulti=TRUE))
names(tcr_combined) <- names(tcr_list)[sapply(tcr_list, function(x) !is.null(x))]

bcr_combined <- sapply(which(sapply(bcr_list, function(x) !is.null(x))),
                       function(i) combineBCR(bcr_list[[i]],
                                              samples=c(names(bcr_list)[[i]]),
                                              ID=c(names(bcr_list)[[i]])))
names(bcr_combined) <- names(bcr_list)[sapply(bcr_list, function(x) !is.null(x))]
```

```{r save_all}
saveRDS(pbmc,file.path(path,'seurat','PBMC_combined.rds'))
saveRDS(tcr_combined,file.path(path,'seurat','TCR_combined.rds'))
saveRDS(bcr_combined,file.path(path,'seurat','BCR_combined.rds'))
```

```{r sessionInfo}
sessionInfo()
```

### 2023-06-30 scRNAseq/recluster_CD4.R

@@ -1,0 +1,62 @@
+library(Seurat)
+library(dplyr)
+
+path <- getwd()
+pbmc <- readRDS(file.path(path,'seurat','PBMC_combined.rds'))
+
+study <- read.csv(file.path(path,'sample_sheet','s_hegazy_IBD.txt'),
+                 sep='\t') %>%
+  dplyr::mutate(Sample.Name=gsub('p','P',Sample.Name))
+assay <- read.csv(file.path(path,'sample_sheet','a_hegazy_IBD_totalseq_nucleotide_sequencing.txt'),
+                 sep='\t') %>%
+  dplyr::mutate(Sample.Name=gsub('p','P',Sample.Name),
+                Library.Name=gsub('p','P',Library.Name))
+
+CD4.list <- lapply(pbmc, function(x) subset(x, predicted.celltype.l2 %in% c('CD4 CTL','CD4 Proliferating','CD4 TCM','CD4 TEM','Treg')))
+for (nn in names(CD4.list)) {
+  DefaultAssay(CD4.list[[nn]]) <- 'SCT'
+}
+CD4.list <- lapply(CD4.list, RunPCA, verbose=FALSE)
+
+features <- SelectIntegrationFeatures(object.list = CD4.list, nfeatures = 2000, verbose=FALSE)
+features <- features[!grepl('MT-|TR[AB][VDJ][0-9]|^IG[HLK][VDJ][0-9]',features)]
+
+CD4 <- merge(CD4.list[[1]],CD4.list[2:length(CD4.list)],add.cell.ids=names(CD4.list)) %>%
+  RunPCA(features=features, verbose = FALSE) %>%
+  RunUMAP(reduction = "pca", dims = 1:20, verbose=FALSE) %>%
+  FindNeighbors(dims = 1:20, verbose = FALSE) %>%
+  FindClusters(verbose = FALSE)
+
+CD4@meta.data <- CD4@meta.data %>%
+  tibble::rownames_to_column('cell') %>%
+  dplyr::left_join(study %>% 
+                     dplyr::mutate(Patient.ID=gsub('[P]*([0-9]*)[\\/_][12]','P\\1',Characteristics.Patient.ID.),
+                                   Responder=Characteristics.Responder.,
+                                   Diagnosis=Characteristics.Diagnosis.,
+                                   Batch=gsub(' ','',gsub('_scRNAseq','',Parameter.Value.Method.)),
+                                   Timepoint=ifelse(grepl('_1_',Sample.Name),'Vedo_0d',
[…]

### 2023-06-30 scRNAseq/paper_figures.Rmd
update

@@
 ---
-title: "20230627 paper figures"
+title: "paper figures"
 output: 
@@
 - K955601_PBMC
-- P1_1_PBMC (re-sequenced)
+- P1_1_PBMC
 - P1_2_PBMC
@@
 - P8_1_PBMC
-- P8_2_PBMC (re-sequenced)
+- P8_2_PBMC
 - P22_1_PBMC

### 2023-09-08 scRNAseq/paper_figures.Rmd
update figures

@@
 library(Binarize)
+library(limma)
 options(future.globals.maxSize = 10 * 1024 * 1024^2)
@@
 
-```{r normalize}
+```{r normalize, cache=FALSE}
 for (nn in names(pbmc)) {
@@
 
-# Fig 1 + supplements
-
 show all data in the reference embedding; match cluster granularity to CyTOF data;
 
-```{r Fig_1C,fig.width=7,fig.height=5}
+```{r Fig_S2A,fig.width=4.5,fig.height=3.5}
 ct_map <- c('CD8 TCM'='Mmry CD8+ T cells',
@@
 
-```{r Fig_S3B,fig.width=4.5,fig.height=3.5}
+```{r Fig_S2B,fig.width=4.5,fig.height=3.5}
 ggplot(pbmc_meta %>% dplyr::sample_frac(1L), 
@@
 
-```{r Fig_S3A,fig.width=7,fig.height=4.5, cache=FALSE}
+```{r Fig_S2C,fig.width=7,fig.height=4.5, cache=FALSE}
 ABs <- c("CD3","CD8a","CD40","CD19","CD20","CD45RA","CD38","CD69","HLA-DR",
@@
 
-```{r Fig_S3C, fig.width=7,fig.height=4}
+```{r Fig_S2D, fig.width=7,fig.height=4}
 #library(lme4)
@@
 
-```{r Fig_1E,fig.width=4.5,fig.height=4, cache=FALSE}
+```{r Fig_S3A,fig.width=4.5,fig.height=4, cache=FALSE}
 for (nn in grep('PBMC',names(pbmc),value=TRUE)) {
@@
 
-```{r Fig_S8A,fig.width=7,fig.height=3}
+```{r Fig_S3D,fig.width=7,fig.height=3}
 RNA %>% 
@@
 
-# Fig 2
-
 differential expression of ITGA4 and ITGB7
 
-```{r Fig_2A,fig.width=5,fig.height=4}
+```{r Fig_S3B,fig.width=5,fig.height=4}
 RNA <- do.call(rbind,
@@
 
-```{r Fig_2B,fig.width=3.75,fig.height=4, cache=FALSE}
+```{r Fig_S3C,fig.width=3.75,fig.height=4, cache=FALSE}
 for (nn in grep('PBMC',names(pbmc),value=TRUE)) {
@@
 
-# Fig. 3 + supplements
-
 show where we detect clonotypes (not yet filtered for CD4/CD8 T cells!)
 
-```{r Fig_3B,fig.width=4.5,fig.height=3.5}
+```{r Fig_S4B,fig.width=4.5,fig.height=3.5, cache=FALSE}
 pbmc.T <- lapply(names(pbmc), function(sample) {
@@
 
-```{r clonotype_celltypes_TCR,fig.width=6.5,fig.height=10}
-tmp <- list()
-for (sample in grep('PBMC',names(pbmc.T),value=TRUE)) {
-  sobj <- pbmc.T[[sample]]
-  sobj$celltype <- plyr::revalue(sobj$predicted.celltype.l2,ct_map)
-  sobj$cloneType <- factor(tolower(gsub('Single','Small',gsub(' .*','',sobj$cloneType))),
-                           levels=c('small','medium','large','hyperexpanded'))
-  tmp[[sample]] <- occupiedscRepertoire(sobj,
-                                        x.axis = "celltype", 
-                                        exportTable=TRUE) %>%
-    dplyr::mutate(sample=gsub('_PBMC','',sample))
-}
-
-ggplot(do.call(rbind,tmp),aes(x=celltype,y=value,fill=cloneType)) + 
-  geom_col() + 
-  theme_classic() +
-  facet_wrap(~sample,scales='free_y',ncol=2,strip.position='left') +
-  scale_fill_manual(values=cloneType_colors) +
-  theme(axis.text.x=element_text(angle=90,hjust=1,vjust=.5),
-        strip.background=element_blank(),
-        strip.placement='outside',
-        legend.key.size=unit(3,'mm')) +
-  labs(x='',y='')
-```
+<!-- ```{r clonotype_celltypes_TCR,fig.width=6.5,fig.height=10} -->
+<!-- tmp <- list() -->
+<!-- for (sample in grep('PBMC',names(pbmc.T),value=TRUE)) { -->
+<!--   sobj <- pbmc.T[[sample]] -->
+<!--   sobj$celltype <- plyr::revalue(sobj$predicted.celltype.l2,ct_map) -->
+<!--   sobj$cloneType <- factor(tolower(gsub('Single','Small',gsub(' .*','',sobj$cloneType))), -->
+<!--                            levels=c('small','medium','large','hyperexpanded')) -->
+<!--   tmp[[sample]] <- occupiedscRepertoire(sobj, -->
+<!--                                         x.axis = "celltype",  -->
+<!--                                         exportTable=TRUE) %>% -->
+<!--     dplyr::mutate(sample=gsub('_PBMC','',sample)) -->
+<!-- } -->
+
+<!-- ggplot(do.call(rbind,tmp),aes(x=celltype,y=value,fill=cloneType)) +  -->
+<!--   geom_col() +  -->
+<!--   theme_classic() + -->
+<!--   facet_wrap(~sample,scales='free_y',ncol=2,strip.position='left') + -->
+<!--   scale_fill_manual(values=cloneType_colors) + -->
+<!--   theme(axis.text.x=element_text(angle=90,hjust=1,vjust=.5), -->
+<!--         strip.background=element_blank(), -->
+<!--         strip.placement='outside', -->
+<!--         legend.key.size=unit(3,'mm')) + -->
+<!--   labs(x='',y='') -->
+<!-- ``` -->
 
@@
 
-```{r Fig_3C,fig.width=6,fig.height=2}
+```{r Fig_S4C,fig.width=6,fig.height=2}
 cols <- c('sample','orig.ident','predicted.celltype.l2')
@@
                                    Batch=gsub(' ','',gsub('_scRNAseq','',Parameter.Value.Method.)),
-                                   Timepoint=ifelse(grepl('_1_',Sample.Name),'Vedo_0d',
+                                   Timepoint =ifelse(grepl('_1_',Sample.Name),'Vedo_0d',
                                                     ifelse(grepl('_2_',Sample.Name),'Vedo_6wk','ctrl'))) %>%
@@
 
-```{r Fig_S11F,fig.width=6,fig.height=3}
-pl <- ggplot(df %>% 
-               dplyr::filter(Timepoint!='ctrl',celltype!='Tregs') %>%
-               gather(metric,value,-c(sample,Patient.ID,Responder,Diagnosis,Timepoint,Batch,celltype)) %>%
-               dplyr::filter(metric=='Inv.Simpson') %>%
-               dplyr::arrange(Responder,celltype,Timepoint,Patient.ID),
-             aes(x=Responder,y=value)) +
-  geom_boxplot(aes(fill=Responder,color=Responder),outlier.shape=NA) +
-  geom_point(size=1,color='gray30') +
-  theme_classic() +
-  coord_cartesian(clip='off') +
-  scale_y_log10() +
-  facet_grid(Timepoint ~ celltype,scales='free_y',switch='y') +
-  stat_compare_means(comparisons=list(c('NR','R')),label = "p.format",size=2,bracket.size=.25,tip.length=.01) +
-  scale_fill_manual(values=responder_colors,limits=force) +
-  scale_color_manual(values=responder_colors,limits=force) +
-  labs(x='',y='Inv. Simpson',title='') +
-  theme(legend.key.size=unit(4,'mm'),
-        legend.box='horizontal',
-        axis.text.x=element_blank(),
-        strip.background=element_blank(),
-        strip.text=element_text(size=8),
-        strip.placement='outside')
-
-pg <- ggplotGrob(pl)
-for(i in which(grepl("strip-t", pg$layout$name))){
-  pg$grobs[[i]]$layout$clip <- "off"
-}
-
-grid::grid.draw(pg)
-#pl
-```
+<!-- ```{r Fig_S11F,fig.width=6,fig.height=3} -->
+<!-- pl <- ggplot(df %>%  -->
+<!--                dplyr::filter(Timepoint!='ctrl',celltype!='Tregs') %>% -->
+<!--                gather(metric,value,-c(sample,Patient.ID,Responder,Diagnosis,Timepoint,Batch,celltype)) %>% -->
+<!--                dplyr::filter(metric=='Inv.Simpson') %>% -->
+<!--                dplyr::arrange(Responder,celltype,Timepoint,Patient.ID), -->
+<!--              aes(x=Responder,y=value)) + -->
+<!--   geom_boxplot(aes(fill=Responder,color=Responder),outlier.shape=NA) + -->
+<!--   geom_point(size=1,color='gray30') + -->
+<!--   theme_classic() + -->
+<!--   coord_cartesian(clip='off') + -->
+<!--   scale_y_log10() + -->
+<!--   facet_grid(Timepoint ~ celltype,scales='free_y',switch='y') + -->
+<!--   stat_compare_means(comparisons=list(c('NR','R')),label = "p.format",size=2,bracket.size=.25,tip.length=.01) + -->
+<!--   scale_fill_manual(values=responder_colors,limits=force) + -->
+<!--   scale_color_manual(values=responder_colors,limits=force) + -->
+<!--   labs(x='',y='Inv. Simpson',title='') + -->
+<!--   theme(legend.key.size=unit(4,'mm'), -->
+<!--         legend.box='horizontal', -->
+<!--         axis.text.x=element_blank(), -->
+<!--         strip.background=element_blank(), -->
+<!--         strip.text=element_text(size=8), -->
+<!--         strip.placement='outside') -->
+
+<!-- pg <- ggplotGrob(pl) -->
+<!-- for(i in which(grepl("strip-t", pg$layout$name))){ -->
+<!--   pg$grobs[[i]]$layout$clip <- "off" -->
+<!-- } -->
+
+<!-- grid::grid.draw(pg) -->
+<!-- #pl -->
+<!-- ``` -->
 
@@
 
-```{r Figs_3E_3F_S11B_S11C_S11D,fig.width=2.5,fig.height=2.5, cache=FALSE}
+```{r Fig_S4E_S4H_S5C_S5D,fig.width=2.5,fig.height=2.5, cache=FALSE}
 library(Binarize)
@@
 
-for (ss in c('CCR9','GPR15','Integrin_b7','ITGA4','ITGAE','ITGB7')) {
+for (ss in c('GPR15','Integrin_b7','ITGA4','ITGAE','ITGB7')) {
   for (ct in c('Mmry CD4','Mmry CD8')) {
@@
 
-```{r Fig_3H,fig.width=2.5,fig.height=2}
+```{r Fig_S4H,fig.width=2.5,fig.height=2}
 ADT %>% 
@@
 
-```{r Fig_3G, fig.width=2.5,fig.height=2.}
+```{r Fig_S4G, fig.width=2.5,fig.height=2.}
 pbmc_frac %>%
@@
 
-```{r Fig_S11E, fig.width=2.5,fig.height=2.}
+```{r Fig_S5E, fig.width=2.5,fig.height=2.}
 pbmc_frac %>%
@@
 
-```{r Fig_3D,fig.width=3,fig.height=2.5}
+```{r Fig_S4D,fig.width=3,fig.height=2.5}
 cols <- c('sample','orig.ident','predicted.celltype.l2')
@@
 
-```{r Fig_S11A,fig.width=6,fig.height=4}
+```{r Fig_S4D_S5B,fig.width=4,fig.height=4}
 dfs <- list()
-for (ct in c('CD4 Naive','CD4 Proliferating','CD4 TCM','CD4 TEM',
-             'CD8 Naive','CD8 Proliferating','CD8 TCM','CD8 TEM')) {
+for (ct in c('CD4 TCM','CD4 TEM',
+             'CD8 TCM','CD8 TEM')) {
   bcs <- pbmc_md$cell[(pbmc_md$celltype==ct) & (pbmc_md$Material=='PBMC')]
@@
 
-```{r Fig_S10B,fig.width=4.5,fig.height=3.5,cache=FALSE}
+```{r Fig_S5F,fig.width=4.5,fig.height=3.5,cache=FALSE}
 pbmc.B <- lapply(names(bcr_combined), function(sample) {
@@
 
-```{r Fig_S10C,fig.width=4.5,fig.height=3}
+```{r Fig_S5G,fig.width=4.5,fig.height=3}
 dfs <- list()
@@
 
-```{r Fig_S10F,fig.width=4.5,fig.height=3}
-ggplot(df %>% 
-         dplyr::filter(Diagnosis=='UC') %>%
-         gather(metric,value,-c(sample,Patient.ID,Responder,Diagnosis,Timepoint,Batch,celltype,Material)) %>%
-         dplyr::filter(metric=='Inv.Simpson',Responder!='control') %>%
-         dplyr::mutate(Responder=ifelse(Responder=='yes','R','NR')), 
-       aes(x=Responder,y=value)) +
-  geom_boxplot(aes(fill=Responder,color=Responder),outlier.shape=NA) +
-  geom_point(size=1,color='gray30') +
-  theme_classic() +
-  coord_cartesian(clip='off') +
-  scale_y_log10() +
-  facet_grid(Timepoint~celltype,scales='free_y',switch='y') +
-  stat_compare_means(comparisons=list(c('NR','R')),label = "p.format",size=2,bracket.size=.25,tip.length=.01) +
-  scale_fill_manual(values=responder_colors) +
-  scale_color_manual(values=responder_colors) +
-  labs(x='',y='Inv. Simpson',title='') +
-  theme(legend.key.size=unit(4,'mm'),
-        legend.box='horizontal',
-        axis.text.x=element_text(angle=90,hjust=1,vjust=0),
-        strip.background=element_blank(),
-        strip.placement='outside')
-```
-
-```{r clonotype_celltypes_BCR,fig.width=6.5,fig.height=10}
-tmp <- list()
-for (sample in grep('PBMC',names(pbmc.B),value=TRUE)) {
-  sobj <- pbmc.B[[sample]]
-  sobj$celltype <- plyr::revalue(sobj$predicted.celltype.l2,ct_map)
-  sobj$cloneType <- factor(tolower(gsub('Single','Small',gsub(' .*','',sobj$cloneType))),
-                           levels=c('small','medium','large','hyperexpanded'))
-  tmp[[sample]] <- occupiedscRepertoire(sobj,
-                                        x.axis = "celltype",
-                                        exportTable=TRUE) %>%
-    dplyr::mutate(sample=gsub('_PBMC','',sample))
-}
-
-ggplot(do.call(rbind,tmp),aes(x=celltype,y=value,fill=cloneType)) +
-  geom_col() +
-  theme_classic() +
-  facet_wrap(~sample,scales='free_y',ncol=2,strip.position='left') +
-  scale_fill_manual(values=cloneType_colors) +
-  theme(axis.text.x=element_text(angle=90,hjust=1,vjust=.5),
-        strip.background=element_blank(),
-        strip.placement='outside',
-        legend.key.size=unit(3,'mm')) +
-  labs(x='',y='')
-```
-
-```{r Fig_S10D, fig.width=4.5,fig.height=2.}
+<!-- ```{r Fig_S10F,fig.width=4.5,fig.height=3} -->
+<!-- ggplot(df %>%  -->
+<!--          dplyr::filter(Diagnosis=='UC') %>% -->
+<!--          gather(metric,value,-c(sample,Patient.ID,Responder,Diagnosis,Timepoint,Batch,celltype,Material)) %>% -->
+<!--          dplyr::filter(metric=='Inv.Simpson',Responder!='control') %>% -->
+<!--          dplyr::mutate(Responder=ifelse(Responder=='yes','R','NR')),  -->
+<!--        aes(x=Responder,y=value)) + -->
+<!--   geom_boxplot(aes(fill=Responder,color=Responder),outlier.shape=NA) + -->
+<!--   geom_point(size=1,color='gray30') + -->
+<!--   theme_classic() + -->
+<!--   coord_cartesian(clip='off') + -->
+<!--   scale_y_log10() + -->
+<!--   facet_grid(Timepoint~celltype,scales='free_y',switch='y') + -->
+<!--   stat_compare_means(comparisons=list(c('NR','R')),label = "p.format",size=2,bracket.size=.25,tip.length=.01) + -->
+<!--   scale_fill_manual(values=responder_colors) + -->
+<!--   scale_color_manual(values=responder_colors) + -->
+<!--   labs(x='',y='Inv. Simpson',title='') + -->
+<!--   theme(legend.key.size=unit(4,'mm'), -->
+<!--         legend.box='horizontal', -->
+<!--         axis.text.x=element_text(angle=90,hjust=1,vjust=0), -->
+<!--         strip.background=element_blank(), -->
+<!--         strip.placement='outside') -->
+<!-- ``` -->
+
+<!-- ```{r clonotype_celltypes_BCR,fig.width=6.5,fig.height=10} -->
+<!-- tmp <- list() -->
+<!-- for (sample in grep('PBMC',names(pbmc.B),value=TRUE)) { -->
+<!--   sobj <- pbmc.B[[sample]] -->
+<!--   sobj$celltype <- plyr::revalue(sobj$predicted.celltype.l2,ct_map) -->
+<!--   sobj$cloneType <- factor(tolower(gsub('Single','Small',gsub(' .*','',sobj$cloneType))), -->
+<!--                            levels=c('small','medium','large','hyperexpanded')) -->
+<!--   tmp[[sample]] <- occupiedscRepertoire(sobj, -->
+<!--                                         x.axis = "celltype", -->
+<!--                                         exportTable=TRUE) %>% -->
+<!--     dplyr::mutate(sample=gsub('_PBMC','',sample)) -->
+<!-- } -->
+
+<!-- ggplot(do.call(rbind,tmp),aes(x=celltype,y=value,fill=cloneType)) + -->
+<!--   geom_col() + -->
+<!--   theme_classic() + -->
+<!--   facet_wrap(~sample,scales='free_y',ncol=2,strip.position='left') + -->
+<!--   scale_fill_manual(values=cloneType_colors) + -->
+<!--   theme(axis.text.x=element_text(angle=90,hjust=1,vjust=.5), -->
+<!--         strip.background=element_blank(), -->
+<!--         strip.placement='outside', -->
+<!--         legend.key.size=unit(3,'mm')) + -->
+<!--   labs(x='',y='') -->
+<!-- ``` -->
+
+```{r Fig_S5H, fig.width=4.5,fig.height=2.}
 pbmc_frac %>%
@@
 
-# Fig. 4 supplements
-
-```{r get_GSE73661}
-olink_genes <- c('CXCL8','VEGFA','CD8A','CCL7','GDNF','CDCP1','CD244','IL7','TNFRSF11B','TGFB1','PLAU','IL6','IL17C','CCL2','IL17A','CXCL11',
-                 'AXIN1','TNFSF10','CXCL9','CST5','OSM','CXCL1','CCL4','CD6','SCF','IL18','SLAMF1','TGFA','CCL13','CCL11','TNFSF14','IL10RA',
-                 'FGF5','MMP1','LIFR','FGF21','CCL19','IL15RA','IL10RB','IL18R1','CD274','CXCL5','TNFSF11','HGF','IL12B','MMP10','IL10','TNF',
-                 'CCL23','CD5','CCL3','FLT3LG','CXCL6','CXCL10','EIF4EBP1','SIRT2','CCL28','DNER','S100A12','CD40','IFNG','FGF19','CCL8','CASP8',
-                 'CCL25','CX3CL1','TNFRSF9','NTF3','TNFSF12','CCL20','SULT1A1','STAMBP','ADA','LTA','CSF1')
-
-meta <- read.table(file.path(path,'GSE73661','GSE73661_Vedo.txt'),header=1,sep='\t') %>%
-  dplyr::mutate(sample=make.names(geo_accession)) %>%
-  tibble::rownames_to_column("accession")  %>%
-  dplyr::mutate(diagnosis=factor(ifelse(grepl('UC',sample),'UC','control'),levels=c('control','UC')),
-                treatment=factor(ifelse(diagnosis=='control','control',
-                                        ifelse(grepl('plac',sample),'placebo',
-                                               ifelse(grepl('IFX',sample),'IFX',
-                                                      ifelse(grepl('vdz4w',sample),'vdz4w','vdz8w')))),
-                                 levels=c('control','placebo','IFX','vdz4w','vdz8w')),
-                timepoint=factor(ifelse(diagnosis=='control','control',
-                                        ifelse(grepl('W0',sample),'w0',
-                                               ifelse(grepl('W4|W6',sample),'w6',
-                                                      ifelse(grepl('W12',sample),'w12',
-                                                             ifelse(grepl('W52',sample),'w52','other'))))),
-                                 levels=c('control','other','w0','w6','w12','w52')),
-                responder=ifelse(diagnosis=='control','control',ifelse(Response %in% c(0,'Placebo'),'other',Response))) %>%
-  dplyr::filter(!grepl('earlier.stop',sample),
-                !grepl('_47$',sample),
-                !grepl('_73$',sample),
-                (diagnosis=='control') | ((timepoint %in% c('w0','w6')) & (treatment=='vdz4w') & (responder!='other'))) %>%
-  droplevels()
-
-row.names(meta) <- meta$sample
-colnames(meta) <- gsub('.ch1','',colnames(meta))
-data <- read.table(file.path(path,'GSE73661','GSE73661.txt'),header=1,sep='\t')
-data <- data[,colnames(data) %in% row.names(meta)]
-
-GSE73661 <- SummarizedExperiment(assays=list(data),colData=meta[colnames(data),])
-```
-
-show a heatmap of gene expression
-
-```{r Fig_S17C,fig.width=10,fig.height=10}
-olink_genes_here <- setdiff(olink_genes, c('SCF','TNFSF12'))
-
-ha <- HeatmapAnnotation(df=colData(GSE73661)[,c('timepoint','responder','diagnosis')],
-                        show_legend=TRUE,
-                        col=list('timepoint'=c('control'='#000000',
-                                               'w0'='#ff7700',
-                                               'w6'='#882e72'),
-                                 'responder'=c('control'='#000000',
-                                               'R'='#db2F26',
-                                               'NR'='#34a048'),
-                                 'diagnosis'=c('UC'='#1965b0',
-                                               'control'='#000000')))
-
-mat <- t(scale(t(assay(GSE73661)[olink_genes_here,])))
-Heatmap(mat,
-        cluster_columns=TRUE,
-        show_row_names=TRUE,
-        show_column_names=FALSE,
-        top_annotation=ha,
-        column_names_gp=gpar(fontsize=6),
-        row_names_gp=gpar(fontsize=6),
-        name='z-score') 
-```
-
-run a few contrasts:
-
-- w6vsw0 @ UC
-- UCvscontrol
-
-```{r GSE73661_DE,fig.width=6,fig.height=5}
-library(limma)
-res <- list()
-
-take <- colData(GSE73661)$diagnosis=='UC'
-design <- model.matrix(~0+timepoint,data=colData(GSE73661)[take,])
-fit <- eBayes(lmFit(assay(GSE73661)[,take], design))
-contrast.matrix <- makeContrasts(w6vsw0=timepointw6-timepointw0,levels=design)
-contrast_fit <- eBayes(contrasts.fit(fit, contrast.matrix))
-res[['w6vsw0@UC']] <- topTable(contrast_fit, coef='w6vsw0', n=Inf)  %>%
-  tibble::rownames_to_column('gene') %>%
-  dplyr::filter(gene %in% olink_genes)  %>%
-  dplyr::mutate(contrast='w6vsw0@UC')
-
-design <- model.matrix(~0+diagnosis,data=colData(GSE73661))
-fit <- eBayes(lmFit(assay(GSE73661), design))
-contrast.matrix <- makeContrasts(UCvscontrol=diagnosisUC-diagnosiscontrol,levels=design)
-contrast_fit <- eBayes(contrasts.fit(fit, contrast.matrix))
-res[['UCvscontrol']] <- topTable(contrast_fit, coef='UCvscontrol', n=Inf)  %>%
-  tibble::rownames_to_column('gene') %>%
-  dplyr::filter(gene %in% olink_genes)  %>%
-  dplyr::mutate(contrast='UCvscontrol')
-
-GSE73661_DE <- do.call(rbind,res) %>%
-    dplyr::select(gene,AveExpr,logFC,P.Value,adj.P.Val,contrast)
-```
-
-show volcano plots (only for Olink genes!)
-
-```{r Fig_S17A_B,fig.width=4.5,fig.height=4}
-for (ct in c('w6vsw0@UC','UCvscontrol')) {
-  res <- GSE73661_DE %>% dplyr::filter(contrast==ct)
-  pl <-  res %>%
-    ggplot(aes(x=logFC,y=-log10(P.Value))) + 
-    geom_point(aes(color=(P.Value < .05)),size=.5) + 
-    geom_text_repel(data=res %>% dplyr::filter(P.Value < .05),
-                    aes(label=gene),segment.size=.25,segment.alpha=.25,color='red') +
-    theme_classic() +
-    geom_hline(yintercept=-log10(.05),linetype='dashed',size=.25) +
-    geom_vline(xintercept=0,linetype='dashed',size=.25) +
-    scale_color_manual(values=c('black','red'),guide=FALSE) +
-    labs(title=ct)
-  print(pl)
-}
-```
-
-# Fig. 6 + supplements
-
-```{r Fig_6B,fig.width=4.5,fig.height=3.5}
+<!-- # Fig. 4 supplements -->
+
+<!-- ```{r get_GSE73661} -->
+<!-- olink_genes <- c('CXCL8','VEGFA','CD8A','CCL7','GDNF','CDCP1','CD244','IL7','TNFRSF11B','TGFB1','PLAU','IL6','IL17C','CCL2','IL17A','CXCL11', -->
+<!--                  'AXIN1','TNFSF10','CXCL9','CST5','OSM','CXCL1','CCL4','CD6','SCF','IL18','SLAMF1','TGFA','CCL13','CCL11','TNFSF14','IL10RA', -->
+<!--                  'FGF5','MMP1','LIFR','FGF21','CCL19','IL15RA','IL10RB','IL18R1','CD274','CXCL5','TNFSF11','HGF','IL12B','MMP10','IL10','TNF', -->
+<!--                  'CCL23','CD5','CCL3','FLT3LG','CXCL6','CXCL10','EIF4EBP1','SIRT2','CCL28','DNER','S100A12','CD40','IFNG','FGF19','CCL8','CASP8', -->
+<!--                  'CCL25','CX3CL1','TNFRSF9','NTF3','TNFSF12','CCL20','SULT1A1','STAMBP','ADA','LTA','CSF1') -->
+
+<!-- meta <- read.table(file.path(path,'GSE73661','GSE73661_Vedo.txt'),header=1,sep='\t') %>% -->
+<!--   dplyr::mutate(sample=make.names(geo_accession)) %>% -->
+<!--   tibble::rownames_to_column("accession")  %>% -->
+<!--   dplyr::mutate(diagnosis=factor(ifelse(grepl('UC',sample),'UC','control'),levels=c('control','UC')), -->
+<!--                 treatment=factor(ifelse(diagnosis=='control','control', -->
+<!--                                         ifelse(grepl('plac',sample),'placebo', -->
+<!--                                                ifelse(grepl('IFX',sample),'IFX', -->
+<!--                                                       ifelse(grepl('vdz4w',sample),'vdz4w','vdz8w')))), -->
+<!--                                  levels=c('control','placebo','IFX','vdz4w','vdz8w')), -->
+<!--                 timepoint=factor(ifelse(diagnosis=='control','control', -->
+<!--                                         ifelse(grepl('W0',sample),'w0', -->
+<!--                                                ifelse(grepl('W4|W6',sample),'w6', -->
+<!--                                                       ifelse(grepl('W12',sample),'w12', -->
+<!--                                                              ifelse(grepl('W52',sample),'w52','other'))))), -->
+<!--                                  levels=c('control','other','w0','w6','w12','w52')), -->
+<!--                 responder=ifelse(diagnosis=='control','control',ifelse(Response %in% c(0,'Placebo'),'other',Response))) %>% -->
+<!--   dplyr::filter(!grepl('earlier.stop',sample), -->
+<!--                 !grepl('_47$',sample), -->
+<!--                 !grepl('_73$',sample), -->
+<!--                 (diagnosis=='control') | ((timepoint %in% c('w0','w6')) & (treatment=='vdz4w') & (responder!='other'))) %>% -->
+<!--   droplevels() -->
+
+<!-- row.names(meta) <- meta$sample -->
+<!-- colnames(meta) <- gsub('.ch1','',colnames(meta)) -->
+<!-- data <- read.table(file.path(path,'GSE73661','GSE73661.txt'),header=1,sep='\t') -->
+<!-- data <- data[,colnames(data) %in% row.names(meta)] -->
+
+<!-- GSE73661 <- SummarizedExperiment(assays=list(data),colData=meta[colnames(data),]) -->
+<!-- ``` -->
+
+<!-- show a heatmap of gene expression -->
+
+<!-- ```{r Fig_S17C,fig.width=10,fig.height=10} -->
+<!-- olink_genes_here <- setdiff(olink_genes, c('SCF','TNFSF12')) -->
+
+<!-- ha <- HeatmapAnnotation(df=colData(GSE73661)[,c('timepoint','responder','diagnosis')], -->
+<!--                         show_legend=TRUE, -->
+<!--                         col=list('timepoint'=c('control'='#000000', -->
+<!--                                                'w0'='#ff7700', -->
+<!--                                                'w6'='#882e72'), -->
+<!--                                  'responder'=c('control'='#000000', -->
+<!--                                                'R'='#db2F26', -->
+<!--                                                'NR'='#34a048'), -->
+<!--                                  'diagnosis'=c('UC'='#1965b0', -->
+<!--                                                'control'='#000000'))) -->
+
+<!-- mat <- t(scale(t(assay(GSE73661)[olink_genes_here,]))) -->
+<!-- Heatmap(mat, -->
+<!--         cluster_columns=TRUE, -->
+<!--         show_row_names=TRUE, -->
+<!--         show_column_names=FALSE, -->
+<!--         top_annotation=ha, -->
+<!--         column_names_gp=gpar(fontsize=6), -->
+<!--         row_names_gp=gpar(fontsize=6), -->
+<!--         name='z-score')  -->
+<!-- ``` -->
+
+<!-- run a few contrasts: -->
+
+<!-- - w6vsw0 @ UC -->
+<!-- - UCvscontrol -->
+
+<!-- ```{r GSE73661_DE,fig.width=6,fig.height=5} -->
+<!-- library(limma) -->
+<!-- res <- list() -->
+
+<!-- take <- colData(GSE73661)$diagnosis=='UC' -->
+<!-- design <- model.matrix(~0+timepoint,data=colData(GSE73661)[take,]) -->
+<!-- fit <- eBayes(lmFit(assay(GSE73661)[,take], design)) -->
+<!-- contrast.matrix <- makeContrasts(w6vsw0=timepointw6-timepointw0,levels=design) -->
+<!-- contrast_fit <- eBayes(contrasts.fit(fit, contrast.matrix)) -->
+<!-- res[['w6vsw0@UC']] <- topTable(contrast_fit, coef='w6vsw0', n=Inf)  %>% -->
+<!--   tibble::rownames_to_column('gene') %>% -->
+<!--   dplyr::filter(gene %in% olink_genes)  %>% -->
+<!--   dplyr::mutate(contrast='w6vsw0@UC') -->
+
+<!-- design <- model.matrix(~0+diagnosis,data=colData(GSE73661)) -->
+<!-- fit <- eBayes(lmFit(assay(GSE73661), design)) -->
+<!-- contrast.matrix <- makeContrasts(UCvscontrol=diagnosisUC-diagnosiscontrol,levels=design) -->
+<!-- contrast_fit <- eBayes(contrasts.fit(fit, contrast.matrix)) -->
+<!-- res[['UCvscontrol']] <- topTable(contrast_fit, coef='UCvscontrol', n=Inf)  %>% -->
+<!--   tibble::rownames_to_column('gene') %>% -->
+<!--   dplyr::filter(gene %in% olink_genes)  %>% -->
+<!--   dplyr::mutate(contrast='UCvscontrol') -->
+
+<!-- GSE73661_DE <- do.call(rbind,res) %>% -->
+<!--     dplyr::select(gene,AveExpr,logFC,P.Value,adj.P.Val,contrast) -->
+<!-- ``` -->
+
+<!-- show volcano plots (only for Olink genes!) -->
+
+<!-- ```{r Fig_S17A_B,fig.width=4.5,fig.height=4} -->
+<!-- for (ct in c('w6vsw0@UC','UCvscontrol')) { -->
+<!--   res <- GSE73661_DE %>% dplyr::filter(contrast==ct) -->
+<!--   pl <-  res %>% -->
+<!--     ggplot(aes(x=logFC,y=-log10(P.Value))) +  -->
+<!--     geom_point(aes(color=(P.Value < .05)),size=.5) +  -->
+<!--     geom_text_repel(data=res %>% dplyr::filter(P.Value < .05), -->
+<!--                     aes(label=gene),segment.size=.25,segment.alpha=.25,color='red') + -->
+<!--     theme_classic() + -->
+<!--     geom_hline(yintercept=-log10(.05),linetype='dashed',size=.25) + -->
+<!--     geom_vline(xintercept=0,linetype='dashed',size=.25) + -->
+<!--     scale_color_manual(values=c('black','red'),guide=FALSE) + -->
+<!--     labs(title=ct) -->
+<!--   print(pl) -->
+<!-- } -->
+<!-- ``` -->
+
+```{r Fig_2a,fig.width=4.5,fig.height=3.5}
 DimPlot(CD4.integrated,label=TRUE) +
@@
 
-```{r CD4_material,fig.width=4.5,fig.height=3.5}
-DimPlot(CD4.integrated,group.by='Material') +
-  scale_color_brewer(palette='Set1') +
-  theme_void() +
-  theme(legend.key.size=unit(3,'mm')) +
-  labs(title='')
-```
-
-```{r Fig_S19D,fig.width=6,fig.height=20,cache=FALSE}
-DefaultAssay(CD4.integrated) <- 'RNA'
-if (!file.exists(file.path(path,'seurat','CD4_RNA_markers.csv'))) {
-  markers <- FindAllMarkers(CD4.integrated, only.pos = TRUE, min.pct = 0.25,
-                            thresh.use = 0.25,
-                            print.bar=FALSE, verbose=FALSE)
-  write.csv(markers,file.path(path,'seurat','CD4_RNA_markers.csv'))
-}
-
-markers <- read.csv(file.path(path,'seurat','CD4_RNA_markers.csv'))
-
-top10 <- markers %>% 
-  group_by(cluster) %>% 
-  top_n(10, avg_log2FC)
-
-DotPlot(CD4.integrated, features = rev(unique(top10$gene))) + 
-  theme(axis.text.x=element_text(angle=90,hjust=1,vjust=.5)) +
-  labs(x='',y='cluster') + 
-  coord_flip()
-```
-
-```{r Fig_6F,fig.width=12,fig.height=2,cache=FALSE}
-DefaultAssay(CD4.integrated) <- 'RNA'
-
-CD4.integrated$cluster1 <- factor(ifelse(Idents(CD4.integrated)=='1','1','other'),levels=c('1','other'))
-
-# c('CXCR3','CXCR5','CCR2','CCR4','CCR5','CCR6','CCR7','CCR9',
-#                 'IL1R1','IL1RAP','IL1RL1','IL2RA','IL4R','IL7R','IL12RB2','IL18RAP','IL18R1',
-#                 'IL23R','IFNGR1','IFNGR2','TNFRSF1A','TNFRSF1B','TNFRSF4',
-tcell_genes <- c('GPR15','ITGA4','ITGB1','ITGB7',
-                 'AREG','CD40LG','CD38','CD69','CTLA4','GZMA','GZMB','HLA-DRA','ICOS','LAG3',
-                 'MKI67','PDCD1','SELL','SLAMF6','HAVCR2','TIGIT',
-                 'AHR','BATF','BCL6','CREM','EOMES','FOSB','GATA3','IKZF2','JUN','JUNB','MAF',
-                 'NFATC1','PRDM1','RORA','RORC','TBX21','TCF7','TOX','ZBTB7B',
-                 'CSF2','IFNG','IL2','IL4','IL5','IL9','IL10','IL17A','IL22','TNF')
-
-DotPlot(CD4.integrated, features=tcell_genes,
-        group.by='cluster1') +
-  scale_color_viridis_c() +
-  scale_y_discrete(limits=rev) +
-  theme(axis.text.x=element_text(angle=90,hjust=1,vjust=.5)) +
-  labs(x='',y='')  
-```
-
-```{r Fig_6G,fig.width=6,fig.height=2,cache=FALSE}
-DefaultAssay(CD4.integrated) <- 'orig_ADT'
-DotPlot(CD4.integrated, features = c('CD25','CD127','PD-1','CD38','HLA-DR'),
-        group.by='cluster1') + 
-  scale_color_viridis_c() +
-  scale_y_discrete(limits=rev) +
-  theme(axis.text.x=element_text(angle=90,hjust=1,vjust=.5),
-        legend.key.size=unit(3,'mm'),
-        legend.box='horizontal') +
-  labs(x='',y='')
-```
-
-```{r Fig_6C,fig.width=4,fig.height=3,cache=FALSE}
+```{r Fig_2B,fig.width=4,fig.height=3,cache=FALSE}
 DefaultAssay(CD4.integrated) <- 'RNA'
@@
 
-```{r Fig_6E, fig.width=2,fig.height=2,cache=FALSE}
+<!-- ```{r CD4_material,fig.width=4.5,fig.height=3.5} -->
+<!-- DimPlot(CD4.integrated,group.by='Material') + -->
+<!--   scale_color_brewer(palette='Set1') + -->
+<!--   theme_void() + -->
+<!--   theme(legend.key.size=unit(3,'mm')) + -->
+<!--   labs(title='') -->
+<!-- ``` -->
+
+<!-- ```{r Fig_S19D,fig.width=6,fig.height=20,cache=FALSE} -->
+<!-- DefaultAssay(CD4.integrated) <- 'RNA' -->
+<!-- if (!file.exists(file.path(path,'seurat','CD4_RNA_markers.csv'))) { -->
+<!--   markers <- FindAllMarkers(CD4.integrated, only.pos = TRUE, min.pct = 0.25, -->
+<!--                             thresh.use = 0.25, -->
+<!--                             print.bar=FALSE, verbose=FALSE) -->
+<!--   write.csv(markers,file.path(path,'seurat','CD4_RNA_markers.csv')) -->
+<!-- } -->
+
+<!-- markers <- read.csv(file.path(path,'seurat','CD4_RNA_markers.csv')) -->
+
+<!-- top10 <- markers %>%  -->
+<!--   group_by(cluster) %>%  -->
+<!--   top_n(10, avg_log2FC) -->
+
+<!-- DotPlot(CD4.integrated, features = rev(unique(top10$gene))) +  -->
+<!--   theme(axis.text.x=element_text(angle=90,hjust=1,vjust=.5)) + -->
+<!--   labs(x='',y='cluster') +  -->
+<!--   coord_flip() -->
+<!-- ``` -->
+
+```{r Fig_2D, fig.width=2,fig.height=1,cache=FALSE}
 library(gtools)
@@
   geom_boxplot(outlier.shape=NA) +
-  geom_point(size=1,aes(shape=Material),color='gray30') +
+  #geom_point(size=1,aes(shape=Material),color='gray30') +
+  geom_point(size=1,color='gray30') +
   stat_compare_means(aes(group=Patient.ID),comparisons=list(c('NR','R')),label = "p.format", 
@@
   coord_cartesian(clip='off') +
+  coord_flip() +
+  scale_x_discrete(limits=rev) +
   scale_fill_manual(values=responder_colors,limits=force) +
@@
   scale_shape_manual(values=c(1,4)) +
-  theme(axis.text.x=element_blank(),
-        legend.position='right',
+  theme(legend.position='none',
         legend.key.size=unit(3.5,'mm'),
@@
 
-```{r Fig_S19E, fig.width=5,fig.height=2.5}
-CD4_frac %>%
-    dplyr::filter(Timepoint!='ctrl', celltype %in% seq(1,6)) %>%
-    ggplot(aes(x=Responder,y=frac,fill=Responder,color=Responder)) + 
-    geom_boxplot(outlier.shape=NA) +
-    geom_point(size=1,aes(shape=Material),color='gray30') +
-    stat_compare_means(aes(group=Patient.ID),comparisons=list(c('NR','R')),label = "p.format", 
-                       method='wilcox.test',bracket.size=.25,tip.length=.01,size=2) +
-    theme_classic() + 
-    coord_cartesian(clip='off') +
-    facet_grid(Timepoint~celltype,scales='free_y',switch='y') +
-    scale_fill_manual(values=responder_colors,limits=force) +
-    scale_color_manual(values=responder_colors,limits=force) +
-    scale_shape_manual(values=c(1,4)) +
-    theme(axis.text.x=element_blank(),
-          legend.position='right',
-          legend.key.size=unit(3.5,'mm'),
-          strip.background=element_blank(),
-          strip.placement='outside')  +
-    labs(x='',y='cell fraction',fill='',color='',shape='')
-
-CD4_frac %>%
-    dplyr::filter(Timepoint!='ctrl', celltype %in% seq(7,12)) %>%
-    ggplot(aes(x=Responder,y=frac,fill=Responder,color=Responder)) + 
-    geom_boxplot(outlier.shape=NA) +
-    geom_point(size=1,aes(shape=Material),color='gray30') +
-    stat_compare_means(aes(group=Patient.ID),comparisons=list(c('NR','R')),label = "p.format", 
-                       method='wilcox.test',bracket.size=.25,tip.length=.01,size=2) +
-    theme_classic() + 
-    coord_cartesian(clip='off') +
-    facet_grid(Timepoint~celltype,scales='free_y',switch='y') +
-    scale_fill_manual(values=responder_colors,limits=force) +
-    scale_color_manual(values=responder_colors,limits=force) +
-    scale_shape_manual(values=c(1,4)) +
-    theme(axis.text.x=element_blank(),
-          legend.position='right',
-          legend.key.size=unit(3.5,'mm'),
-          strip.background=element_blank(),
-          strip.placement='outside')  +
-    labs(x='',y='cell fraction',fill='',color='',shape='')
-```
-
-```{r Fig_6D, fig.width=3,fig.height=2,cache=FALSE}
+<!-- ```{r Fig_S19E, fig.width=5,fig.height=2.5} -->
+<!-- CD4_frac %>% -->
+<!--     dplyr::filter(Timepoint!='ctrl', celltype %in% seq(1,6)) %>% -->
+<!--     ggplot(aes(x=Responder,y=frac,fill=Responder,color=Responder)) +  -->
+<!--     geom_boxplot(outlier.shape=NA) + -->
+<!--     geom_point(size=1,aes(shape=Material),color='gray30') + -->
+<!--     stat_compare_means(aes(group=Patient.ID),comparisons=list(c('NR','R')),label = "p.format",  -->
+<!--                        method='wilcox.test',bracket.size=.25,tip.length=.01,size=2) + -->
+<!--     theme_classic() +  -->
+<!--     coord_cartesian(clip='off') + -->
+<!--     facet_grid(Timepoint~celltype,scales='free_y',switch='y') + -->
+<!--     scale_fill_manual(values=responder_colors,limits=force) + -->
+<!--     scale_color_manual(values=responder_colors,limits=force) + -->
+<!--     scale_shape_manual(values=c(1,4)) + -->
+<!--     theme(axis.text.x=element_blank(), -->
+<!--           legend.position='right', -->
+<!--           legend.key.size=unit(3.5,'mm'), -->
+<!--           strip.background=element_blank(), -->
+<!--           strip.placement='outside')  + -->
+<!--     labs(x='',y='cell fraction',fill='',color='',shape='') -->
+
+<!-- CD4_frac %>% -->
+<!--     dplyr::filter(Timepoint!='ctrl', celltype %in% seq(7,12)) %>% -->
+<!--     ggplot(aes(x=Responder,y=frac,fill=Responder,color=Responder)) +  -->
+<!--     geom_boxplot(outlier.shape=NA) + -->
+<!--     geom_point(size=1,aes(shape=Material),color='gray30') + -->
+<!--     stat_compare_means(aes(group=Patient.ID),comparisons=list(c('NR','R')),label = "p.format",  -->
+<!--                        method='wilcox.test',bracket.size=.25,tip.length=.01,size=2) + -->
+<!--     theme_classic() +  -->
+<!--     coord_cartesian(clip='off') + -->
+<!--     facet_grid(Timepoint~celltype,scales='free_y',switch='y') + -->
+<!--     scale_fill_manual(values=responder_colors,limits=force) + -->
+<!--     scale_color_manual(values=responder_colors,limits=force) + -->
+<!--     scale_shape_manual(values=c(1,4)) + -->
+<!--     theme(axis.text.x=element_blank(), -->
+<!--           legend.position='right', -->
+<!--           legend.key.size=unit(3.5,'mm'), -->
+<!--           strip.background=element_blank(), -->
+<!--           strip.placement='outside')  + -->
+<!--     labs(x='',y='cell fraction',fill='',color='',shape='') -->
+<!-- ``` -->
+
+```{r Fig_2C, fig.width=3,fig.height=2,cache=FALSE}
 library(gtools)
@@
 
+```{r CD4_regression_DE,cache=FALSE}
+DefaultAssay(CD4.integrated) <- 'RNA'
+CD4.integrated$cluster1 <- factor(ifelse(Idents(CD4.integrated)=='1','1','other'),levels=c('1','other'))
+
+genes <- list('activation'=c('HLA-DRA','CD40LG','CD38','CD69','CTLA4','AREG','ICOS','LAG3','MKI67','PDCD1','SELL','SLAMF6','HAVCR2','TIGIT'),
+              'homing'=c('CXCR3','CXCR5','CCR2','CCR4','CCR5','CCR6','CCR7','CCR9','GPR15','ITGA4','ITGB1','ITGB7'),
+              'cytokines'=c('CSF2','IFNG','IL2','IL4','IL5','IL9','IL10','IL17A','IL22','TNF','GZMA','GZMB'),
+              'receptors'=c('IL1R1','IL1RAP','IL2RA','IL4R','IL7R','IL12RB2','IL18RAP','IL18R1','IL23R','IFNGR1','IFNGR2','TNFRSF1A','TNFRSF1B','TNFRSF14'),
+              'TFs'=c('AHR','BATF','BCL6','CREM','EOMES','FOSB','GATA3','IKZF2','JUN','JUNB','MAF','NFATC1','PRDM1','RORA','RORC','TCF7','TOX','ZBTB7B','TBX21'))
+
+if (!file.exists(file.path(path,'seurat','CD4_cluster1_markers.csv'))) { 
+  CD4_cluster1_markers <- FindMarkers(CD4.integrated,test.use='LR',logfc.threshold=0,
+                                      latent.vars=c('Material'),min.pct=0.0,
+                                      group.by='cluster1',ident.1='1',ident.2='other',
+                                      features=intersect(rownames(CD4.integrated),unname(unlist(genes))))
+  write.csv(CD4_cluster1_markers,file.path(path,'seurat','CD4_cluster1_markers.csv')) 
+} else {
+  CD4_cluster1_markers <- read.csv(file.path(path,'seurat','CD4_cluster1_markers.csv'),row.names=1)
+}
+```
+
+as scatter plot
+
+```{r Fig_2F_alt1,fig.width=6,fig.height=4}
+tmp <- CD4_cluster1_markers %>% 
+  tibble::rownames_to_column('gene') %>%
+  dplyr::inner_join(data.frame(gene=unlist(genes)) %>% tibble::rownames_to_column('group') %>% dplyr::mutate(group=gsub('[0-9]*$','',group)),
+                    by='gene') %>%
+  dplyr::mutate(group=factor(group),
+                gene=factor(gene,levels=unname(unlist(genes))))
+
+ggplot(tmp %>% dplyr::filter(p_val_adj < .01), 
+  aes(x=1.e-4+pct.2,y=1.e-4+pct.1,fill=avg_log2FC)) +
+  geom_abline(slope=1,intercept=0,size=.5,color='gray') +
+  #geom_point(size=1,pch=20,alpha=.5,color='gray') +
+  geom_point(aes(shape=group),alpha=.5,size=2) +
+  geom_text_repel(aes(label=gene), force=5,max.overlaps=25,
+                  size=2.5,segment.alpha=.5,segment.size=.5) + #,nudge_x=.0,nudge_y=.1) +
+  scale_fill_gradient2(low='blue3',high='red3') +
+  labs(x='% positive cells other clusters',
+       y='% positive cells cluster 1',
+       fill='log2 fold change',
+       size='p-value') +
+  theme_classic() +
+  coord_cartesian(xlim=c(-.1,1),ylim=c(-.1,1)) +
+  guides(color=guide_legend(override.aes=list(shape=15,size=3,label="o"))) +
+  scale_shape_manual(values=seq(21,26)) +
+  theme(legend.key.size=unit(3,'mm'))
+```
+
+as lollipop chart
+
+```{r Fig_2F_alt2,fig.width=10,fig.height=3}
+ggplot(tmp %>% dplyr::arrange(group,avg_log2FC), aes(x=gene,y=avg_log2FC,color=avg_log2FC)) +
+  geom_segment(aes(xend=gene,yend=0)) +
+  geom_point(data=tmp,aes(size=-log10(p_val+1.e-30)),pch=19) +
+  scale_color_gradient2(low='blue3',high='red3') +
+  labs(y='avg. log2 fold change',
+       x='',
+       fill='log2 fold change',
+       size='p-value') +
+  facet_grid(.~group,scales='free_x',space='free_x') +
+  theme_classic() +
+  coord_cartesian(clip='off') +
+  scale_size_continuous(limits=c(0,50),breaks=c(10,20,30),labels=c('1.e-10','1.e-20','<1.e-30')) +
+  theme(legend.key.size=unit(3,'mm'),
+        axis.text.x=element_text(angle=90,hjust=1,vjust=.5),
+        strip.background=element_blank(),
+        strip.clip='off')
+```
+
+```{r Fig_S8D,fig.width=6,fig.height=2,cache=FALSE}
+DefaultAssay(CD4.integrated) <- 'orig_ADT'
+DotPlot(CD4.integrated, features = c('CD25','CD127','PD-1','CD38','HLA-DR'),
+        group.by='cluster1', scale=FALSE) + 
+  scale_color_viridis_c() +
+  scale_y_discrete(limits=rev) +
+  theme(axis.text.x=element_text(angle=90,hjust=1,vjust=.5),
+        legend.key.size=unit(3,'mm'),
+        legend.box='horizontal') +
+  labs(x='',y='')
+```
+
 ```{r sessionInfo}

### 2024-03-11 scRNAseq/paper_figures.Rmd
update figures for revision

@@
 ```{r Fig_S3D,fig.width=7,fig.height=3}
-RNA %>% 
+pl <- RNA %>% 
   spread(gene,expression) %>% 
@@
   labs(x='',y='fraction ITGA4/ITGB7\ndouble-positive cells',fill='',color='')
+
+pg <- ggplotGrob(pl)
+for(i in which(grepl("strip-b", pg$layout$name))){
+  pg$grobs[[i]]$layout$clip <- "off"
+}
+
+grid::grid.draw(pg)
 ```
@@
 
-<!-- ```{r Fig_S11F,fig.width=6,fig.height=3} -->
-<!-- pl <- ggplot(df %>%  -->
-<!--                dplyr::filter(Timepoint!='ctrl',celltype!='Tregs') %>% -->
-<!--                gather(metric,value,-c(sample,Patient.ID,Responder,Diagnosis,Timepoint,Batch,celltype)) %>% -->
-<!--                dplyr::filter(metric=='Inv.Simpson') %>% -->
-<!--                dplyr::arrange(Responder,celltype,Timepoint,Patient.ID), -->
-<!--              aes(x=Responder,y=value)) + -->
-<!--   geom_boxplot(aes(fill=Responder,color=Responder),outlier.shape=NA) + -->
-<!--   geom_point(size=1,color='gray30') + -->
-<!--   theme_classic() + -->
-<!--   coord_cartesian(clip='off') + -->
-<!--   scale_y_log10() + -->
-<!--   facet_grid(Timepoint ~ celltype,scales='free_y',switch='y') + -->
-<!--   stat_compare_means(comparisons=list(c('NR','R')),label = "p.format",size=2,bracket.size=.25,tip.length=.01) + -->
-<!--   scale_fill_manual(values=responder_colors,limits=force) + -->
-<!--   scale_color_manual(values=responder_colors,limits=force) + -->
-<!--   labs(x='',y='Inv. Simpson',title='') + -->
-<!--   theme(legend.key.size=unit(4,'mm'), -->
-<!--         legend.box='horizontal', -->
-<!--         axis.text.x=element_blank(), -->
-<!--         strip.background=element_blank(), -->
-<!--         strip.text=element_text(size=8), -->
-<!--         strip.placement='outside') -->
-
-<!-- pg <- ggplotGrob(pl) -->
-<!-- for(i in which(grepl("strip-t", pg$layout$name))){ -->
-<!--   pg$grobs[[i]]$layout$clip <- "off" -->
-<!-- } -->
-
-<!-- grid::grid.draw(pg) -->
-<!-- #pl -->
-<!-- ``` -->
+```{r Fig_S11F,fig.width=6,fig.height=3}
+pl <- ggplot(df %>%
+               dplyr::filter(Timepoint!='ctrl',celltype!='Tregs') %>%
+               gather(metric,value,-c(sample,Patient.ID,Responder,Diagnosis,Timepoint,Batch,celltype)) %>%
+               dplyr::filter(metric=='Inv.Simpson') %>%
+               dplyr::arrange(Responder,celltype,Timepoint,Patient.ID),
+             aes(x=Responder,y=value)) +
+  geom_boxplot(aes(fill=Responder,color=Responder),outlier.shape=NA) +
+  geom_point(size=1,color='gray30') +
+  theme_classic() +
+  coord_cartesian(clip='off') +
+  scale_y_log10() +
+  facet_grid(Timepoint ~ celltype,scales='free_y',switch='y') +
+  stat_compare_means(comparisons=list(c('NR','R')),label = "p.format",size=2,bracket.size=.25,tip.length=.01) +
+  scale_fill_manual(values=responder_colors,limits=force) +
+  scale_color_manual(values=responder_colors,limits=force) +
+  labs(x='',y='Inv. Simpson',title='') +
+  theme(legend.key.size=unit(4,'mm'),
+        legend.box='horizontal',
+        axis.text.x=element_blank(),
+        strip.background=element_blank(),
+        strip.text=element_text(size=8),
+        strip.placement='outside')
+
+pg <- ggplotGrob(pl)
+for(i in which(grepl("strip-t", pg$layout$name))){
+  pg$grobs[[i]]$layout$clip <- "off"
+}
+
+grid::grid.draw(pg)
+#pl
+```
 
@@
 
-- ITGA4 or ITGB7 positive
+- ITGA4, ITGB7 or ITGB1 positive
 - Integrin-b7 positive (after determining a threshold)
 - GPR15,  CCR9 or ITGAE positive 
+- ITGA4, ITGB7 or ITGB1 negative
+- Integrin-b7 negative
 
@@
 
-cols <- c('sample','orig.ident','predicted.celltype.l2','ITGA4','ITGB7','Integrin-b7','GPR15','CCR9','ITGAE','MKI67')
+cols <- c('sample','orig.ident','predicted.celltype.l2','ITGA4','ITGB7','ITGB1','Integrin-b7','GPR15','CCR9','ITGAE','MKI67')
 pbmc_md <- do.call(rbind,lapply(pbmc[grepl('PBMC',names(pbmc))], function(x) FetchData(x,cols))) %>%
@@
               ITGAE=pbmc_md$cell[grepl(ct,pbmc_md$celltype) & (pbmc_md$ITGAE > 0)],
-              MKI67=pbmc_md$cell[grepl(ct,pbmc_md$celltype) & (pbmc_md$MKI67 > 0)])
+              MKI67=pbmc_md$cell[grepl(ct,pbmc_md$celltype) & (pbmc_md$MKI67 > 0)],
+              ITGB1=pbmc_md$cell[grepl(ct,pbmc_md$celltype) & (pbmc_md$ITGB1 > 0)],
+              ITGA4neg=pbmc_md$cell[grepl(ct,pbmc_md$celltype) & (pbmc_md$ITGA4==0)],
+              ITGB7neg=pbmc_md$cell[grepl(ct,pbmc_md$celltype) & (pbmc_md$ITGB7==0)],
+              ITGB1neg=pbmc_md$cell[grepl(ct,pbmc_md$celltype) & (pbmc_md$ITGB1==0)],
+              Integrin_b7_neg=pbmc_md$cell[grepl(ct,pbmc_md$celltype) & (pbmc_md$Integrinb7=='neg')])
   for (nn in names(bcs)) {
@@
 
-for (ss in c('GPR15','Integrin_b7','ITGA4','ITGAE','ITGB7')) {
+for (ss in c('GPR15','Integrin_b7','ITGA4','ITGAE','ITGB7','ITGB1','ITGA4neg','ITGB7neg','ITGB1neg','Integrin_b7_neg')) {
   for (ct in c('Mmry CD4','Mmry CD8')) {
@@
 
-<!-- ```{r Fig_S10F,fig.width=4.5,fig.height=3} -->
-<!-- ggplot(df %>%  -->
-<!--          dplyr::filter(Diagnosis=='UC') %>% -->
-<!--          gather(metric,value,-c(sample,Patient.ID,Responder,Diagnosis,Timepoint,Batch,celltype,Material)) %>% -->
-<!--          dplyr::filter(metric=='Inv.Simpson',Responder!='control') %>% -->
-<!--          dplyr::mutate(Responder=ifelse(Responder=='yes','R','NR')),  -->
-<!--        aes(x=Responder,y=value)) + -->
-<!--   geom_boxplot(aes(fill=Responder,color=Responder),outlier.shape=NA) + -->
-<!--   geom_point(size=1,color='gray30') + -->
-<!--   theme_classic() + -->
-<!--   coord_cartesian(clip='off') + -->
-<!--   scale_y_log10() + -->
-<!--   facet_grid(Timepoint~celltype,scales='free_y',switch='y') + -->
-<!--   stat_compare_means(comparisons=list(c('NR','R')),label = "p.format",size=2,bracket.size=.25,tip.length=.01) + -->
-<!--   scale_fill_manual(values=responder_colors) + -->
-<!--   scale_color_manual(values=responder_colors) + -->
-<!--   labs(x='',y='Inv. Simpson',title='') + -->
-<!--   theme(legend.key.size=unit(4,'mm'), -->
-<!--         legend.box='horizontal', -->
-<!--         axis.text.x=element_text(angle=90,hjust=1,vjust=0), -->
-<!--         strip.background=element_blank(), -->
-<!--         strip.placement='outside') -->
-<!-- ``` -->
+```{r Fig_S10F,fig.width=4.5,fig.height=3}
+ggplot(df %>%
+         dplyr::filter(Diagnosis=='UC') %>%
+         gather(metric,value,-c(sample,Patient.ID,Responder,Diagnosis,Timepoint,Batch,celltype,Material)) %>%
+         dplyr::filter(metric=='Inv.Simpson',Responder!='control') %>%
+         dplyr::mutate(Responder=ifelse(Responder=='yes','R','NR')),
+       aes(x=Responder,y=value)) +
+  geom_boxplot(aes(fill=Responder,color=Responder),outlier.shape=NA) +
+  geom_point(size=1,color='gray30') +
+  theme_classic() +
+  coord_cartesian(clip='off') +
+  scale_y_log10() +
+  facet_grid(Timepoint~celltype,scales='free_y',switch='y') +
+  stat_compare_means(comparisons=list(c('NR','R')),label = "p.format",size=2,bracket.size=.25,tip.length=.01) +
+  scale_fill_manual(values=responder_colors) +
+  scale_color_manual(values=responder_colors) +
+  labs(x='',y='Inv. Simpson',title='') +
+  theme(legend.key.size=unit(4,'mm'),
+        legend.box='horizontal',
+        axis.text.x=element_text(angle=90,hjust=1,vjust=0),
+        strip.background=element_blank(),
+        strip.placement='outside')
+```
 
@@
 
-<!-- ```{r get_GSE73661} -->
-<!-- olink_genes <- c('CXCL8','VEGFA','CD8A','CCL7','GDNF','CDCP1','CD244','IL7','TNFRSF11B','TGFB1','PLAU','IL6','IL17C','CCL2','IL17A','CXCL11', -->
-<!--                  'AXIN1','TNFSF10','CXCL9','CST5','OSM','CXCL1','CCL4','CD6','SCF','IL18','SLAMF1','TGFA','CCL13','CCL11','TNFSF14','IL10RA', -->
-<!--                  'FGF5','MMP1','LIFR','FGF21','CCL19','IL15RA','IL10RB','IL18R1','CD274','CXCL5','TNFSF11','HGF','IL12B','MMP10','IL10','TNF', -->
-<!--                  'CCL23','CD5','CCL3','FLT3LG','CXCL6','CXCL10','EIF4EBP1','SIRT2','CCL28','DNER','S100A12','CD40','IFNG','FGF19','CCL8','CASP8', -->
-<!--                  'CCL25','CX3CL1','TNFRSF9','NTF3','TNFSF12','CCL20','SULT1A1','STAMBP','ADA','LTA','CSF1') -->
-
-<!-- meta <- read.table(file.path(path,'GSE73661','GSE73661_Vedo.txt'),header=1,sep='\t') %>% -->
-<!--   dplyr::mutate(sample=make.names(geo_accession)) %>% -->
-<!--   tibble::rownames_to_column("accession")  %>% -->
-<!--   dplyr::mutate(diagnosis=factor(ifelse(grepl('UC',sample),'UC','control'),levels=c('control','UC')), -->
-<!--                 treatment=factor(ifelse(diagnosis=='control','control', -->
-<!--                                         ifelse(grepl('plac',sample),'placebo', -->
-<!--                                                ifelse(grepl('IFX',sample),'IFX', -->
-<!--                                                       ifelse(grepl('vdz4w',sample),'vdz4w','vdz8w')))), -->
-<!--                                  levels=c('control','placebo','IFX','vdz4w','vdz8w')), -->
-<!--                 timepoint=factor(ifelse(diagnosis=='control','control', -->
-<!--                                         ifelse(grepl('W0',sample),'w0', -->
-<!--                                                ifelse(grepl('W4|W6',sample),'w6', -->
-<!--                                                       ifelse(grepl('W12',sample),'w12', -->
-<!--                                                              ifelse(grepl('W52',sample),'w52','other'))))), -->
-<!--                                  levels=c('control','other','w0','w6','w12','w52')), -->
-<!--                 responder=ifelse(diagnosis=='control','control',ifelse(Response %in% c(0,'Placebo'),'other',Response))) %>% -->
-<!--   dplyr::filter(!grepl('earlier.stop',sample), -->
-<!--                 !grepl('_47$',sample), -->
-<!--                 !grepl('_73$',sample), -->
-<!--                 (diagnosis=='control') | ((timepoint %in% c('w0','w6')) & (treatment=='vdz4w') & (responder!='other'))) %>% -->
-<!--   droplevels() -->
-
-<!-- row.names(meta) <- meta$sample -->
-<!-- colnames(meta) <- gsub('.ch1','',colnames(meta)) -->
-<!-- data <- read.table(file.path(path,'GSE73661','GSE73661.txt'),header=1,sep='\t') -->
-<!-- data <- data[,colnames(data) %in% row.names(meta)] -->
-
-<!-- GSE73661 <- SummarizedExperiment(assays=list(data),colData=meta[colnames(data),]) -->
-<!-- ``` -->
-
-<!-- show a heatmap of gene expression -->
-
-<!-- ```{r Fig_S17C,fig.width=10,fig.height=10} -->
-<!-- olink_genes_here <- setdiff(olink_genes, c('SCF','TNFSF12')) -->
-
-<!-- ha <- HeatmapAnnotation(df=colData(GSE73661)[,c('timepoint','responder','diagnosis')], -->
-<!--                         show_legend=TRUE, -->
-<!--                         col=list('timepoint'=c('control'='#000000', -->
-<!--                                                'w0'='#ff7700', -->
-<!--                                                'w6'='#882e72'), -->
-<!--                                  'responder'=c('control'='#000000', -->
-<!--                                                'R'='#db2F26', -->
-<!--                                                'NR'='#34a048'), -->
-<!--                                  'diagnosis'=c('UC'='#1965b0', -->
-<!--                                                'control'='#000000'))) -->
-
-<!-- mat <- t(scale(t(assay(GSE73661)[olink_genes_here,]))) -->
-<!-- Heatmap(mat, -->
-<!--         cluster_columns=TRUE, -->
-<!--         show_row_names=TRUE, -->
-<!--         show_column_names=FALSE, -->
-<!--         top_annotation=ha, -->
-<!--         column_names_gp=gpar(fontsize=6), -->
-<!--         row_names_gp=gpar(fontsize=6), -->
-<!--         name='z-score')  -->
-<!-- ``` -->
-
-<!-- run a few contrasts: -->
-
-<!-- - w6vsw0 @ UC -->
-<!-- - UCvscontrol -->
-
-<!-- ```{r GSE73661_DE,fig.width=6,fig.height=5} -->
-<!-- library(limma) -->
-<!-- res <- list() -->
-
-<!-- take <- colData(GSE73661)$diagnosis=='UC' -->
-<!-- design <- model.matrix(~0+timepoint,data=colData(GSE73661)[take,]) -->
-<!-- fit <- eBayes(lmFit(assay(GSE73661)[,take], design)) -->
-<!-- contrast.matrix <- makeContrasts(w6vsw0=timepointw6-timepointw0,levels=design) -->
-<!-- contrast_fit <- eBayes(contrasts.fit(fit, contrast.matrix)) -->
-<!-- res[['w6vsw0@UC']] <- topTable(contrast_fit, coef='w6vsw0', n=Inf)  %>% -->
-<!--   tibble::rownames_to_column('gene') %>% -->
-<!--   dplyr::filter(gene %in% olink_genes)  %>% -->
-<!--   dplyr::mutate(contrast='w6vsw0@UC') -->
-
-<!-- design <- model.matrix(~0+diagnosis,data=colData(GSE73661)) -->
-<!-- fit <- eBayes(lmFit(assay(GSE73661), design)) -->
-<!-- contrast.matrix <- makeContrasts(UCvscontrol=diagnosisUC-diagnosiscontrol,levels=design) -->
-<!-- contrast_fit <- eBayes(contrasts.fit(fit, contrast.matrix)) -->
-<!-- res[['UCvscontrol']] <- topTable(contrast_fit, coef='UCvscontrol', n=Inf)  %>% -->
-<!--   tibble::rownames_to_column('gene') %>% -->
-<!--   dplyr::filter(gene %in% olink_genes)  %>% -->
-<!--   dplyr::mutate(contrast='UCvscontrol') -->
-
-<!-- GSE73661_DE <- do.call(rbind,res) %>% -->
-<!--     dplyr::select(gene,AveExpr,logFC,P.Value,adj.P.Val,contrast) -->
-<!-- ``` -->
-
-<!-- show volcano plots (only for Olink genes!) -->
-
-<!-- ```{r Fig_S17A_B,fig.width=4.5,fig.height=4} -->
-<!-- for (ct in c('w6vsw0@UC','UCvscontrol')) { -->
-<!--   res <- GSE73661_DE %>% dplyr::filter(contrast==ct) -->
-<!--   pl <-  res %>% -->
-<!--     ggplot(aes(x=logFC,y=-log10(P.Value))) +  -->
-<!--     geom_point(aes(color=(P.Value < .05)),size=.5) +  -->
-<!--     geom_text_repel(data=res %>% dplyr::filter(P.Value < .05), -->
-<!--                     aes(label=gene),segment.size=.25,segment.alpha=.25,color='red') + -->
-<!--     theme_classic() + -->
-<!--     geom_hline(yintercept=-log10(.05),linetype='dashed',size=.25) + -->
-<!--     geom_vline(xintercept=0,linetype='dashed',size=.25) + -->
-<!--     scale_color_manual(values=c('black','red'),guide=FALSE) + -->
-<!--     labs(title=ct) -->
-<!--   print(pl) -->
-<!-- } -->
-<!-- ``` -->
-
-```{r Fig_2a,fig.width=4.5,fig.height=3.5}
+```{r get_GSE73661}
+library(limma)
+library(SummarizedExperiment)
+olink_genes <- c('CXCL8','VEGFA','CD8A','CCL7','GDNF','CDCP1','CD244','IL7','TNFRSF11B','TGFB1','PLAU','IL6','IL17C','CCL2','IL17A','CXCL11',
+                 'AXIN1','TNFSF10','CXCL9','CST5','OSM','CXCL1','CCL4','CD6','SCF','IL18','SLAMF1','TGFA','CCL13','CCL11','TNFSF14','IL10RA',
+                 'FGF5','MMP1','LIFR','FGF21','CCL19','IL15RA','IL10RB','IL18R1','CD274','CXCL5','TNFSF11','HGF','IL12B','MMP10','IL10','TNF',
+                 'CCL23','CD5','CCL3','FLT3LG','CXCL6','CXCL10','EIF4EBP1','SIRT2','CCL28','DNER','S100A12','CD40','IFNG','FGF19','CCL8','CASP8',
+                 'CCL25','CX3CL1','TNFRSF9','NTF3','TNFSF12','CCL20','SULT1A1','STAMBP','ADA','LTA','CSF1')
+
+meta <- read.table(file.path(path,'GSE73661','GSE73661_Vedo.txt'),header=1,sep='\t') %>%
+  dplyr::mutate(sample=make.names(geo_accession)) %>%
+  tibble::rownames_to_column("accession")  %>%
+  dplyr::mutate(diagnosis=factor(ifelse(grepl('UC',sample),'UC','control'),levels=c('control','UC')),
+                treatment=factor(ifelse(diagnosis=='control','control',
+                                        ifelse(grepl('plac',sample),'placebo',
+                                               ifelse(grepl('IFX',sample),'IFX',
+                                                      ifelse(grepl('vdz4w',sample),'vdz4w','vdz8w')))),
+                                 levels=c('control','placebo','IFX','vdz4w','vdz8w')),
+                timepoint=factor(ifelse(diagnosis=='control','control',
+                                        ifelse(grepl('W0',sample),'w0',
+                                               ifelse(grepl('W4|W6',sample),'w6',
+                                                      ifelse(grepl('W12',sample),'w12',
+                                                             ifelse(grepl('W52',sample),'w52','other'))))),
+                                 levels=c('control','other','w0','w6','w12','w52')),
+                responder=ifelse(diagnosis=='control','control',ifelse(Response %in% c(0,'Placebo'),'other',Response))) %>%
+  dplyr::filter(!grepl('earlier.stop',sample),
+                !grepl('_47$',sample),
+                !grepl('_73$',sample),
+                (diagnosis=='control') | ((timepoint %in% c('w0','w6')) & (treatment=='vdz4w') & (responder!='other'))) %>%
+  droplevels()
+
+row.names(meta) <- meta$sample
+colnames(meta) <- gsub('.ch1','',colnames(meta))
+data <- read.table(file.path(path,'GSE73661','GSE73661.txt'),header=1,sep='\t')
+data <- data[,colnames(data) %in% row.names(meta)]
+
+GSE73661 <- SummarizedExperiment(assays=list(data),colData=meta[colnames(data),])
+```
+
+show a heatmap of gene expression
+
+```{r Fig_S17C,fig.width=10,fig.height=10}
+olink_genes_here <- setdiff(olink_genes, c('SCF','TNFSF12'))
+
+ha <- HeatmapAnnotation(df=colData(GSE73661)[,c('timepoint','responder','diagnosis')],
+                        show_legend=TRUE,
+                        col=list('timepoint'=c('control'='#000000',
+                                               'w0'='#ff7700',
+                                               'w6'='#882e72'),
+                                 'responder'=c('control'='#000000',
+                                               'R'='#db2F26',
+                                               'NR'='#34a048'),
+                                 'diagnosis'=c('UC'='#1965b0',
+                                               'control'='#000000')))
+
+mat <- t(scale(t(assay(GSE73661)[olink_genes_here,])))
+Heatmap(mat,
+        cluster_columns=TRUE,
+        show_row_names=TRUE,
+        show_column_names=FALSE,
+        top_annotation=ha,
+        column_names_gp=gpar(fontsize=6),
+        row_names_gp=gpar(fontsize=6),
+        name='z-score')
+```
+
+run a few contrasts:
+
+- w6vsw0 @ UC
+- UCvscontrol
+
+```{r GSE73661_DE,fig.width=6,fig.height=5}
+library(limma)
+res <- list()
+
+take <- colData(GSE73661)$diagnosis=='UC'
+design <- model.matrix(~0+timepoint,data=colData(GSE73661)[take,])
+fit <- eBayes(lmFit(assay(GSE73661)[,take], design))
+contrast.matrix <- makeContrasts(w6vsw0=timepointw6-timepointw0,levels=design)
+contrast_fit <- eBayes(contrasts.fit(fit, contrast.matrix))
+res[['w6vsw0@UC']] <- topTable(contrast_fit, coef='w6vsw0', n=Inf)  %>%
+  tibble::rownames_to_column('gene') %>%
+  dplyr::filter(gene %in% olink_genes)  %>%
+  dplyr::mutate(contrast='w6vsw0@UC')
+
+design <- model.matrix(~0+diagnosis,data=colData(GSE73661))
+fit <- eBayes(lmFit(assay(GSE73661), design))
+contrast.matrix <- makeContrasts(UCvscontrol=diagnosisUC-diagnosiscontrol,levels=design)
+contrast_fit <- eBayes(contrasts.fit(fit, contrast.matrix))
+res[['UCvscontrol']] <- topTable(contrast_fit, coef='UCvscontrol', n=Inf)  %>%
+  tibble::rownames_to_column('gene') %>%
+  dplyr::filter(gene %in% olink_genes)  %>%
+  dplyr::mutate(contrast='UCvscontrol')
+
+GSE73661_DE <- do.call(rbind,res) %>%
+    dplyr::select(gene,AveExpr,logFC,P.Value,adj.P.Val,contrast)
+```
+
+show volcano plots (only for Olink genes!)
+
+```{r Fig_S17A_B,fig.width=4.5,fig.height=4}
+for (ct in c('w6vsw0@UC','UCvscontrol')) {
+  res <- GSE73661_DE %>% dplyr::filter(contrast==ct)
+  pl <-  res %>%
+    ggplot(aes(x=logFC,y=-log10(P.Value))) +
+    geom_point(aes(color=(P.Value < .05)),size=.5) +
+    geom_text_repel(data=res %>% dplyr::filter(P.Value < .05),
+                    aes(label=gene),segment.size=.25,segment.alpha=.25,color='red') +
+    theme_classic() +
+    geom_hline(yintercept=-log10(.05),linetype='dashed',size=.25) +
+    geom_vline(xintercept=0,linetype='dashed',size=.25) +
+    scale_color_manual(values=c('black','red'),guide=FALSE) +
+    labs(title=ct)
+  print(pl)
+}
+```
+
+```{r Fig_2A,fig.width=4,fig.height=3}
 DimPlot(CD4.integrated,label=TRUE) +
@@
 
-<!-- ```{r CD4_material,fig.width=4.5,fig.height=3.5} -->
-<!-- DimPlot(CD4.integrated,group.by='Material') + -->
-<!--   scale_color_brewer(palette='Set1') + -->
-<!--   theme_void() + -->
-<!--   theme(legend.key.size=unit(3,'mm')) + -->
-<!--   labs(title='') -->
-<!-- ``` -->
-
-<!-- ```{r Fig_S19D,fig.width=6,fig.height=20,cache=FALSE} -->
-<!-- DefaultAssay(CD4.integrated) <- 'RNA' -->
-<!-- if (!file.exists(file.path(path,'seurat','CD4_RNA_markers.csv'))) { -->
-<!--   markers <- FindAllMarkers(CD4.integrated, only.pos = TRUE, min.pct = 0.25, -->
-<!--                             thresh.use = 0.25, -->
-<!--                             print.bar=FALSE, verbose=FALSE) -->
-<!--   write.csv(markers,file.path(path,'seurat','CD4_RNA_markers.csv')) -->
-<!-- } -->
-
-<!-- markers <- read.csv(file.path(path,'seurat','CD4_RNA_markers.csv')) -->
-
-<!-- top10 <- markers %>%  -->
-<!--   group_by(cluster) %>%  -->
-<!--   top_n(10, avg_log2FC) -->
-
-<!-- DotPlot(CD4.integrated, features = rev(unique(top10$gene))) +  -->
-<!--   theme(axis.text.x=element_text(angle=90,hjust=1,vjust=.5)) + -->
-<!--   labs(x='',y='cluster') +  -->
-<!--   coord_flip() -->
-<!-- ``` -->
+```{r CD4_umap_material,fig.width=4,fig.height=3}
+DimPlot(CD4.integrated,group.by='Material') +
+  scale_color_brewer(palette='Set1') +
+  theme_void() +
+  theme(legend.key.size=unit(3,'mm')) +
+  labs(title='')
+```
+
+```{r CD4_umap_patient,fig.width=4,fig.height=3}
+DimPlot(CD4.integrated,group.by='Patient.ID') +
+  theme_void() +
+  theme(legend.key.size=unit(3,'mm')) +
+  labs(title='')
+```
+
+plot the cluster composition
+
+```{r CD4_cluster_composition,fig.width=4,fig.height=3}
+CD4.integrated[['cluster']] <- Idents(CD4.integrated)
+for (col in c('Diagnosis','Patient.ID','Material')) {
+  CD4.integrated$tmp <- CD4.integrated@meta.data[,col]
+  pl <- rbind(CD4.integrated@meta.data %>% 
+                dplyr::group_by(cluster,tmp) %>%
+                dplyr::summarise(n=n()) %>%
+                dplyr::mutate(frac=n/sum(n),
+                              group='by cluster'),
+              CD4.integrated@meta.data %>% 
+                dplyr::group_by(tmp) %>%
+                dplyr::summarise(n=n()) %>%
+                dplyr::mutate(frac=n/sum(n),
+                              group='all',
+                              cluster='all')) %>%
+    dplyr::mutate(cluster=factor(cluster,levels=c('all',seq(1,12)))) %>%
+    ggplot(aes(x=cluster,y=frac,fill=tmp)) +
+    geom_col(outlier.shape=NA) +
+    theme_classic() +
+    theme(axis.text.x=element_text(angle=90,hjust=1,vjust=.5),
+          legend.key.size=unit(3,'mm'),
+          strip.background=element_blank()) +
+    facet_grid(.~group,scales='free_x',space='free_x') +
+    labs(x='',y='fraction cells',fill='')
+  print(pl)
+}
+```
+
+```{r CD4_markers,fig.width=14,fig.height=4.5,cache=FALSE}
+DefaultAssay(CD4.integrated) <- 'RNA'
+if (!file.exists(file.path(path,'seurat','CD4_RNA_markers.csv'))) {
+  markers <- FindAllMarkers(CD4.integrated, only.pos = TRUE, min.pct = 0.25,
+                            thresh.use = 0.25,
+                            print.bar=FALSE, verbose=FALSE)
+  write.csv(markers,file.path(path,'seurat','CD4_RNA_markers.csv'))
+}
+
+markers <- read.csv(file.path(path,'seurat','CD4_RNA_markers.csv'))
+
+top10 <- markers %>%
+  group_by(cluster) %>%
+  top_n(10, avg_log2FC)
+
+DotPlot(CD4.integrated, features = unique(top10$gene)) +
+  theme_classic() +
+  theme(axis.text.x=element_text(angle=90,hjust=1,vjust=.5),
+        legend.key.size=unit(3,'mm')) +
+  labs(x='',y='cluster')
+```
 
@@
                                       latent.vars=c('Material'),min.pct=0.0,
-                                      group.by='cluster1',ident.1='1',ident.2='other',
-                                      features=intersect(rownames(CD4.integrated),unname(unlist(genes))))
+                                      group.by='cluster1',ident.1='1',ident.2='other')
+#                                      features=intersect(rownames(CD4.integrated),unname(unlist(genes))))
   write.csv(CD4_cluster1_markers,file.path(path,'seurat','CD4_cluster1_markers.csv')) 
@@
 
-```{r Fig_2F_alt1,fig.width=6,fig.height=4}
-tmp <- CD4_cluster1_markers %>% 
+```{r Fig_2E,fig.width=4,fig.height=3.5}
+top_genes <- CD4_cluster1_markers %>% 
+  tibble::rownames_to_column('gene') %>%
+  dplyr::slice_max(abs(avg_log2FC),n=30) %>%
+  dplyr::pull(gene)
+top_genes <- setdiff(c(top_genes,c('ITGA4','ITGB7','CD38','MKI67','CXCR3','CCR7')),c('C12orf75','FLNA','HLA−DRB5','CD99','YWHAQ','HLA−DPB1'))
+ggplot(CD4_cluster1_markers,
+  aes(x=pct.2,y=pct.1)) +
+  geom_abline(slope=1,intercept=0,size=.5,color='gray') +
+  #geom_point(size=1,pch=20,alpha=.5,color='gray') +
+  geom_point(alpha=.5,size=1,color='gray30') +
+  geom_label_repel(data=CD4_cluster1_markers[top_genes,] %>%
+                     tibble::rownames_to_column('gene'),
+                  aes(label=gene), force=5,max.overlaps=25,
+                  label.padding=unit(.5,'mm'),
+                  label.size=0,
+                  color='black',
+                  size=2.5,segment.alpha=.5,segment.size=.5) +
+  scale_color_gradient2(low='blue3',mid='gray',high='red3') +
+  labs(x='% positive cells other clusters',
+       y='% positive cells cluster 1',
+       fill='log2 fold change',
+       size='p-value') +
+  theme_classic() +
+  #coord_cartesian(xlim=c(-.1,1),ylim=c(-.1,1)) +
+  #guides(color=guide_legend(override.aes=list(shape=15,size=3,label="o"))) +
+  scale_shape_manual(values=seq(21,26)) +
+  theme(legend.key.size=unit(3,'mm'))
+```
+
+as lollipop chart
+
+```{r Fig_2F,fig.width=10,fig.height=3}
+CD4_cluster1_markers %>% 
   tibble::rownames_to_column('gene') %>%
@@
   dplyr::mutate(group=factor(group),
-                gene=factor(gene,levels=unname(unlist(genes))))
-
-ggplot(tmp %>% dplyr::filter(p_val_adj < .01), 
-  aes(x=1.e-4+pct.2,y=1.e-4+pct.1,fill=avg_log2FC)) +
-  geom_abline(slope=1,intercept=0,size=.5,color='gray') +
-  #geom_point(size=1,pch=20,alpha=.5,color='gray') +
-  geom_point(aes(shape=group),alpha=.5,size=2) +
-  geom_text_repel(aes(label=gene), force=5,max.overlaps=25,
-                  size=2.5,segment.alpha=.5,segment.size=.5) + #,nudge_x=.0,nudge_y=.1) +
-  scale_fill_gradient2(low='blue3',high='red3') +
-  labs(x='% positive cells other clusters',
-       y='% positive cells cluster 1',
-       fill='log2 fold change',
-       size='p-value') +
-  theme_classic() +
-  coord_cartesian(xlim=c(-.1,1),ylim=c(-.1,1)) +
-  guides(color=guide_legend(override.aes=list(shape=15,size=3,label="o"))) +
-  scale_shape_manual(values=seq(21,26)) +
-  theme(legend.key.size=unit(3,'mm'))
-```
-
-as lollipop chart
-
-```{r Fig_2F_alt2,fig.width=10,fig.height=3}
-ggplot(tmp %>% dplyr::arrange(group,avg_log2FC), aes(x=gene,y=avg_log2FC,color=avg_log2FC)) +
+                gene=factor(gene,levels=unname(unlist(genes)))) %>%
+  dplyr::arrange(group,avg_log2FC) %>%
+  ggplot(aes(x=gene,y=avg_log2FC,color=avg_log2FC)) +
   geom_segment(aes(xend=gene,yend=0)) +
-  geom_point(data=tmp,aes(size=-log10(p_val+1.e-30)),pch=19) +
+  geom_point(aes(size=-log10(p_val+1.e-30)),pch=19) +
   scale_color_gradient2(low='blue3',high='red3') +
@@
 
-```{r Fig_S8D,fig.width=6,fig.height=2,cache=FALSE}
+```{r Fig_S8D,fig.width=4,fig.height=3.5,cache=FALSE}
 DefaultAssay(CD4.integrated) <- 'orig_ADT'
-DotPlot(CD4.integrated, features = c('CD25','CD127','PD-1','CD38','HLA-DR'),
-        group.by='cluster1', scale=FALSE) + 
+DotPlot(CD4.integrated, features = c('PD-1','CD25','CD45RA','CD127','CD38','CD103','CD4','HLA-DR'),
+        group.by='cluster1', scale=TRUE) + 
   scale_color_viridis_c() +
-  scale_y_discrete(limits=rev) +
+  scale_x_discrete(limits=rev) +
+  coord_flip() +
   theme(axis.text.x=element_text(angle=90,hjust=1,vjust=.5),
         legend.key.size=unit(3,'mm'),
-        legend.box='horizontal') +
+        legend.box='vertical') +
   labs(x='',y='')
@@
 
+```{r CD4_clonotypes,fig.width=3,fig.height=2.5,cache=FALSE}
+CD4.integrated <- RenameCells(CD4.integrated, add.cell.id=CD4.integrated$orig.ident)
+
+tmp <- lapply(tcr_combined, function(x) x[x$barcode %in% Cells(CD4.integrated),])
+CD4.integrated <- combineExpression(tmp, CD4.integrated,  cloneCall="aa")
+CD4.integrated$cloneType <- factor(tolower(gsub('Single','Small',gsub(' .*','',CD4.integrated$cloneType))),
+                                   levels=c('small','medium','large','hyperexpanded'))
+
+CD4_md <- CD4.integrated@meta.data %>%
+  tibble::rownames_to_column('cell')
+
+dfs <- list()
+for (cl in unique(CD4.integrated$cluster)) {
+  bcs <- CD4_md$cell[CD4_md$cluster==cl]
+  tcrcomb <- lapply(tcr_combined,
+                    function(x) x[(x$barcode %in% bcs) & !is.na(x$TCR1) & !is.na(x$TCR2),])
+  dfs[[cl]] <- clonalDiversity(tcrcomb,cloneCall='gene',group='sample',exportTable=TRUE,skip.boots=TRUE) %>%
+    dplyr::mutate(cluster=cl)
+}
+df <- do.call(rbind,dfs) %>%
+  dplyr::left_join(study %>% 
+                     dplyr::mutate(Patient.ID=gsub('[P]*([0-9]*)[\\/_][12]','P\\1',Characteristics.Patient.ID.),
+                                   Responder=Characteristics.Responder.,
+                                   Diagnosis=Characteristics.Diagnosis.,
+                                   Batch=gsub(' ','',gsub('_scRNAseq','',Parameter.Value.Method.)),
+                                   Timepoint=ifelse(grepl('_1_',Sample.Name),'Vedo_0d',
+                                                    ifelse(grepl('_2_',Sample.Name),'Vedo_6wk','ctrl'))) %>%
+                     dplyr::select(Patient.ID,Responder,Diagnosis,Sample.Name,Timepoint,Batch),
+                   by=c('sample'='Sample.Name')) %>%
+  dplyr::arrange(Patient.ID)
+
+ggplot(df %>% 
+         #dplyr::filter(Diagnosis=='UC') %>%
+         dplyr::mutate(cluster=factor(cluster,levels=as.character(seq(0,12)))), 
+       aes(x=cluster,y=`Inv.Simpson`)) +
+  geom_boxplot(aes(fill=cluster),outlier.shape=NA) +
+  geom_point(width=.25,size=1,color='gray30') +
+  theme_classic() +
+  coord_cartesian(clip='off') +
+  scale_y_log10() +
+  labs(x='cluster',y='Inv. Simpson',title='') +
+  theme(legend.key.size=unit(4,'mm'),
+        legend.box='horizontal',
+        legend.position='none',
+        axis.text.x=element_text(angle=90,hjust=1,vjust=.5),
+        strip.background=element_blank(),
+        strip.placement='outside')
+```
+
+<!-- ```{r tmp} -->
+<!-- tmp <- CD4_md %>% dplyr::filter(!is.na(CTaa)) %>% dplyr::select(cluster,sample) %>% table() %>% as.data.frame()  -->
+<!-- dplyr::inner_join(df,tmp,by=c('cluster','sample')) %>%  -->
+<!--   ggplot(aes(x=Freq,y=`Inv.Pielou`,label=cluster,color=cluster=='1')) +  -->
+<!--   geom_text() -->
+<!-- ``` -->
 ```{r sessionInfo}

### 2024-03-11 scRNAseq/paper_figures.Rmd
fixed IDs

@@
 
-- K456985_PBMC
-- K783496_PBMC
-- K693104_PBMC
-- K955601_PBMC
-- P1_1_PBMC
-- P1_2_PBMC
-- P3_1_PBMC
-- P3_2_PBMC
-- P8_1_PBMC
-- P8_2_PBMC
-- P22_1_PBMC
-- P22_2_PBMC
-- P23_1_PBMC
-- P23_2_PBMC
-- P25_1_PBMC
-- P25_2_PBMC
-- P27_1_PBMC
-- P27_2_PBMC
-- P28_1_PBMC
-- P28_2_PBMC
-- P22_P23_K783496_CD4
-- P8_P3_K456985_CD4
-- P27_P28_K693104_CD4
-- P12_1_PBMC
-- P12_2_PBMC
-- P26_1_PBMC
-- P26_2_PBMC
-- K496398_PBMC
-- P12_P26_K496398_CD4
+- HC-1_PBMC
+- HC-5_PBMC
+- HC-2_PBMC
+- HC-3_PBMC
+- IBD-1_1_PBMC
+- IBD-1_2_PBMC
+- IBD-9_1_PBMC
+- IBD-9_2_PBMC
+- IBD-10_1_PBMC
+- IBD-10_2_PBMC
+- IBD-3_1_PBMC
+- IBD-3_2_PBMC
+- IBD-4_1_PBMC
+- IBD-4_2_PBMC
+- IBD-5_1_PBMC
+- IBD-5_2_PBMC
+- IBD-7_1_PBMC
+- IBD-7_2_PBMC
+- IBD-8_1_PBMC
+- IBD-8_2_PBMC
+- IBD-3_IBD-4_HC-5_CD4
+- IBD-10_IBD-9_HC-1_CD4
+- IBD-7_IBD-8_HC-2_CD4
+- IBD-2_1_PBMC
+- IBD-2_2_PBMC
+- IBD-6_1_PBMC
+- IBD-6_2_PBMC
+- HC-4_PBMC
+- IBD-2_IBD-6_HC-4_CD4
 
@@
 ```{r get_metadata}
-study <- read.csv(file.path(path,'sample_sheet','s_hegazy_IBD.txt'),
-                 sep='\t') %>%
-  dplyr::mutate(Sample.Name=gsub('p','P',Sample.Name))
-assay <- read.csv(file.path(path,'sample_sheet','a_hegazy_IBD_totalseq_nucleotide_sequencing.txt'),
-                 sep='\t') %>%
+# old IDs
+study <- read.csv(file.path(path,'sample_sheet','s_hegazy_IBD.txt'),sep='\t') %>%
   dplyr::mutate(Sample.Name=gsub('p','P',Sample.Name),
-                Library.Name=gsub('p','P',Library.Name))
+                Patient.ID=gsub('[P]*([0-9]*)[\\/_][12]','P\\1',Characteristics.Patient.ID.),
+                Responder=Characteristics.Responder.,
+                Diagnosis=Characteristics.Diagnosis.,
+                Batch=gsub(' ','',gsub('_scRNAseq','',Parameter.Value.Method.)),
+                Timepoint=ifelse(grepl('_1_',Sample.Name),'Vedo_0d',
+                                 ifelse(grepl('_2_',Sample.Name),'Vedo_6wk','ctrl')),
+                Material=ifelse(Characteristics.Material.=='PBMC','PBMC','CD4'))
+
+# new IDs
+# study <- read.csv(file.path(path,'sample_sheet.txt'),sep='\t') %>%
+#   dplyr::mutate(Patient.ID=Characteristics.Patient.ID.,
+#                 Responder=Characteristics.Responder.,
+#                 Diagnosis=Characteristics.Diagnosis.,
+#                 Batch=gsub(' ','',gsub('_scRNAseq','',Parameter.Value.Method.)),
+#                 Timepoint=ifelse(grepl('_1_',Sample.Name),'Vedo_0d',
+#                                  ifelse(grepl('_2_',Sample.Name),'Vedo_6wk','ctrl')),
+#                 Material=ifelse(Characteristics.Material.=='PBMC','PBMC','CD4'))
 
@@
   dplyr::left_join(study %>%
-                     dplyr::mutate(Patient.ID=gsub('[P]*([0-9]*)[\\/_][12]','P\\1',Characteristics.Patient.ID.),
-                                   Responder=Characteristics.Responder.,
-                                   Diagnosis=Characteristics.Diagnosis.,
-                                   Batch=gsub(' ','',gsub('_scRNAseq','',Parameter.Value.Method.)),
-                                   Timepoint=ifelse(grepl('_1_',Sample.Name),'Vedo_0d',
-                                                    ifelse(grepl('_2_',Sample.Name),'Vedo_6wk','ctrl')),
-                                   Material=ifelse(Characteristics.Material.=='PBMC','PBMC','CD4')) %>%
                      dplyr::select(Patient.ID,Responder,Diagnosis,Sample.Name,Material,Timepoint,Batch),
@@
   dplyr::left_join(study %>% 
-                     dplyr::mutate(Patient.ID=gsub('[P]*([0-9]*)[\\/_][12]','P\\1',Characteristics.Patient.ID.),
-                                   Responder=Characteristics.Responder.,
-                                   Diagnosis=Characteristics.Diagnosis.,
-                                   Batch=gsub(' ','',gsub('_scRNAseq','',Parameter.Value.Method.)),
-                                   Timepoint=ifelse(grepl('_1_',Sample.Name),'Vedo_0d',
-                                                    ifelse(grepl('_2_',Sample.Name),'Vedo_6wk','ctrl')),
-                                   Material=ifelse(Characteristics.Material.=='PBMC','PBMC','CD4')) %>%
                      dplyr::select(Patient.ID,Responder,Diagnosis,Sample.Name,Material,Timepoint,Batch),
@@
   dplyr::left_join(study %>% 
-                     dplyr::mutate(Patient.ID=gsub('[P]*([0-9]*)[\\/_][12]','P\\1',Characteristics.Patient.ID.),
-                                   Responder=ifelse(Characteristics.Responder.=='yes','R','NR'),
-                                   Diagnosis=Characteristics.Diagnosis.,
-                                   Batch=gsub(' ','',gsub('_scRNAseq','',Parameter.Value.Method.)),
-                                   Timepoint=ifelse(grepl('_1_',Sample.Name),'Vedo_0d',
-                                                    ifelse(grepl('_2_',Sample.Name),'Vedo_6wk','ctrl')),
-                                   Material=ifelse(Characteristics.Material.=='PBMC','PBMC','CD4')) %>%
                      dplyr::select(Patient.ID,Responder,Diagnosis,Sample.Name,Material,Timepoint,Batch),
@@
   dplyr::left_join(study %>% 
-                     dplyr::mutate(Patient.ID=gsub('[P]*([0-9]*)[\\/_][12]','P\\1',Characteristics.Patient.ID.),
-                                   Responder=ifelse(Characteristics.Responder.=='yes','R','NR'),
-                                   Diagnosis=Characteristics.Diagnosis.,
-                                   Batch=gsub(' ','',gsub('_scRNAseq','',Parameter.Value.Method.)),
-                                   Timepoint =ifelse(grepl('_1_',Sample.Name),'Vedo_0d',
-                                                    ifelse(grepl('_2_',Sample.Name),'Vedo_6wk','ctrl'))) %>%
                      dplyr::select(Patient.ID,Responder,Diagnosis,Sample.Name,Timepoint,Batch),
@@
   dplyr::left_join(study %>% 
-                     dplyr::mutate(Patient.ID=gsub('[P]*([0-9]*)[\\/_][12]','P\\1',Characteristics.Patient.ID.),
-                                   Responder=Characteristics.Responder.,
-                                   Diagnosis=Characteristics.Diagnosis.,
-                                   Batch=gsub(' ','',gsub('_scRNAseq','',Parameter.Value.Method.)),
-                                   Timepoint=ifelse(grepl('_1_',Sample.Name),'Vedo_0d',
-                                                    ifelse(grepl('_2_',Sample.Name),'Vedo_6wk','ctrl'))) %>%
                      dplyr::select(Patient.ID,Responder,Diagnosis,Sample.Name,Timepoint,Batch),
@@
   dplyr::left_join(study %>% 
-                     dplyr::mutate(Patient.ID=gsub('[P]*([0-9]*)[\\/_][12]','P\\1',Characteristics.Patient.ID.),
-                                   Responder=ifelse(Characteristics.Responder.=='yes','R','NR'),
-                                   Diagnosis=Characteristics.Diagnosis.,
-                                   Batch=gsub(' ','',gsub('_scRNAseq','',Parameter.Value.Method.)),
-                                   Timepoint=ifelse(grepl('_1_',Sample.Name),'Vedo_0d',
-                                                    ifelse(grepl('_2_',Sample.Name),'Vedo_6wk','ctrl'))) %>%
                      dplyr::select(Patient.ID,Responder,Diagnosis,Sample.Name,Timepoint,Batch),

### 2024-03-11 scRNAseq/processing.Rmd
fixed IDs

@@
 
-- K456985_PBMC
-- K783496_PBMC
-- K693104_PBMC
-- K955601_PBMC
-- P1_1_PBMC 
-- P1_2_PBMC
-- P3_1_PBMC
-- P3_2_PBMC
-- P8_1_PBMC
-- P8_2_PBMC 
-- P22_1_PBMC
-- P22_2_PBMC
-- P23_1_PBMC
-- P23_2_PBMC
-- P25_1_PBMC
-- P25_2_PBMC
-- P27_1_PBMC
-- P27_2_PBMC
-- P28_1_PBMC
-- P28_2_PBMC
-- P22_P23_K783496_CD4
-- P8_P3_K456985_CD4
-- P27_P28_K693104_CD4
-- P12_1_PBMC
-- P12_2_PBMC
-- P26_1_PBMC
-- P26_2_PBMC
-- K496398_PBMC
-- P12_P26_K496398_CD4
+- HC-1_PBMC
+- HC-5_PBMC
+- HC-2_PBMC
+- HC-3_PBMC
+- IBD-1_1_PBMC 
+- IBD-1_2_PBMC
+- IBD-9_1_PBMC
+- IBD-9_2_PBMC
+- IBD-10_1_PBMC
+- IBD-10_2_PBMC 
+- IBD-3_1_PBMC
+- IBD-3_2_PBMC
+- IBD-4_1_PBMC
+- IBD-4_2_PBMC
+- IBD-5_1_PBMC
+- IBD-5_2_PBMC
+- IBD-7_1_PBMC
+- IBD-7_2_PBMC
+- IBD-8_1_PBMC
+- IBD-8_2_PBMC
+- IBD-3_IBD-4_HC-5_CD4
+- IBD-10_IBD-9_HC-1_CD4
+- IBD-7_IBD-8_HC-2_CD4
+- IBD-2_1_PBMC
+- IBD-2_2_PBMC
+- IBD-6_1_PBMC
+- IBD-6_2_PBMC
+- HC-4_PBMC
+- IBD-2_IBD-6_HC-4_CD4
 
 ```{r get_data}
-samples <- c('K456985_PBMC','K783496_PBMC','P1_1_PBMC','P22_1_PBMC','P22_P23_K783496_CD4','P23_2_PBMC','P25_2_PBMC','P27_2_PBMC',
-             'P28_1_PBMC','P3_1_PBMC','P8_1_PBMC','P8_P3_K456985_CD4','K693104_PBMC','K955601_PBMC','P1_2_PBMC','P22_2_PBMC',
-             'P23_1_PBMC','P25_1_PBMC','P27_1_PBMC','P27_P28_K693104_CD4','P28_2_PBMC','P3_2_PBMC','P8_2_PBMC',
-             'P12_1_PBMC','P12_2_PBMC','P26_1_PBMC','P26_2_PBMC','K496398_PBMC','P12_P26_K496398_CD4')
+samples <- c('HC-1_PBMC','HC-5_PBMC','IBD-1_1_PBMC','IBD-3_1_PBMC','IBD-3_IBD-4_HC-5_CD4','IBD-4_2_PBMC','IBD-5_2_PBMC','IBD-7_2_PBMC',
+             'IBD-8_1_PBMC','IBD-9_1_PBMC','IBD-10_1_PBMC','IBD-10_IBD-9_HC-1_CD4','HC-2_PBMC','HC-3_PBMC','IBD-1_2_PBMC','IBD-3_2_PBMC',
+             'IBD-4_1_PBMC','IBD-5_1_PBMC','IBD-7_1_PBMC','IBD-7_IBD-8_HC-2_CD4','IBD-8_2_PBMC','IBD-9_2_PBMC','IBD-10_2_PBMC',
+             'IBD-2_1_PBMC','IBD-2_2_PBMC','IBD-6_1_PBMC','IBD-6_2_PBMC','HC-4_PBMC','IBD-2_IBD-6_HC-4_CD4')
 

### 2024-03-11 scRNAseq/processing.Rmd
fixed IDs

@@
 
-study <- read.csv(file.path(path,'sample_sheet','s_hegazy_IBD.txt'),
-                 sep='\t') %>%
-  dplyr::mutate(Sample.Name=gsub('p','P',Sample.Name))
-
-assay <- read.csv(file.path(path,'sample_sheet','a_hegazy_IBD_totalseq_nucleotide_sequencing.txt'),
-                 sep='\t') %>%
-  dplyr::mutate(Sample.Name=gsub('p','P',Sample.Name),
-                Library.Name=gsub('p','P',Library.Name))
+libraries <- read.csv("libraries.txt", sep='\t')
 
@@
   sobj$vireo_donor_id[is.na(sobj$vireo_donor_id)] <- 'none'
-  samples <- assay %>% 
+  samples <- libraries %>% 
     dplyr::filter(Library.Name==lib) %>% 

### 2024-03-11 scRNAseq/recluster_CD4.R

@@ -4,13 +4,26 @@ library(dplyr)
 path <- getwd()
 pbmc <- readRDS(file.path(path,'seurat','PBMC_combined.rds'))
 
-study <- read.csv(file.path(path,'sample_sheet','s_hegazy_IBD.txt'),
-                 sep='\t') %>%
-  dplyr::mutate(Sample.Name=gsub('p','P',Sample.Name))
-assay <- read.csv(file.path(path,'sample_sheet','a_hegazy_IBD_totalseq_nucleotide_sequencing.txt'),
-                 sep='\t') %>%
+# old IDs
+study <- read.csv(file.path(path,'sample_sheet','s_hegazy_IBD.txt'),sep='\t') %>%
   dplyr::mutate(Sample.Name=gsub('p','P',Sample.Name),
-                Library.Name=gsub('p','P',Library.Name))
+                Patient.ID=gsub('[P]*([0-9]*)[\\/_][12]','P\\1',Characteristics.Patient.ID.),
+                Responder=Characteristics.Responder.,
+                Diagnosis=Characteristics.Diagnosis.,
+                Batch=gsub(' ','',gsub('_scRNAseq','',Parameter.Value.Method.)),
+                Timepoint=ifelse(grepl('_1_',Sample.Name),'Vedo_0d',
+                                 ifelse(grepl('_2_',Sample.Name),'Vedo_6wk','ctrl')),
+                Material=ifelse(Characteristics.Material.=='PBMC','PBMC','CD4')) 
+
+# new IDs
+# study <- read.csv(file.path(path,'sample_sheet.txt'),sep='\t') %>%
+#   dplyr::mutate(Patient.ID=Characteristics.Patient.ID.,
+#                 Responder=Characteristics.Responder.,
+#                 Diagnosis=Characteristics.Diagnosis.,
+#                 Batch=gsub(' ','',gsub('_scRNAseq','',Parameter.Value.Method.)),
+#                 Timepoint=ifelse(grepl('_1_',Sample.Name),'Vedo_0d',
+#                                  ifelse(grepl('_2_',Sample.Name),'Vedo_6wk','ctrl')),
+#                 Material=ifelse(Characteristics.Material.=='PBMC','PBMC','CD4')) 
 
 CD4.list <- lapply(pbmc, function(x) subset(x, predicted.celltype.l2 %in% c('CD4 CTL','CD4 Proliferating','CD4 TCM','CD4 TEM','Treg')))
 for (nn in names(CD4.list)) {
[…]

### 2024-09-25 ML_Response_prediction_visualization.rmd
20240925 Updated scripts for publication

---
title: "Visualization ML prediction"
output: html_notebook
---

```{r lib_load, message = F}
library(tidyverse)
library(magrittr)
library(rstatix)
library(reshape2)
library(ggpattern)
library(ggpubr)

theme_figures <- function() {
  theme_minimal() %+replace%    #replace elements we want to change
    theme(axis.title.x = element_blank(),
          axis.ticks.x = element_blank(),
          legend.position = "bottom",
          panel.grid.major = element_blank(),
          panel.grid.minor = element_blank(),
          panel.background = element_blank(),
          axis.line = element_line(colour = "black"),
          axis.text = element_text(
            size = 8),
          axis.title = element_text(size = 8),
          legend.title = element_text(size = 8),
          legend.text = element_text(size = 8),
          strip.text.x = element_text(size = 8)) }
```
# Model performances

## AUC values

```{r auc_read_in}
files_accurcy <- intersect(list.files(path = "../PredictiveModels/Data/", pattern = "R_response_*"),
                           list.files(path = "../PredictiveModels/Data/", pattern = "*pred_test.csv"))

combined_data <- list()

for (i in 1:length(files_accurcy)) {
  data <- t(read.csv(paste0("../PredictiveModels/Data/", files_accurcy[i]), row.names = 1))
  colnames(data) <- c('TrainingMeanAUC', 'TrainingStdAUC', 'TestingMeanAUC', 'TestingStdAUC')
  data <- t(round(data, 2))
  mat <- matrix(data, nrow = 2, ncol = 2)
  rownames(mat) <- c("MeanAUC", "SEMAUC")
  colnames(mat) <- c("Training", "Test")

  combined_data[[i]] <- mat
}
# Tidy dataframe
names(combined_data) <- gsub(pattern = "R_response_|_pred_test.csv", replacement = "", files_accurcy)
data_tweaking <- names(combined_data)[grepl(pattern = "treatment|response", x = names(combined_data))]
names(combined_data) <- ifelse(names(combined_data) %in% data_tweaking, names(combined_data), paste0(names(combined_data), "_response"))

df_accuracy <- melt(combined_data) %>%
  dcast(L1 + Var2 ~ Var1) %>%
  set_colnames(c('Condition', "Set", "MeanAUC", 'SEMAUC')) %>%
  mutate(Variable = gsub(Condition, pattern = "LinSVM_|PolynomialSVM_|RadialSVM_|LogReg_|RandomForest_", replacement = "")) %>%
  mutate(Classifier = ifelse(grepl(pattern = "treatment", Variable), "Treatment", "Response")) %>%
  mutate(Method = factor(ifelse(grepl(pattern = "LogReg", Condition), "LogReg",
                                ifelse(grepl(pattern = "LinSVM", Condition), "LinSVM",
                                       ifelse(grepl(pattern = "PolynomialSVM", Condition), "PolynomialSVM",
                                              ifelse(grepl(pattern = "RadialSVM", Condition), "RadialSVM",
                                                     ifelse(grepl(pattern = "RandomForest", Condition), "RandomForest", NA)))))), levels = c("LogReg", "LinSVM", "PolynomialSVM", "RadialSVM", "RandomForest")) %>%
  mutate(DataSet = gsub(Variable, pattern = "_response|_treatment", replacement = "")) %>%
  mutate(FeatureCutoff = ifelse(grepl("__0.05", DataSet), "95%",
                                ifelse(grepl("__0.1", DataSet), "90%",
                                       ifelse(grepl("__0.2$", DataSet), "80%",
                                              ifelse(grepl("__0.25", DataSet), "75%",
                                                     ifelse(grepl("__0.5", DataSet), "50%", "none")))))) %>%
  mutate(FeatureEngineering = ifelse(FeatureCutoff == 'none', 'No', 'Yes'))
```

```{r auc_vis_single_modalities}
# Code Supplement 18 A
pdf("../PredictiveModels/results/S_18A.pdf", height = 2, width = 4)
p <- df_accuracy %>%
  filter(DataSet %in% c("Clinical", "CyTOF", "CyTOFab", "OLINK", "FACS")) %>%
  filter(FeatureEngineering == "No") %>%
  mutate(DataSet = factor(DataSet, levels = c("Clinical", "CyTOF", "CyTOFab", "OLINK", "FACS"))) %>%
  mutate(DataSet = gsub("only", replacement = "", DataSet)) %>%
  mutate(DataSet = gsub("CyTOFfreq", replacement = "CyTOF", DataSet)) %>%
  mutate(DataSet = gsub("CyTOFA4B7", replacement = "CyTOFab", DataSet)) %>%
  mutate(DataSet = factor(DataSet, levels = c("Clinical", "CyTOF", "CyTOFab", "OLINK", "FACS"))) %>%
  mutate(Classifier = factor(Classifier, levels = c("Treatment", "Response"))) %>%
  filter(Set == "Test") %>%
  filter(Classifier == "Response") %>%
  # Data visualization -------------------------------------------
  ggplot(aes(x = DataSet, y = MeanAUC, fill = Method)) +
  geom_bar(position = "dodge", stat = "identity", width = 0.6) +
  geom_errorbar(aes(ymin = MeanAUC - SEMAUC, ymax = MeanAUC + SEMAUC), width = .2,
                position = position_dodge(.6)) +
  geom_hline(yintercept = 0.8, linetype = "dashed",
             color = "black") +
  geom_hline(yintercept = 0) +
  geom_hline(yintercept = 0.5, linetype = "dashed",
             color = "red") +
  scale_fill_manual(values = c("#DDDDDD", "#AEADAD", "#7A7A7A", "#505050", "#1B1B1B")) +
  theme_figures() +
  ylab("AUC") +
  coord_cartesian(ylim = c(0, 1))
print(p)
dev.off()
# #-------------------------------------------------------------------------------------
print(p)
```

```{r}
good_models <- df_accuracy %>%
  mutate(SelectionCriteria = MeanAUC - SEMAUC) %>%
  filter(SelectionCriteria >= 0.5) %>%
  filter(FeatureEngineering == "No") %>%
  filter(Method == "LogReg") %>%
  filter(Set == "Test") %>%
  filter(Classifier == "Response") %>%
  pull(DataSet)

# Code Supplement 18 B
pdf("../PredictiveModels/results/S_18B.pdf", height = 2.15, width = 8.1)
p <- df_accuracy %>%
  mutate(SelectionCriteria = MeanAUC - SEMAUC) %>%
  filter(DataSet %in% good_models) %>%
  filter(Method == "LogReg") %>%
  filter(Set == "Test") %>%
  filter(Classifier == "Response") %>%
  mutate(DataSet = gsub(pattern = "_", replacement = "\n", DataSet)) %>%
  mutate(FeatureCutoff = factor(FeatureCutoff, levels = c("none", "95%", "90%", "80%", "75%", "50%"))) %>%
  mutate(DataSet = factor(DataSet, levels = c("OLINK", "FACS", "OLINK\nClinical", "OLINK\nFACS", "FACS\nCyTOF", "FACS\nCyTOFab", "FACS\nCyTOF\nClinical", "FACS\nCyTOFab\nClinical", "FACS\nCyTOF\nOLINK", "FACS\nClinical\nOLINK", "all\nCyTOF"))) %>%
  ggplot(aes(x = DataSet, y = MeanAUC, fill = FeatureCutoff)) +
  geom_bar(position = "dodge", stat = "identity", width = 0.6) +
  geom_errorbar(aes(ymin = MeanAUC - SEMAUC, ymax = MeanAUC + SEMAUC), width = .2,
                position = position_dodge(.6)) +
  geom_hline(yintercept = 0) +
  geom_hline(yintercept = 0.5, linetype = "dashed",
             color = "red") +
  geom_hline(yintercept = 0.8, linetype = "dashed",
             color = "black") +
  scale_fill_manual(values = c("#000000", "#3B3B3B", "#636363", "#929292", "#C3C3C3", "#DFDFDF")) +
  theme_figures() +
  ylab("AUC") +
  coord_cartesian(ylim = c(0, 1)) +
  guides(color = guide_legend(override.aes = list(size = 1))) +
  theme(legend.key.size = unit(0.5, 'lines'))
print(p)
dev.off()

print(p)
```

```{r}
files_accurcy <- intersect(list.files(path = "../PredictiveModels/Data/", pattern = "R_response_*"),
                           list.files(path = "../PredictiveModels/Data/", pattern = "*_coefficients.csv"))

combined_data <- list()

for (i in 1:length(files_accurcy)) {
  data <- read.csv(paste0("../PredictiveModels/Data/", files_accurcy[i]))
  rownames(data) <- paste0("K-fold", 1:dim(data)[1])
  data <- data[, -1]

  combined_data[[i]] <- data
}
names(combined_data) <- gsub(pattern = "R_|_coefficients.csv", replacement = "", files_accurcy)
data_tweaking <- names(combined_data)[grepl(pattern = "treatment|response", x = names(combined_data))]
names(combined_data) <- ifelse(names(combined_data) %in% data_tweaking, names(combined_data), paste0(names(combined_data), "_response"))

df_features <- melt(combined_data) %>%
  set_colnames(c('Feature', "Coefficient", "Condition")) %>%
  mutate(Variable = gsub(Condition, pattern = "LinSVM_|PolynomialSVM_|RadialSVM_|LogReg_|RandomForest_", replacement = "")) %>%
  mutate(Classifier = ifelse(grepl(pattern = "treatment", Variable), "Treatment", "Response")) %>%
  mutate(Method = factor(ifelse(grepl(pattern = "LogReg", Condition), "LogReg",
                                ifelse(grepl(pattern = "LinSVM", Condition), "LinSVM",
                                       ifelse(grepl(pattern = "PolynomialSVM", Condition), "PolynomialSVM",
                                              ifelse(grepl(pattern = "RadialSVM", Condition), "RadialSVM",
                                                     ifelse(grepl(pattern = "RandomForest", Condition), "RandomForest", NA)))))), levels = c("LogReg", "LinSVM", "PolynomialSVM", "RadialSVM", "RandomForest")) %>%
  mutate(DataSet = gsub(Variable, pattern = "_response|_treatment", replacement = "")) %>%
  mutate(FeatureCutoff = ifelse(grepl("__0.05", DataSet), "95%",
                                ifelse(grepl("__0.1", DataSet), "90%",
                                       ifelse(grepl("__0.2$", DataSet), "80%",
                                              ifelse(grepl("__0.25", DataSet), "75%",
                                                     ifelse(grepl("__0.5", DataSet), "50%", "none")))))) %>%
  mutate(FeatureEngineering = ifelse(FeatureCutoff == 'none', 'No', 'Yes'))
```

```{r}
# Code Supplement 18D
pdf("../PredictiveModels/results/S_18C.pdf", height = 3.7, width = 4.8)
df_features %>%
  mutate(absCoef = abs(Coefficient)) %>%
  filter(Method == "LogReg") %>%
  filter(DataSet %in% good_models) %>%
  filter(RedPatientSet == 'No') %>%
  filter(FeatureCutoff == "90%") %>%
  mutate(DataSet = gsub(pattern = "_", replacement = "\n", DataSet)) %>%
  mutate(DataSet = factor(DataSet, levels = c("FACS", "OLINK\nFACS", "FACS\nCyTOF", "FACS\nCyTOFab", "OLINK\nClinical", "FACS\nCyTOF\nClinical", "FACS\nCyTOFab\nClinical", "FACS\nCyTOF\nOLINK", "FACS\nClinical\nOLINK", "all\nCyTOF"))) %>%
  mutate(Feature = factor(Feature, levels = c("Ki67_FACS", "GPR15_FACS", "CXCR3_FACS", "IL4_FACS", "CCR7_FACS", "IFNg_FACS", "Treg_FACS", "B7_FACS", "IL17A_FACS", "HLADR_FACS", 'Tbet_FACS', "CD103_FACS", "IL.10RA", "MCP.3", "TNF", "NT.3", "CCL25", "TNFB", "CASP.8", "AXIN1", "OSM", "CRP", "IL8", "NK.cells"))) %>%
  group_by(Feature, DataSet) %>%
  dplyr::count(Feature, DataSet) %>%
  mutate(n_norm = n / 10) %>%
  ggplot(aes(x = Feature, y = DataSet)) +
  geom_point() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1),
        text = element_text(size = 8)) +
  theme(axis.title.x = element_blank(),
        legend.position = "bottom",
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.background = element_blank(),
        axis.line = element_line(colour = "black"),
        axis.text = element_text(
          size = 8),
        axis.title = element_text(size = 8),
        legend.title = element_text(size = 8),
        legend.text = element_text(size = 8),
        axis.text.x = element_text(size = 8),
        strip.text.x = element_text(size = 8)) +
  coord_flip()
dev.off()
```

### 2024-09-25 ML_IterativeRun_HelperFunctionsUpdated.py

@@ -1,0 +1,250 @@
+import pandas as pd
+import numpy as np
+# from numpy import mean, std
+# import seaborn as sns
+import matplotlib.pyplot as plt
+from sklearn.metrics import roc_auc_score, RocCurveDisplay, auc
+# from sklearn.model_selection import StratifiedKFold, permutation_test_score
+from numpy import mean
+from scipy import stats
+from sklearn.model_selection import permutation_test_score
+from sklearn.preprocessing import StandardScaler
+from sklearn.svm import SVC
+from sklearn.metrics import roc_auc_score
+from sklearn.inspection import permutation_importance
+from sklearn.ensemble import RandomForestClassifier
+from sklearn.linear_model import LogisticRegression
+from sklearn.model_selection import RepeatedStratifiedKFold
+# from sklearn import linear_model
+import warnings
+
+warnings.filterwarnings('ignore')
+scaler = StandardScaler()
+
+rc_ticks = {
+    "text.usetex": False,
+    "font.size": 8,
+    "legend.fontsize": 4,
+    "legend.title_fontsize": 4,
+    "xtick.labelsize": 6,
+    "ytick.labelsize": 6,
+    "axes.labelsize": 8,
+    "axes.titlesize": 8,
+    "lines.linewidth": 0.5,
+    "axes.linewidth": 0.4,
+    "lines.markersize": 3,
+    "xtick.major.size": 2.5,
+    "xtick.major.width": 0.5,
+    "xtick.major.pad": 1,
+    "xtick.minor.size": 1.5,
+    "xtick.minor.width": 0.5,
+    "xtick.minor.pad": 1,
+    "ytick.major.size": 2.5,
+    "ytick.major.width": 0.5,
+    "ytick.major.pad": 1,
+    "ytick.minor.size": 1.5,
+    "ytick.minor.width": 0.5,
+    "ytick.minor.pad": 1,
+    "axes.labelpad": 1,
+    "axes.titlepad": 1
+}
+
+def run_k_fold_CV(X, X_scaler, y, classifier, coeff_or_not, run_perm, data_type):
+    data_set_size = len(y)
+    cv = RepeatedStratifiedKFold(n_splits=2, n_repeats=5, random_state=36851234)
+    # initiate all the empty variables 2 be filled
+    tprs = []
+    aucs = []
+    data = pd.DataFrame([])
+    y_true, y_pred_train, y_auc_test = list(), list(), list()
+    mean_fpr = np.linspace(0, 1, 100)
[…]

### 2024-09-25 ML_Iterative_kfold_model_valiation_Run.py

@@ -1,0 +1,223 @@
+exec(open("/home/lisbet/git-lisbet/code/2021_Hegazy/2022_MachineLearning/20240308_HelperFunctionsUpdated.py").read())
+import sys
+from sklearn.model_selection import train_test_split
+import random
+from sklearn.metrics import accuracy_score, f1_score
+
+# Open the file in write mode.
+sys.stdout = open('../ML_outputLog_Feature_StabilityAnalysis.txt', 'w')
+
+# clinicaliloc = [13,14,15,16,17,21]
+clinicaliloc = [12, 13, 14, 15, 16, 19]
+clinicalsubiloc = [13,  15]
+cytof_freiloc = range(22, 38)
+cytof_totaliloc = range(38, 54)
+cytof_totalAbiloc = range(54, 70)
+cytof_Abiloc = range(71, 87)
+facsiloc = range(163, 193)
+olinikiloc = range(87, 161)
+
+
+# Define a dataframe that combines each non-responder with 2 different responders in 4 runs
+# ------------------------------------------------------------------------------------------
+
+
+metadata = pd.read_csv("../Metadata_file.csv", sep=';', index_col=0, decimal=',')
+
+# Filter for timepoint 1 and VDZ treatment
+only_timepoint_one = metadata[metadata["timepoint"] == 1]
+only_timepoint_one_filtered = only_timepoint_one[
+    (only_timepoint_one["response"] == "R") | (only_timepoint_one["response"] == "NR")]
+
+
+def Run_Stability_Check(data_type_name, feature_indexes, number_iterations):
+    data_type = data_type_name
+    contrast_data = only_timepoint_one_filtered.iloc[:, np.r_[feature_indexes]]
+    contrast_data['Response'] = only_timepoint_one_filtered['response']
+    contrast_data.dropna(axis=0, how='any', inplace=True)
+
+    for i in range(0, number_iterations):
+        # random.seed(i)
+        print("=======================Iteration" + str(i) + "==========================")
+        train, test = train_test_split(contrast_data, test_size=0.2, stratify=contrast_data['Response'], random_state=i)
+        # Documentation which patients were used in held out data
+        print(test["Response"])
+        # remove the NaN entry rows = these proteins were not detected
+        X, y = Run_analysis(data_frame=train)
+        X_scaler = scaler.fit_transform(X)
+        X_scaler = scaler.transform(X)
+        print("=======================" + str(data_type) + "==========================")
+        print("=======================Linear Regression==========================")
+        print("=======================Model training")
[…]

### 2024-09-25 ML_TreatmentPrediction_ResponsePrediction_AUC_Visualization.py

@@ -1,0 +1,845 @@
+exec(open("../ML_Response_prediction_HelperScript.py").read())
+rc_ticks = {
+    "figure.figsize": (1.5, 1.2),
+    "text.usetex": False,
+    "font.size": 8,
+    "legend.fontsize":4,
+    "legend.title_fontsize":4,
+    "xtick.labelsize":6,
+    "ytick.labelsize": 6,
+    "axes.labelsize":8,
+    "axes.titlesize":8,
+    "lines.linewidth": 0.5,
+    "axes.linewidth": 0.4,
+    "lines.markersize": 3,
+    "xtick.major.size": 2.5,
+    "xtick.major.width": 0.5,
+    "xtick.major.pad": 1,
+    "xtick.minor.size": 1.5,
+    "xtick.minor.width": 0.5,
+    "xtick.minor.pad": 1,
+    "ytick.major.size": 2.5,
+    "ytick.major.width": 0.5,
+    "ytick.major.pad": 1,
+    "ytick.minor.size": 1.5,
+    "ytick.minor.width": 0.5,
+    "ytick.minor.pad": 1,
+    "axes.labelpad": 1,
+    "axes.titlepad": 1
+}
+sns.set_theme(context="talk", style="ticks", rc=rc_ticks)
+
+clinicaliloc = range(12, 14)
+cytof_freiloc = range(22, 38)
+cytof_totaliloc = range(38, 54)
+cytof_totalAbiloc = range(55, 70)
+cytof_Abiloc = range(71, 87)
+facsiloc = range(163, 193)
+olinikiloc = range(87, 161)
+
+metadata = pd.read_csv("../Metadata_file.csv", sep=';', index_col=0, decimal=',')
+
+# Code Figure 4B
+# ---------------------------------------------------------------------------------------------------
+selected_features = ['IL8', 'VEGFA', 'CD8A', 'MCP-3', 'GDNF', 'CDCP1', 'CD244', 'IL7', 'OPG',
+       'LAP TGF-beta-1', 'uPA', 'IL6', 'IL-17C', 'MCP-1', 'IL-17A', 'CXCL11',
+       'AXIN1', 'TRAIL', 'CXCL9', 'CST5', 'OSM', 'CXCL1', 'CCL4', 'CD6', 'SCF',
[…]

### 2024-09-25 ML_Treatment_effficacy_SystematicScanning.py

@@ -1,0 +1,68 @@
+exec(open("../ML_Response_prediction_HelperScript.py").read())
+import sys
+class Logger(object):
+    def __init__(self, filename="Default.log"):
+        self.terminal = sys.stdout
+        self.log = open(filename, "a")
+
+    def write(self, message):
+        self.terminal.write(message)
+        self.log.write(message)
+
+sys.stdout = Logger('../ML_outputLog_treatment_efficacy.txt')
+
+# Where are modality information stored?
+clinicaliloc = range(12, 14)
+cytof_freiloc = range(22, 38)
+cytof_totaliloc = range(38, 54)
+cytof_totalAbiloc = range(55, 70)
+cytof_Abiloc = range(71, 87)
+facsiloc = range(163, 193)
+olinikiloc = range(87, 161)
+
+metadata = pd.read_csv("../Metadata_file.csv", sep='\t', index_col=0, decimal=',')
+only_treated = metadata[metadata["treatment"] == "VDZ"]
+
+# single
+run_modality(dataframe = only_treated,input_features =olinikiloc , data_modality_name = "onlyOLINK", treatment_or_response = "treatment")
+
+contrast_data = only_treated.iloc[:, np.r_[clinicaliloc]]
+contrast_data['Response'] = only_treated['timepoint']
+X, y = Prepare_input(data_frame=contrast_data, treatment_or_response = "treatment")
+print(str("onlyClinical") + ":\n================================================")
+features = scan_over_models(X=X, y=y, data_type=str("onlyClinical"))
+
+run_modality(dataframe = only_treated,input_features =cytof_freiloc , data_modality_name = "onlyCyTOF", treatment_or_response = "treatment")
+run_modality(dataframe = only_treated,input_features =cytof_Abiloc , data_modality_name = "onlyCyTOFab", treatment_or_response = "treatment")
+run_modality(dataframe = only_treated,input_features =facsiloc , data_modality_name = "onlyFACS", treatment_or_response = "treatment")
+
+#duos
+run_modality(dataframe = only_treated,input_features =(olinikiloc,clinicaliloc) ,data_modality_name = "OLINK_Clinical", treatment_or_response = "treatment")
+run_modality(dataframe = only_treated,input_features =(olinikiloc,facsiloc) ,data_modality_name = "OLINK_FACS", treatment_or_response = "treatment")
[…]

### 2024-10-14 machine learning/ML_Iterative_model_validation_StabilitySelection.rmd
Add files via upload
Added ML code into desired folder.

---
title: "Visualizations of Classification pipeline"
author: "Lisa M. Steinheuer"
date: "`r format(Sys.time(), '%B %d, %Y')`"
abstract: |
  This script visualizes the output of the python pipeline to predict Vedolizumab treatment response in IBD patients. After the revision we decided to include LASSO and ELASTICNET to validate the feature selection process.
output:
  html_document:
    toc: true # table of content true
    toc_depth: 5  # upto five depths of headings (specified by #, ## and ###)
    number_sections: true  ## if you want number sections at each table header
    theme: united  # many options for theme, this one is my favorite.
    highlight: tango  # specifies the syntax highlighting style
---

```{r}
knitr::opts_chunk$set(message = FALSE, warning = FALSE)
knitr::opts_chunk$set(fig.width = 5, fig.height = 4)
```

# Preface

```{r lib_load, message = F}
library(tidyverse)
library(magrittr)
library(rstatix)
library(reshape2)
library(ggpubr)

project_name <- "Vedo_Stab"

theme_lisa <- function() {

  theme_minimal() %+replace%    #replace elements we want to change

    theme(axis.title.x = element_blank(),
          #axis.text.x = element_blank(),
          axis.ticks.x = element_blank(),
          legend.position = "bottom",
          panel.grid.major = element_blank(),
          panel.grid.minor = element_blank(),
          panel.background = element_blank(),
          axis.line = element_line(colour = "black"),
          axis.text = element_text(
            size = 6),
          axis.title = element_text(size = 8),
          legend.title = element_text(size = 8),
          legend.text = element_text(size = 6),
          strip.text.x = element_text(size = 6)) }

color_vector <- c("Logistic Regression" = "#762a83",
                  "LASSO" = "#9970ab",
                  "ELASTICNET" = "#e7d4e8",
                  "Linear SVM" = "#d9f0d3",
                  "Polynomial SVM" = "#a6dba0",
                  "Radial SVM" = "#5aae61",
                  "Random Forest" = "#1b7837")

color_vector_paper <- c("Logistic Regression" = "#C20000",
                        #"LASSO" = "#9970ab",
                        #"ELASTICNET" = "#e7d4e8",
                        "Linear SVM" = "#cccccc",
                        "Polynomial SVM" = "#969696",
                        "Radial SVM" = "#636363",
                        "Random Forest" = "#252525")

cols_modality <- c("FACS" = "#c01c2c",
                   "OLINK" = "#2e3e9b",
                   "CyTOF" = "#5e2590",
                   "CyTOF a4b7" = "#ab43a8",
                   "Clinical" = "#656565")

cols_feature_cutoff <- c("none" = "#252525", "50%" = "#636363", "75%" = "#969696", "80%" = "#bdbdbd", "90%" = "#C20000", "95%" = "#d9d9d9")

check_robustness <- function(n_features, threshold, reference_bootstrap, which_modality) {
  robust_features <- df_features %>%
    filter(DataSet == that_modality) %>%
    filter(Classifier == "Response") %>%
    filter(Method == "Logistic Regression") %>%
    filter(Combination == "Combination") %>%
    select(Feature, Coefficient, DataSet, Run, Origin) %>% #filter(Coefficient ! = 0 ) %>%
    mutate(absCoef = abs(Coefficient)) %>%
    group_by(DataSet, Feature, Run) %>%
    summarise(mean_features = mean(absCoef), sem_features = sd(absCoef) / sqrt(length(absCoef))) %>%
    mutate(adjusted_mean = mean_features - sem_features) %>%
    arrange(DataSet, desc(adjusted_mean)) %>%
    group_by(DataSet, Run) %>%
    top_n(n = round(n() * 0.1), wt = adjusted_mean) %>%
    mutate(rank_per_run = row_number(desc(adjusted_mean))) %>%
    mutate(Bootstrap = factor(as.numeric(Run) + 1, levels = c(1:50))) %>%
    group_by(Feature) %>%
    #mutate(Feature2 = factor(Feature, levels = order_features)) %>%
    filter(Bootstrap == reference_bootstrap) %>%
    filter(rank_per_run <= n_features) %>%
    pull(Feature)

  in_how_many_runs <- df_features %>%
    filter(DataSet == that_modality) %>%
    filter(Classifier == "Response") %>%
    filter(Method == "Logistic Regression") %>%
    filter(Combination == "Combination") %>%
    select(Feature, Coefficient, DataSet, Run, Origin) %>% #filter(Coefficient ! = 0 ) %>%
    mutate(absCoef = abs(Coefficient)) %>%
    group_by(DataSet, Feature, Run) %>%
    summarise(mean_features = mean(absCoef), sem_features = sd(absCoef) / sqrt(length(absCoef))) %>%
    mutate(adjusted_mean = mean_features - sem_features) %>%
    arrange(DataSet, desc(adjusted_mean)) %>%
    group_by(DataSet, Run) %>%
    top_n(n = round(n() * threshold), wt = adjusted_mean) %>%
    mutate(rank_per_run = row_number(desc(adjusted_mean))) %>%
    mutate(Bootstrap = factor(as.numeric(Run) + 1, levels = c(1:50))) %>%
    filter(Bootstrap != reference_bootstrap) %>%
    group_by(Bootstrap) %>%
    summarise(contains_features = all(robust_features %in% Feature)) %>%
    summarise(count = sum(contains_features))
  output <- round(in_how_many_runs$count * 100 / 49, 2)
  return(output)

}

```

# Evaluation

## Model performances

```{r auc_read_in}

# Data read in
files_accurcy <- intersect(list.files(path = "/Iterative_model_validation/", pattern = "R_*"),
                           list.files(path = "/Iterative_model_validation/", pattern = "*_testing.csv"))

only_individual <- files_accurcy[grepl(pattern = "only", files_accurcy)]

combined_data <- list()

for (i in 1:length(files_accurcy)) {
  data <- t(read.csv(paste0("/Iterative_model_validation/", files_accurcy[i]), row.names = 1))

  combined_data[[i]] <- data
}
# Data tweaking
names(combined_data) <- gsub(pattern = "R_|_testing.csv", replacement = "", files_accurcy)

df_testing_performance <- melt(combined_data) %>%
  dcast(L1 + Var2 ~ Var1) %>%
  set_colnames(c('Condition', "remove_col", "Run", "Accuracy", 'F1')) %>%
  select(-remove_col) %>%
  mutate(Variable = gsub(Condition, pattern = "LogisticRegression_|LASSO_|ELASTICNET_|ELASTICNET_|LinearSVM_|PolynomialSVM_|RadialSVM_|RandomForest_", replacement = "")) %>%
  mutate(Method = factor(ifelse(grepl(pattern = "LogisticRegression_", Condition), "Logistic Regression",
                                ifelse(grepl(pattern = "LASSO_", Condition), "LASSO",
                                       ifelse(grepl(pattern = "ELASTICNET_", Condition), "ELASTICNET",
                                              ifelse(grepl(pattern = "LinearSVM_", Condition), "Linear SVM",
                                                     ifelse(grepl(pattern = "PolynomialSVM_", Condition), "Polynomial SVM",
                                                            ifelse(grepl(pattern = "RadialSVM_", Condition), "Radial SVM",
                                                                   ifelse(grepl(pattern = "RandomForest_", Condition), "Random Forest", NA))))))), levels = c("Logistic Regression", "LASSO", "ELASTICNET", "Linear SVM", "Polynomial SVM", "Radial SVM", "Random Forest"))) %>%
  mutate(DataSet = gsub(Variable, pattern = "LogisticRegression_|LASSO_|ELASTICNET_|ELASTICNET_|LinearSVM_|PolynomialSVM_|RadialSVM_|RandomForest_", replacement = "")) %>%
  mutate(DataSet = gsub("^_", replacement = "", DataSet)) %>%
  mutate(Run = gsub(pattern = ".*_Run", replacement = "", x = DataSet)) %>%
  mutate(DataSet = gsub(pattern = "_Run[0-9]|_Run[0-9][0-9]", replacement = "", DataSet)) %>%
  mutate(Combination = ifelse(DataSet %in% c("Clinical", "CyTOF", "CyTOFa4b7", "OLINK", "FACS"), yes = "Single", no = "Combination")) %>%
  mutate(Bootstrap = factor(as.numeric(Run) + 1, levels = c(1:50)))

```

## The model coefficients

```{r prepare_coef}
files_accurcy <- intersect(list.files(path = "/Iterative_model_validation/", pattern = "R_*"),
                           list.files(path = "/Iterative_model_validation/", pattern = "*_coefficients.csv"))

only_individual <- files_accurcy[grepl(pattern = "only", files_accurcy)]

combined_data <- list()

for (i in 1:length(files_accurcy)) {
  data <- read.csv(paste0("/Iterative_model_validation/", files_accurcy[i]))
  rownames(data) <- paste0("K-fold", 1:dim(data)[1])
  data <- data[, -1]

  combined_data[[i]] <- data
}
names(combined_data) <- gsub(pattern = "R_|_coefficients.csv", replacement = "", files_accurcy)
data_tweaking <- names(combined_data)[grepl(pattern = "treatment|response", x = names(combined_data))]
names(combined_data) <- ifelse(names(combined_data) %in% data_tweaking, names(combined_data), paste0(names(combined_data), "_response"))

df_features <- melt(combined_data) %>%
  #dcast(L1 + Var2 ~ Var1) %>%
  set_colnames(c('Feature', "Coefficient", "Condition")) %>%
  mutate(Variable = gsub(Condition, pattern = "LinSVM_|PolynomialSVM_|RadialSVM_|LogReg_|RandomForest_", replacement = "")) %>%
  mutate(Classifier = ifelse(grepl(pattern = "treatment", Variable), "Treatment", "Response")) %>%
  mutate(Method = factor(ifelse(grepl(pattern = "LogisticRegression_", Condition), "Logistic Regression",
                                ifelse(grepl(pattern = "LASSO_", Condition), "LASSO",
                                       ifelse(grepl(pattern = "ELASTICNET_", Condition), "ELASTICNET",
                                              ifelse(grepl(pattern = "LinearSVM_", Condition), "Linear SVM",
                                                     ifelse(grepl(pattern = "PolynomialSVM_", Condition), "Polynomial SVM",
                                                            ifelse(grepl(pattern = "RadialSVM_", Condition), "Radial SVM",
                                                                   ifelse(grepl(pattern = "RandomForest_", Condition), "Random Forest", NA))))))), levels = c("Logistic Regression", "LASSO", "ELASTICNET", "Linear SVM", "Polynomial SVM", "Radial SVM", "Random Forest"))) %>%
  mutate(DataSet = gsub(Variable, pattern = "_response|_treatment", replacement = "")) %>%
  mutate(DataSet = gsub(DataSet, pattern = "LogisticRegression_|LASSO_|ELASTICNET_|ELASTICNET_|LinearSVM_|PolynomialSVM_|RadialSVM_|RandomForest_", replacement = "")) %>%
  mutate(DataSet = gsub("^_", replacement = "", DataSet)) %>%
  mutate(Run = gsub(pattern = ".*_Run", replacement = "", x = DataSet)) %>%
  mutate(DataSet = gsub(pattern = "_Run[0-9]|_Run[0-9][0-9]", replacement = "", DataSet)) %>%
  mutate(Combination = ifelse(DataSet %in% c("Clinical", "CyTOF", "CyTOFa4b7", "OLINK", "FACS"), yes = "Single", no = "Combination"))
```

## Adding Origin

```{r fig.height = 3}
df_features <- df_features %>%
  mutate(Origin = factor(ifelse(grepl(patter = "_FACS", x = Feature), yes = "FACS", no =
    ifelse(grepl(patter = "_A4B7", x = Feature), yes = "CyTOF a4b7", no =
      ifelse(Feature %in% c("Basophils", "cDCs", "cMonocytes", "Eosinophils", "gd.T.cells", "Memory.B.cells", "Naive.B.cells", "MAIT.cells", "Memory.CD4", "Memory.CD8", "Naive.CD4", "Naive.CD8", "ncMonocytes", "NK.cells", "pDCs", "Plasmablasts"), yes = "CyTOF", no =
        ifelse(Feature %in% c('Leukocytes', 'CRP', 'Thrombocytes', 'Hemoglobin', 'Neutrophils', 'TNF_pretreated'), yes = "Clinical", no = "OLINK")))), levels = c("FACS", "OLINK", "CyTOF", "CyTOF a4b7", "Clinical")))

```

This code chunk checks how often a feature was included in the top 10 features from the non-feature selected data. I used the non-feature engineered data because otherwise I cannot extract the top 10 features per run. The smallest feature space was 22.
All in all, we have 18 model combinations.
I implemented it in a way, that we have for each non-responder (4) 4 different responder pairs

## Functions

```{r fig.width = 15, fig.height = 10}

Analyse_Modality_overview <- function(which_modality, which_classifier) {

  if (which_classifier %in% c("LASSO", "ELASTICNET")) {
    order_features <- df_features %>%
      filter(DataSet == that_modality) %>%
      filter(Classifier == "Response") %>%
      filter(Method == which_classifier) %>%
      filter(Combination == "Combination") %>%
      select(Feature, Coefficient, DataSet, Run, Origin) %>%
      filter(Coefficient != 0) %>%
      mutate(absCoef = abs(Coefficient)) %>%
      group_by(DataSet, Feature, Run) %>%
      summarise(mean_features = mean(absCoef), sem_features = sd(absCoef) / sqrt(length(absCoef))) %>%
      mutate(adjusted_mean = mean_features - sem_features) %>%
      arrange(DataSet, desc(adjusted_mean)) %>%
      group_by(Feature, Run) %>%
      dplyr::count(Feature) %>%
      mutate(Bootstrap = factor(as.numeric(Run) + 1, levels = c(1:50))) %>%
      group_by(Feature) %>%
      add_count(name = "Frequency") %>%
      arrange(-Frequency) %>%
      pull(Feature) %>%
      unique()

    p1 <- df_features %>%
      filter(DataSet == which_modality) %>%
      filter(Classifier == "Response") %>%
      filter(Method == which_classifier) %>%
      filter(Combination == "Combination") %>%
      select(Feature, Coefficient, DataSet, Run, Origin) %>%
      filter(Coefficient != 0) %>%
      mutate(absCoef = abs(Coefficient)) %>%
      group_by(DataSet, Feature, Run) %>%
      summarise(mean_features = mean(absCoef), sem_features = sd(absCoef) / sqrt(length(absCoef))) %>%
      mutate(adjusted_mean = mean_features - sem_features) %>%
      arrange(DataSet, desc(adjusted_mean)) %>%
      mutate(rank_per_run = row_number(desc(adjusted_mean))) %>%
      mutate(Bootstrap = factor(as.numeric(Run) + 1, levels = c(1:50))) %>%
      group_by(Feature) %>%
      mutate(Feature2 = factor(Feature, levels = order_features)) %>%
      ggplot(aes(x = Feature2, y = Bootstrap, fill = rank_per_run)) +
      geom_tile() +
      theme_minimal() +
      geom_text(aes(label = rank_per_run), size = 2.3) +
      theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
      scale_fill_gradient("Rank", low = "#737373", high = "#f0f0f0", limits = c(0, 10)) +
      xlab("Features") +
      ylab("Within \n each iteration") +
      theme(axis.title.x = element_blank(),
            axis.text.x = element_blank(),
            axis.ticks.x = element_blank()) +
      theme(legend.position = "none")

  } else if (which_classifier == "Logistic Regression") {
    order_features <- df_features %>%
      filter(DataSet == which_modality) %>%
      filter(Classifier == "Response") %>%
      filter(Method == which_classifier) %>%
      filter(Combination == "Combination") %>%
      select(Feature, Coefficient, DataSet, Run, Origin) %>%
      mutate(absCoef = abs(Coefficient)) %>%
      group_by(DataSet, Feature, Run) %>%
      summarise(mean_features = mean(absCoef), sem_features = sd(absCoef) / sqrt(length(absCoef))) %>%
      mutate(adjusted_mean = mean_features - sem_features) %>%
      arrange(DataSet, desc(adjusted_mean)) %>%
      group_by(DataSet, Run) %>%
      top_n(n = round(n() * 0.1), wt = adjusted_mean) %>%
      group_by(Feature, Run) %>%
      dplyr::count(Feature) %>%
      mutate(Bootstrap = factor(as.numeric(Run) + 1, levels = c(1:50))) %>%
      group_by(Feature) %>%
      add_count(name = "Frequency") %>%
      arrange(-Frequency) %>%
      pull(Feature) %>%
      unique()

    p1 <- df_features %>%
      filter(DataSet == which_modality) %>%
      filter(Classifier == "Response") %>%
      filter(Method == which_classifier) %>%
      filter(Combination == "Combination") %>%
      select(Feature, Coefficient, DataSet, Run, Origin) %>% #filter(Coefficient ! = 0 ) %>%
      mutate(absCoef = abs(Coefficient)) %>%
      group_by(DataSet, Feature, Run) %>%
      summarise(mean_features = mean(absCoef), sem_features = sd(absCoef) / sqrt(length(absCoef))) %>%
      mutate(adjusted_mean = mean_features - sem_features) %>%
      arrange(DataSet, desc(adjusted_mean)) %>%
      group_by(DataSet, Run) %>%
      top_n(n = round(n() * 0.1), wt = adjusted_mean) %>%
      mutate(rank_per_run = row_number(desc(adjusted_mean))) %>%
      mutate(Bootstrap = factor(as.numeric(Run) + 1, levels = c(1:50))) %>%
      group_by(Feature) %>%
      mutate(Feature2 = factor(Feature, levels = order_features)) %>%
      ggplot(aes(x = Feature2, y = Bootstrap, fill = rank_per_run)) +
      geom_tile() +
      theme_minimal() +
      geom_text(aes(label = rank_per_run), size = 2.3) +
      theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
      scale_fill_gradient("Rank", low = "#737373", high = "#f0f0f0", limits = c(0, 10)) +
      xlab("Features") +
      ylab("Within \n each iteration") +
      theme(axis.title.x = element_blank(),
            axis.text.x = element_blank(),
            axis.ticks.x = element_blank()) +
      theme(legend.position = "none") }

  p2 <- df_features %>%
    filter(DataSet == which_modality) %>%
    filter(Classifier == "Response") %>%
    filter(Method == which_classifier) %>%
    filter(Combination == "Combination") %>%
    select(Feature, Coefficient, DataSet, Run, Origin) %>%
    filter(Feature %in% order_features) %>%
    mutate(Feature2 = factor(Feature, levels = order_features)) %>%
    ggplot(aes(x = Feature2, y = 1, fill = Origin)) +
    geom_tile() +
    theme_minimal() +
    theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
    scale_fill_manual(values = cols_modality) +
    ylab("") +
    xlab("Feature") +
    theme(axis.title.y = element_blank(),
          axis.text.y = element_blank(),
          axis.ticks.y = element_blank()) +
    theme(legend.position = "bottom")

  p3 <- df_testing_performance %>%
    filter(DataSet == which_modality) %>%
    filter(Method == which_classifier) %>%
    filter(Combination == "Combination") %>%
    ggplot(aes(x = Method, y = Bootstrap, fill = Accuracy)) +
    geom_tile() +
    geom_text(aes(label = round(Accuracy, 1))) +
    scale_fill_gradient(low = "white", high = "#fc9272", limits = c(0, 1)) +
    theme_minimal() +
    theme(axis.text.x = element_blank(), axis.title.x = element_blank()) +
    theme(axis.title.y = element_blank(),
          # axis.text.y = element_blank(),
          axis.ticks.y = element_blank())

  p4 <- cowplot::plot_grid(p1, p3, p2, labels = c('', '', ''), ncol = 2, rel_heights = c(0.8, 0.3), rel_widths = c(1, 0.15), align = 'v')

  ggsave(p4, file = paste0("../", project_name, "_Stability_Top10_", which_modality, "_", which_classifier, ".pdf"), width = 14, height = 9)

  print(p4)
  return }
```

```{r}
run_stability_analysis <- function(which_modality, rerun) {

  rerun_script <- rerun

  if (rerun_script == 1) {
    stability_landscape <- data.frame("Number of features" = integer(), "Threshold (top X%)" = integer(), "Reference Run" = integer(), "Stability" = integer())
    for (run in 1:50) {
      stability_landscape_iteration <- data.frame("Number of features" = integer(), "Threshold (top X%)" = integer(), "Reference Run" = integer(), "Stability" = integer())
      for (n_feat in 1:10) {
        for (cutoff in seq(0.1, 1.0, by = 0.1)) {
          how_many_features <- n_feat
          define_threshold <- cutoff
          define_reference <- run
          stability_value <- check_robustness(n_features = how_many_features, threshold = define_threshold, reference_bootstrap = define_reference, which_modality = that_modality)
          stability_landscape_int <- data.frame("Number of features" = how_many_features, "Threshold (top X%)" = define_threshold * 100, "Reference Run" = define_reference, "Stability" = stability_value)
          stability_landscape_iteration <- rbind(stability_landscape_iteration, stability_landscape_int) } }

      print(mean(stability_landscape$Stability))

      stability_landscape <- rbind(stability_landscape, stability_landscape_iteration)
    }
    save(stability_landscape, file = paste0('../StabilityRun_', that_modality, ".RData"))

  p1 <- stability_landscape %>%
    set_colnames(c("Number of features", "Threshold (top X%)", "Reference Run", "Stability")) %>%
    group_by(`Reference Run`) %>%
    summarise(`Mean Stability` = mean(Stability)) %>%
    ungroup() %>%
    ggplot(aes(x = 1, y = `Reference Run`, fill = `Mean Stability`)) +
    geom_tile() +
    theme_minimal() +
    geom_text(aes(label = round(`Mean Stability`, 2)), size = 3.5) +
    theme(axis.title.x = element_blank(),
          axis.text.x = element_blank(),
          axis.ticks.x = element_blank()) +
    scale_fill_gradient("Stability", low = "#fee5d9", high = "#a50f15", limits = c(0, 100)) +
    scale_x_continuous(breaks = seq(1, 1, by = 1)) +
    scale_y_continuous(breaks = seq(0, 100, by = 10)) +
    ylab(paste("Iterations"))

  print(p1)
  ggsave(p1, file = paste0("../", project_name, "_StabilityAnalysis_", that_modality, "_Run.pdf"), width = 2, height = 10)

  refine_reference_run <- stability_landscape %>%
    set_colnames(c("Number of features", "Threshold (top X%)", "Reference Run", "Stability")) %>%
    group_by(`Reference Run`) %>%
    summarise(`Mean Stability` = mean(Stability)) %>%
    filter(`Mean Stability` == max(`Mean Stability`)) %>%
    slice(1)

  print(refine_reference_run)

  p2 <- stability_landscape %>%
    set_colnames(c("Number of features", "Threshold (top X%)", "Reference Run", "Stability")) %>%
    filter(`Reference Run` == refine_reference_run$`Reference Run`) %>%
    ggplot(aes(x = `Number of features`, y = `Threshold (top X%)`, fill = Stability)) +
    geom_tile() +
    theme_minimal() +
    geom_text(aes(label = Stability), size = 2.5) +
    scale_fill_gradient("Stability", low = "#fee5d9", high = "#a50f15", limits = c(0, 100)) +
    scale_x_continuous(breaks = seq(1, 10, by = 1)) +
    scale_y_continuous(breaks = seq(10, 100, by = 10)) +
    xlab(paste("Number of features")) +
    labs(title = paste("Reference Run", refine_reference_run$`Reference Run`))
  ggsave(p2, file = paste0("../StabilityAnalysis_", that_modality, "_RunRef.pdf"), width = 5, height = 4)
  data.frame(stability_landscape %>%
               set_colnames(c("Number of features", "Threshold (top X%)", "Reference Run", "Stability")) %>%
               filter(`Reference Run` == refine_reference_run$`Reference Run`) %>%
               arrange(desc(Stability), `Threshold (top X%)`) %>%
               filter(`Number of features` >= 5) %>%
               filter(`Threshold (top X%)` < 40))

  return(stability_landscape) }

```

## 4 best modalities

### OLINK_FACS

```{r fig.width = 15, fig.height = 10}
that_modality <- "OLINK_FACS"
that_classifier <- "Logistic Regression"
Analyse_Modality_overview(which_modality = that_modality, which_classifier = that_classifier)

number_of_features <- df_features %>%
  filter(DataSet == that_modality) %>%
  filter(Classifier == "Response") %>%
  filter(Method == "Logistic Regression") %>%
  filter(Combination == "Combination") %>%
  select(Feature, Coefficient, DataSet, Run, Origin) %>% #filter(Coefficient ! = 0 ) %>%
  mutate(absCoef = abs(Coefficient)) %>%
  group_by(DataSet, Feature, Run) %>%
  summarise(mean_features = mean(absCoef), sem_features = sd(absCoef) / sqrt(length(absCoef))) %>%
  mutate(adjusted_mean = mean_features - sem_features) %>%
  arrange(DataSet, desc(adjusted_mean)) %>%
  group_by(DataSet, Run) %>%
  top_n(n = round(n() * 0.1), wt = adjusted_mean) %>%
  filter(Run == 0) %>%
  pull(Feature) %>%
  length()
auc_landscape <- data.frame("Number of features" = integer(), "Reference Run" = integer(), "AUC" = integer())

for (run in 1:50) {
  for (n_feat in 1:number_of_features) {
    stability_landscape_iteration <- data.frame("Number of features" = integer(), "Threshold (top X%)" = integer(), "Reference Run" = integer(), "Included" = integer(), "Excluded" = integer())
    for (cutoff in seq(0, 1.0, by = 0.1)) {
      robust_features <- df_features %>%
        filter(DataSet == that_modality) %>%
        filter(Classifier == "Response") %>%
        filter(Method == "Logistic Regression") %>%
        filter(Combination == "Combination") %>%
        select(Feature, Coefficient, DataSet, Run, Origin) %>% #filter(Coefficient ! = 0 ) %>%
        mutate(absCoef = abs(Coefficient)) %>%
        group_by(DataSet, Feature, Run) %>%
        summarise(mean_features = mean(absCoef), sem_features = sd(absCoef) / sqrt(length(absCoef))) %>%
        mutate(adjusted_mean = mean_features - sem_features) %>%
        arrange(DataSet, desc(adjusted_mean)) %>%
        group_by(DataSet, Run) %>%
        top_n(n = round(n() * 0.1), wt = adjusted_mean) %>%
        mutate(rank_per_run = row_number(desc(adjusted_mean))) %>%
        mutate(Bootstrap = factor(as.numeric(Run) + 1, levels = c(1:50))) %>%
        group_by(Feature) %>%
        #mutate(Feature2 = factor(Feature, levels = order_features)) %>%
        filter(Bootstrap == run) %>%
        filter(rank_per_run <= n_feat) %>%
        pull(Feature)

      stability_value <- df_features %>%
        filter(DataSet == that_modality) %>%
        filter(Classifier == "Response") %>%
        filter(Method == "Logistic Regression") %>%
        filter(Combination == "Combination") %>%
        select(Feature, Coefficient, DataSet, Run, Origin) %>% #filter(Coefficient ! = 0 ) %>%
        mutate(absCoef = abs(Coefficient)) %>%
        group_by(DataSet, Feature, Run) %>%
        summarise(mean_features = mean(absCoef), sem_features = sd(absCoef) / sqrt(length(absCoef))) %>%
        mutate(adjusted_mean = mean_features - sem_features) %>%
        arrange(DataSet, desc(adjusted_mean)) %>%
        group_by(DataSet, Run) %>%
        top_n(n = round(n() * cutoff), wt = adjusted_mean) %>%
        mutate(rank_per_run = row_number(desc(adjusted_mean))) %>%
        mutate(Bootstrap = factor(as.numeric(Run) + 1, levels = c(1:50))) %>%
        filter(Bootstrap != run) %>%
        group_by(Bootstrap) %>%
        summarise(contains_features = all(robust_features %in% Feature)) %>%
        summarise(count = sum(contains_features)) %>%
        mutate(Included = count / 49) %>%
        mutate(Excluded = (49 - count) / 49)

      stability_landscape_int <- data.frame("Number of features" = n_feat, "Threshold (top X%)" = cutoff, "Reference Run" = run, "Included" = stability_value$Included, "Excluded" = stability_value$Excluded)
      stability_landscape_iteration <- rbind(stability_landscape_iteration, stability_landscape_int) }
    auc_it <- sum(diff(stability_landscape_iteration$Threshold..top.X..) *
                    (stability_landscape_iteration$Included[-1] + stability_landscape_iteration$Included[-length(stability_landscape_iteration$Included)]) / 2)
    auc_landscape_it <- data.frame("Number of features" = n_feat, "Reference Run" = run, "AUC" = auc_it)
    auc_landscape <- rbind(auc_landscape, auc_landscape_it)
  }
}

save(auc_landscape, file = paste0('../StabilityRun_', that_modality, ".RData"))

p1 <- auc_landscape %>%
  set_colnames(c("Number of features", "Reference Run", "AUC")) %>%
  group_by(`Reference Run`) %>%
  summarise(`Mean AUC` = mean(AUC)) %>%
  ungroup() %>%
  ggplot(aes(x = 1, y = `Reference Run`, fill = `Mean AUC`)) +
  geom_tile() +
  theme_minimal() +
  geom_text(aes(label = round(`Mean AUC`, 2)), size = 3.5) +
  theme(axis.title.x = element_blank(),
        axis.text.x = element_blank(),
        axis.ticks.x = element_blank()) +
  scale_fill_gradient("AUC", low = "#fee5d9", high = "#a50f15", limits = c(0, 1)) +
  scale_x_continuous(breaks = seq(1, 1, by = 1)) +
  scale_y_continuous(breaks = seq(0, 100, by = 10)) +
  ylab(paste("Iterations"))

print(p1)
ggsave(p1, file = paste0("../", project_name, "_StabilityAnalysis_", that_modality, "_Run.pdf"), width = 2, height = 10)

refine_reference_run <- auc_landscape %>%
  set_colnames(c("Number of features", "Reference Run", "AUC")) %>%
  group_by(`Reference Run`) %>%
  summarise(`Mean AUC` = mean(AUC)) %>%
  filter(`Mean AUC` == max(`Mean AUC`)) %>%
  slice(1)

print(refine_reference_run)

p2 <- auc_landscape %>%
  set_colnames(c("Number of features", "Reference Run", "AUC")) %>%
  filter(`Reference Run` == refine_reference_run$`Reference Run`) %>%
  ggplot(aes(x = `Number of features`, y = AUC, fill = AUC)) +
  geom_bar(stat = "identity") +
  theme_minimal() +
  geom_hline(yintercept = 0.7, linetype = "dashed", color = "red") +
  scale_fill_gradient("AUC", low = "#fee5d9", high = "#a50f15", limits = c(0, 1)) +
  scale_x_continuous(breaks = seq(1, number_of_features, by = 1)) +
  scale_y_continuous(limits = c(0, 1), breaks = seq(0, 1, by = 0.25)) +
  xlab(paste("Number of features")) +
  labs(title = paste("Reference Run", refine_reference_run$`Reference Run`))
ggsave(p2, file = paste0("~/Desktop/", project_name, "_StabilityAnalysis_", that_modality, "_RunRef.pdf"), width = 5, height = 2)

how_many_features <- 5

feature_panel_olink_facs <- df_features %>%
  filter(DataSet == that_modality) %>%
  filter(Classifier == "Response") %>%
  filter(Method == that_classifier) %>%
  filter(Combination == "Combination") %>%
  select(Feature, Coefficient, DataSet, Run, Origin) %>% #filter(Coefficient ! = 0 ) %>%
  mutate(absCoef = abs(Coefficient)) %>%
  group_by(DataSet, Feature, Run) %>%
  summarise(mean_features = mean(absCoef), sem_features = sd(absCoef) / sqrt(length(absCoef))) %>%
  mutate(adjusted_mean = mean_features - sem_features) %>%
  arrange(DataSet, desc(adjusted_mean)) %>%
  group_by(DataSet, Run) %>%
  top_n(n = how_many_features, wt = adjusted_mean) %>%
  mutate(rank_per_run = row_number(desc(adjusted_mean))) %>%
  mutate(Bootstrap = factor(as.numeric(Run) + 1, levels = c(1:50))) %>%
  group_by(Feature) %>%
  filter(Bootstrap == refine_reference_run$`Reference Run`) %>%
  pull(Feature)

print(feature_panel_olink_facs)

feature_info_OLINK_FACS <- df_features %>%
  filter(DataSet == that_modality) %>%
  filter(Classifier == "Response") %>%
  filter(Method == that_classifier) %>%
  filter(Combination == "Combination") %>%
  select(Feature, Coefficient, DataSet, Run, Origin) %>% #filter(Coefficient ! = 0 ) %>%
  mutate(absCoef = abs(Coefficient)) %>%
  mutate(Bootstrap = factor(as.numeric(Run) + 1, levels = c(1:50))) %>%
  filter(Bootstrap == refine_reference_run$`Reference Run`) %>%
  filter(Feature %in% feature_panel_olink_facs)

p1 <- df_features %>%
  filter(DataSet == that_modality) %>%
  filter(Classifier == "Response") %>%
  filter(Method == that_classifier) %>%
  filter(Combination == "Combination") %>%
  select(Feature, Coefficient, DataSet, Run, Origin) %>% #filter(Coefficient ! = 0 ) %>%
  mutate(absCoef = abs(Coefficient)) %>%
  mutate(Bootstrap = factor(as.numeric(Run) + 1, levels = c(1:50))) %>%
  filter(Bootstrap == refine_reference_run$`Reference Run`) %>%
  ungroup %>%
  filter(Feature %in% feature_panel_olink_facs) %>%
  mutate(Feature = factor(Feature, levels = feature_panel_olink_facs)) %>%
  ggplot(aes(x = Feature, y = absCoef)) +
  geom_boxplot() +
  expand_limits(y = c(0, 0.3)) +
  theme_minimal() +
  ylab("Absolute Feature Coefficient") +
  xlab("Features") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))

print(p1)

ggsave(p1, file = paste0("../", project_name, "_Bestfeatures_", that_modality, ".pdf"), width = 3, height = 2)

p1 <- df_features %>%
  filter(DataSet == that_modality) %>%
  filter(Classifier == "Response") %>%
  filter(Method == that_classifier) %>%
  filter(Combination == "Combination") %>%
  select(Feature, Coefficient, DataSet, Run, Origin) %>% #filter(Coefficient ! = 0 ) %>%
  mutate(Bootstrap = factor(as.numeric(Run) + 1, levels = c(1:50))) %>%
  filter(Bootstrap == refine_reference_run$`Reference Run`) %>%
  ungroup %>%
  filter(Feature %in% feature_panel_olink_facs) %>%
  mutate(Feature = factor(Feature, levels = feature_panel_olink_facs)) %>%
  ggplot(aes(x = Feature, y = Coefficient)) +
  geom_boxplot() +
  expand_limits(y = c(0, 0.3)) +
  theme_minimal() +
  ylab("Absolute Feature Coefficient") +
  xlab("Features") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))

print(p1)

```

#### Using LASS0

```{r fig.width = 15, fig.height = 10}
Analyse_Modality_overview(which_modality = that_modality, which_classifier = "LASSO")
df_features %>%
  filter(DataSet == that_modality) %>%
  filter(Classifier == "Response") %>%
  filter(Method == "LASSO") %>%
  filter(Combination == "Combination") %>%
  select(Feature, Coefficient, DataSet, Run, Origin) %>%
  filter(Coefficient != 0) %>%
  mutate(absCoef = abs(Coefficient)) %>%
  group_by(DataSet, Feature, Run) %>%
  summarise(mean_features = mean(absCoef), sem_features = sd(absCoef) / sqrt(length(absCoef))) %>%
  mutate(adjusted_mean = mean_features - sem_features) %>%
  arrange(DataSet, desc(adjusted_mean)) %>%
  mutate(rank_per_run = row_number(desc(adjusted_mean))) %>%
  mutate(Bootstrap = factor(as.numeric(Run) + 1, levels = c(1:50))) %>%
  group_by(Bootstrap) %>%
  summarise(contains_features = all(feature_panel_olink_facs %in% Feature)) %>%
  summarise(count = sum(contains_features))

lasso_run_olink_facs <- df_features %>%
  filter(DataSet == that_modality) %>%
  filter(Classifier == "Response") %>%
  filter(Method == "LASSO") %>%
  filter(Combination == "Combination") %>%
  select(Feature, Coefficient, DataSet, Run, Origin) %>%
  filter(Coefficient != 0) %>%
  mutate(absCoef = abs(Coefficient)) %>%
  group_by(DataSet, Feature, Run) %>%
  summarise(mean_features = mean(absCoef), sem_features = sd(absCoef) / sqrt(length(absCoef))) %>%
  mutate(adjusted_mean = mean_features - sem_features) %>%
  arrange(DataSet, desc(adjusted_mean)) %>%
  mutate(rank_per_run = row_number(desc(adjusted_mean))) %>%
  mutate(Bootstrap = factor(as.numeric(Run) + 1, levels = c(1:50))) %>%
  filter(Bootstrap == refine_reference_run$`Reference Run`) %>%
  select(DataSet, Feature) %>%
  mutate(Method = "LASSO")

intersect(unique(lasso_run_olink_facs$Feature), feature_panel_olink_facs)

```

#### Using ELASTICNET

```{r fig.width = 15, fig.height = 10}
Analyse_Modality_overview(which_modality = that_modality, which_classifier = "ELASTICNET")
df_features %>%
  filter(DataSet == that_modality) %>%
  filter(Classifier == "Response") %>%
  filter(Method == "ELASTICNET") %>%
  filter(Combination == "Combination") %>%
  select(Feature, Coefficient, DataSet, Run, Origin) %>%
  filter(Coefficient != 0) %>%
  mutate(absCoef = abs(Coefficient)) %>%
  group_by(DataSet, Feature, Run) %>%
  summarise(mean_features = mean(absCoef), sem_features = sd(absCoef) / sqrt(length(absCoef))) %>%
  mutate(adjusted_mean = mean_features - sem_features) %>%
  arrange(DataSet, desc(adjusted_mean)) %>%
  mutate(rank_per_run = row_number(desc(adjusted_mean))) %>%
  mutate(Bootstrap = factor(as.numeric(Run) + 1, levels = c(1:50))) %>%
  group_by(Bootstrap) %>%
  summarise(contains_features = all(feature_panel_olink_facs %in% Feature)) %>%
  summarise(count = sum(contains_features))

el_run_olink_facs <- df_features %>%
  filter(DataSet == that_modality) %>%
  filter(Classifier == "Response") %>%
  filter(Method == "ELASTICNET") %>%
  filter(Combination == "Combination") %>%
  select(Feature, Coefficient, DataSet, Run, Origin) %>%
  filter(Coefficient != 0) %>%
  mutate(absCoef = abs(Coefficient)) %>%
  group_by(DataSet, Feature, Run) %>%
  summarise(mean_features = mean(absCoef), sem_features = sd(absCoef) / sqrt(length(absCoef))) %>%
  mutate(adjusted_mean = mean_features - sem_features) %>%
  arrange(DataSet, desc(adjusted_mean)) %>%
  mutate(rank_per_run = row_number(desc(adjusted_mean))) %>%
  mutate(Bootstrap = factor(as.numeric(Run) + 1, levels = c(1:50))) %>%
  filter(Bootstrap == refine_reference_run$`Reference Run`) %>%
  select(DataSet, Feature) %>%
  mutate(Method = "ELASTICNET")

intersect(unique(el_run_olink_facs$Feature), feature_panel_olink_facs)
```

```{r}
summary_reg_olink_facs <- rbind(el_run_olink_facs, lasso_run_olink_facs) %>%
  group_by(DataSet, Feature) %>%
  summarise(
    Methods = paste(unique(Method), collapse = ","),
    .groups = 'drop'
  ) %>%
  mutate(Occurrence = case_when(
    Methods == "ELASTICNET,LASSO" ~ "Both",
    Methods == "ELASTICNET" ~ "ELASTICNET only",
    Methods == "LASSO" ~ "LASSO only",
    TRUE ~ "None"
  ))
```

### OLINK_FACS_CyTOF

```{r fig.width = 15, fig.height = 10}
that_modality <- "OLINK_FACS_CyTOF"
that_classifier <- "Logistic Regression"
Analyse_Modality_overview(which_modality = that_modality, which_classifier = that_classifier)

number_of_features <- df_features %>%
  filter(DataSet == that_modality) %>%
  filter(Classifier == "Response") %>%
  filter(Method == "Logistic Regression") %>%
  filter(Combination == "Combination") %>%
  select(Feature, Coefficient, DataSet, Run, Origin) %>% #filter(Coefficient ! = 0 ) %>%
  mutate(absCoef = abs(Coefficient)) %>%
  group_by(DataSet, Feature, Run) %>%
  summarise(mean_features = mean(absCoef), sem_features = sd(absCoef) / sqrt(length(absCoef))) %>%
  mutate(adjusted_mean = mean_features - sem_features) %>%
  arrange(DataSet, desc(adjusted_mean)) %>%
  group_by(DataSet, Run) %>%
  top_n(n = round(n() * 0.1), wt = adjusted_mean) %>%
  filter(Run == 0) %>%
  pull(Feature) %>%
  length()
auc_landscape <- data.frame("Number of features" = integer(), "Reference Run" = integer(), "AUC" = integer())

for (run in 1:50) {
  for (n_feat in 1:number_of_features) {
    stability_landscape_iteration <- data.frame("Number of features" = integer(), "Threshold (top X%)" = integer(), "Reference Run" = integer(), "Included" = integer(), "Excluded" = integer())
    for (cutoff in seq(0, 1.0, by = 0.1)) {
      robust_features <- df_features %>%
        filter(DataSet == that_modality) %>%
        filter(Classifier == "Response") %>%
        filter(Method == "Logistic Regression") %>%
        filter(Combination == "Combination") %>%
        select(Feature, Coefficient, DataSet, Run, Origin) %>% #filter(Coefficient ! = 0 ) %>%
        mutate(absCoef = abs(Coefficient)) %>%
        group_by(DataSet, Feature, Run) %>%
        summarise(mean_features = mean(absCoef), sem_features = sd(absCoef) / sqrt(length(absCoef))) %>%
        mutate(adjusted_mean = mean_features - sem_features) %>%
        arrange(DataSet, desc(adjusted_mean)) %>%
        group_by(DataSet, Run) %>%
        top_n(n = round(n() * 0.1), wt = adjusted_mean) %>%
        mutate(rank_per_run = row_number(desc(adjusted_mean))) %>%
        mutate(Bootstrap = factor(as.numeric(Run) + 1, levels = c(1:50))) %>%
        group_by(Feature) %>%
        #mutate(Feature2 = factor(Feature, levels = order_features)) %>%
        filter(Bootstrap == run) %>%
        filter(rank_per_run <= n_feat) %>%
        pull(Feature)

      stability_value <- df_features %>%
        filter(DataSet == that_modality) %>%
        filter(Classifier == "Response") %>%
        filter(Method == "Logistic Regression") %>%
        filter(Combination == "Combination") %>%
        select(Feature, Coefficient, DataSet, Run, Origin) %>% #filter(Coefficient ! = 0 ) %>%
        mutate(absCoef = abs(Coefficient)) %>%
        group_by(DataSet, Feature, Run) %>%
        summarise(mean_features = mean(absCoef), sem_features = sd(absCoef) / sqrt(length(absCoef))) %>%
        mutate(adjusted_mean = mean_features - sem_features) %>%
        arrange(DataSet, desc(adjusted_mean)) %>%
        group_by(DataSet, Run) %>%
        top_n(n = round(n() * cutoff), wt = adjusted_mean) %>%
        mutate(rank_per_run = row_number(desc(adjusted_mean))) %>%
        mutate(Bootstrap = factor(as.numeric(Run) + 1, levels = c(1:50))) %>%
        filter(Bootstrap != run) %>%
        group_by(Bootstrap) %>%
        summarise(contains_features = all(robust_features %in% Feature)) %>%
        summarise(count = sum(contains_features)) %>%
        mutate(Included = count / 49) %>%
        mutate(Excluded = (49 - count) / 49)

      stability_landscape_int <- data.frame("Number of features" = n_feat, "Threshold (top X%)" = cutoff, "Reference Run" = run, "Included" = stability_value$Included, "Excluded" = stability_value$Excluded)
      stability_landscape_iteration <- rbind(stability_landscape_iteration, stability_landscape_int) }
    auc_it <- sum(diff(stability_landscape_iteration$Threshold..top.X..) *
                    (stability_landscape_iteration$Included[-1] + stability_landscape_iteration$Included[-length(stability_landscape_iteration$Included)]) / 2)
    auc_landscape_it <- data.frame("Number of features" = n_feat, "Reference Run" = run, "AUC" = auc_it)
    auc_landscape <- rbind(auc_landscape, auc_landscape_it)
  }
}

save(auc_landscape, file = paste0('../StabilityRun_', that_modality, ".RData"))
#

p1 <- auc_landscape %>%
  set_colnames(c("Number of features", "Reference Run", "AUC")) %>%
  group_by(`Reference Run`) %>%
  summarise(`Mean AUC` = mean(AUC)) %>%
  ungroup() %>%
  ggplot(aes(x = 1, y = `Reference Run`, fill = `Mean AUC`)) +
  geom_tile() +
  theme_minimal() +
  geom_text(aes(label = round(`Mean AUC`, 2)), size = 3.5) +
  theme(axis.title.x = element_blank(),
        axis.text.x = element_blank(),
        axis.ticks.x = element_blank()) +
  scale_fill_gradient("AUC", low = "#fee5d9", high = "#a50f15", limits = c(0, 1)) +
  scale_x_continuous(breaks = seq(1, 1, by = 1)) +
  scale_y_continuous(breaks = seq(0, 100, by = 10)) +
  ylab(paste("Iterations"))

print(p1)
ggsave(p1, file = paste0("../", project_name, "_StabilityAnalysis_", that_modality, "_Run.pdf"), width = 2, height = 10)

refine_reference_run <- auc_landscape %>%
  set_colnames(c("Number of features", "Reference Run", "AUC")) %>%
  group_by(`Reference Run`) %>%
  summarise(`Mean AUC` = mean(AUC)) %>%
  filter(`Mean AUC` == max(`Mean AUC`)) %>%
  slice(1)

print(refine_reference_run)

p2 <- auc_landscape %>%
  set_colnames(c("Number of features", "Reference Run", "AUC")) %>%
  filter(`Reference Run` == refine_reference_run$`Reference Run`) %>%
  ggplot(aes(x = `Number of features`, y = AUC, fill = AUC)) +
  geom_bar(stat = "identity") +
  theme_minimal() +
  #geom_text(aes(label = AUC), size = 2.5) +
  geom_hline(yintercept = 0.7, linetype = "dashed", color = "red") +
  scale_fill_gradient("AUC", low = "#fee5d9", high = "#a50f15", limits = c(0, 1)) +
  scale_x_continuous(breaks = seq(1, number_of_features, by = 1)) +
  scale_y_continuous(limits = c(0, 1), breaks = seq(0, 1, by = 0.25)) +
  #scale_y_continuous(breaks = seq(10, 100, by = 10)) +
  xlab(paste("Number of features")) +
  labs(title = paste("Reference Run", refine_reference_run$`Reference Run`))
ggsave(p2, file = paste0("~/Desktop/", project_name, "_StabilityAnalysis_", that_modality, "_RunRef.pdf"), width = 5, height = 2)

how_many_features <- 5

feature_panel_OLINK_FACS_CyTOF <- df_features %>%
  filter(DataSet == that_modality) %>%
  filter(Classifier == "Response") %>%
  filter(Method == that_classifier) %>%
  filter(Combination == "Combination") %>%
  select(Feature, Coefficient, DataSet, Run, Origin) %>% #filter(Coefficient ! = 0 ) %>%
  mutate(absCoef = abs(Coefficient)) %>%
  group_by(DataSet, Feature, Run) %>%
  summarise(mean_features = mean(absCoef), sem_features = sd(absCoef) / sqrt(length(absCoef))) %>%
  mutate(adjusted_mean = mean_features - sem_features) %>%
  arrange(DataSet, desc(adjusted_mean)) %>%
  group_by(DataSet, Run) %>%
  top_n(n = how_many_features, wt = adjusted_mean) %>%
  mutate(rank_per_run = row_number(desc(adjusted_mean))) %>%
  mutate(Bootstrap = factor(as.numeric(Run) + 1, levels = c(1:50))) %>%
  group_by(Feature) %>%
  filter(Bootstrap == refine_reference_run$`Reference Run`) %>%
  pull(Feature)

feature_info_OLINK_FACS_CyTOF <- df_features %>%
  filter(DataSet == that_modality) %>%
  filter(Classifier == "Response") %>%
  filter(Method == that_classifier) %>%
  filter(Combination == "Combination") %>%
  select(Feature, Coefficient, DataSet, Run, Origin) %>% #filter(Coefficient ! = 0 ) %>%
  mutate(absCoef = abs(Coefficient)) %>%
  mutate(Bootstrap = factor(as.numeric(Run) + 1, levels = c(1:50))) %>%
  filter(Bootstrap == refine_reference_run$`Reference Run`) %>%
  filter(Feature %in% feature_panel_OLINK_FACS_CyTOF)

print(feature_panel_OLINK_FACS_CyTOF)

p1 <- df_features %>%
  filter(DataSet == that_modality) %>%
  filter(Classifier == "Response") %>%
  filter(Method == that_classifier) %>%
  filter(Combination == "Combination") %>%
  select(Feature, Coefficient, DataSet, Run, Origin) %>% #filter(Coefficient ! = 0 ) %>%
  mutate(absCoef = abs(Coefficient)) %>%
  mutate(Bootstrap = factor(as.numeric(Run) + 1, levels = c(1:50))) %>%
  filter(Bootstrap == refine_reference_run$`Reference Run`) %>%
  ungroup %>%
  filter(Feature %in% feature_panel_OLINK_FACS_CyTOF) %>%
  mutate(Feature = factor(Feature, levels = feature_panel_OLINK_FACS_CyTOF)) %>%
  ggplot(aes(x = Feature, y = absCoef)) +
  geom_boxplot() +
  expand_limits(y = c(0, 0.3)) +
  theme_minimal() +
  ylab("Absolute Feature Coefficient") +
  xlab("Features") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))

print(p1)

ggsave(p1, file = paste0("../", project_name, "_Bestfeatures_", that_modality, ".pdf"), width = 3, height = 2)

p1 <- df_features %>%
  filter(DataSet == that_modality) %>%
  filter(Classifier == "Response") %>%
  filter(Method == that_classifier) %>%
  filter(Combination == "Combination") %>%
  select(Feature, Coefficient, DataSet, Run, Origin) %>% #filter(Coefficient ! = 0 ) %>%

  mutate(Bootstrap = factor(as.numeric(Run) + 1, levels = c(1:50))) %>%
  filter(Bootstrap == refine_reference_run$`Reference Run`) %>%
  ungroup %>%
  filter(Feature %in% feature_panel_OLINK_FACS_CyTOF) %>%
  mutate(Feature = factor(Feature, levels = feature_panel_OLINK_FACS_CyTOF)) %>%
  ggplot(aes(x = Feature, y = Coefficient)) +
  geom_boxplot() +
  expand_limits(y = c(0, 0.3)) +
  theme_minimal() +
  ylab("Absolute Feature Coefficient") +
  xlab("Features") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))

print(p1)

```

#### Using LASS0

```{r fig.width = 15, fig.height = 10}
Analyse_Modality_overview(which_modality = that_modality, which_classifier = "LASSO")
df_features %>%
  filter(DataSet == that_modality) %>%
  filter(Classifier == "Response") %>%
  filter(Method == "LASSO") %>%
  filter(Combination == "Combination") %>%
  select(Feature, Coefficient, DataSet, Run, Origin) %>%
  filter(Coefficient != 0) %>%
  mutate(absCoef = abs(Coefficient)) %>%
  group_by(DataSet, Feature, Run) %>%
  summarise(mean_features = mean(absCoef), sem_features = sd(absCoef) / sqrt(length(absCoef))) %>%
  mutate(adjusted_mean = mean_features - sem_features) %>%
  arrange(DataSet, desc(adjusted_mean)) %>%
  mutate(rank_per_run = row_number(desc(adjusted_mean))) %>%
  mutate(Bootstrap = factor(as.numeric(Run) + 1, levels = c(1:50))) %>%
  group_by(Bootstrap) %>%
  summarise(contains_features = all(feature_panel_OLINK_FACS_CyTOF %in% Feature)) %>%
  summarise(count = sum(contains_features))

lasso_run_OLINK_FACS_CyTOF <- df_features %>%
  filter(DataSet == that_modality) %>%
  filter(Classifier == "Response") %>%
  filter(Method == "LASSO") %>%
  filter(Combination == "Combination") %>%
  select(Feature, Coefficient, DataSet, Run, Origin) %>%
  filter(Coefficient != 0) %>%
  mutate(absCoef = abs(Coefficient)) %>%
  group_by(DataSet, Feature, Run) %>%
  summarise(mean_features = mean(absCoef), sem_features = sd(absCoef) / sqrt(length(absCoef))) %>%
  mutate(adjusted_mean = mean_features - sem_features) %>%
  arrange(DataSet, desc(adjusted_mean)) %>%
  mutate(rank_per_run = row_number(desc(adjusted_mean))) %>%
  mutate(Bootstrap = factor(as.numeric(Run) + 1, levels = c(1:50))) %>%
  filter(Bootstrap == refine_reference_run$`Reference Run`) %>%
  select(DataSet, Feature) %>%
  mutate(Method = "LASSO")

intersect(unique(lasso_run_OLINK_FACS_CyTOF$Feature), feature_panel_OLINK_FACS_CyTOF)
```

#### Using ELASTICNET

```{r fig.width = 15, fig.height = 10}
Analyse_Modality_overview(which_modality = that_modality, which_classifier = "ELASTICNET")
df_features %>%
  filter(DataSet == that_modality) %>%
  filter(Classifier == "Response") %>%
  filter(Method == "ELASTICNET") %>%
  filter(Combination == "Combination") %>%
  select(Feature, Coefficient, DataSet, Run, Origin) %>%
  filter(Coefficient != 0) %>%
  mutate(absCoef = abs(Coefficient)) %>%
  group_by(DataSet, Feature, Run) %>%
  summarise(mean_features = mean(absCoef), sem_features = sd(absCoef) / sqrt(length(absCoef))) %>%
  mutate(adjusted_mean = mean_features - sem_features) %>%
  arrange(DataSet, desc(adjusted_mean)) %>%
  mutate(rank_per_run = row_number(desc(adjusted_mean))) %>%
  mutate(Bootstrap = factor(as.numeric(Run) + 1, levels = c(1:50))) %>%
  group_by(Bootstrap) %>%
  summarise(contains_features = all(feature_panel_OLINK_FACS_CyTOF %in% Feature)) %>%
  summarise(count = sum(contains_features))

el_run_OLINK_FACS_CyTOF <- df_features %>%
  filter(DataSet == that_modality) %>%
  filter(Classifier == "Response") %>%
  filter(Method == "ELASTICNET") %>%
  filter(Combination == "Combination") %>%
  select(Feature, Coefficient, DataSet, Run, Origin) %>%
  filter(Coefficient != 0) %>%
  mutate(absCoef = abs(Coefficient)) %>%
  group_by(DataSet, Feature, Run) %>%
  summarise(mean_features = mean(absCoef), sem_features = sd(absCoef) / sqrt(length(absCoef))) %>%
  mutate(adjusted_mean = mean_features - sem_features) %>%
  arrange(DataSet, desc(adjusted_mean)) %>%
  mutate(rank_per_run = row_number(desc(adjusted_mean))) %>%
  mutate(Bootstrap = factor(as.numeric(Run) + 1, levels = c(1:50))) %>%
  filter(Bootstrap == refine_reference_run$`Reference Run`) %>%
  select(DataSet, Feature) %>%
  mutate(Method = "ELASTICNET")

intersect(unique(el_run_OLINK_FACS_CyTOF$Feature), feature_panel_OLINK_FACS_CyTOF)
```
```{r}
summary_reg_OLINK_FACS_CyTOF <- rbind(el_run_OLINK_FACS_CyTOF, lasso_run_OLINK_FACS_CyTOF) %>%
  group_by(DataSet, Feature) %>%
  summarise(
    Methods = paste(unique(Method), collapse = ","),
    .groups = 'drop'
  ) %>%
  mutate(Occurrence = case_when(
    Methods == "ELASTICNET,LASSO" ~ "Both",
    Methods == "ELASTICNET" ~ "ELASTICNET only",
    Methods == "LASSO" ~ "LASSO only",
    TRUE ~ "None"
  ))
```

### FACS_CyTOF_Clinical

```{r fig.width = 15, fig.height = 10}
that_modality <- "FACS_CyTOF_Clinical"
that_classifier <- "Logistic Regression"
Analyse_Modality_overview(which_modality = that_modality, which_classifier = that_classifier)
number_of_features <- df_features %>%
  filter(DataSet == that_modality) %>%
  filter(Classifier == "Response") %>%
  filter(Method == "Logistic Regression") %>%
  filter(Combination == "Combination") %>%
  select(Feature, Coefficient, DataSet, Run, Origin) %>% #filter(Coefficient ! = 0 ) %>%
  mutate(absCoef = abs(Coefficient)) %>%
  group_by(DataSet, Feature, Run) %>%
  summarise(mean_features = mean(absCoef), sem_features = sd(absCoef) / sqrt(length(absCoef))) %>%
  mutate(adjusted_mean = mean_features - sem_features) %>%
  arrange(DataSet, desc(adjusted_mean)) %>%
  group_by(DataSet, Run) %>%
  top_n(n = round(n() * 0.1), wt = adjusted_mean) %>%
  filter(Run == 0) %>%
  pull(Feature) %>%
  length()
auc_landscape <- data.frame("Number of features" = integer(), "Reference Run" = integer(), "AUC" = integer())

for (run in 1:50) {
  for (n_feat in 1:number_of_features) {
    stability_landscape_iteration <- data.frame("Number of features" = integer(), "Threshold (top X%)" = integer(), "Reference Run" = integer(), "Included" = integer(), "Excluded" = integer())
    for (cutoff in seq(0, 1.0, by = 0.1)) {
      robust_features <- df_features %>%
        filter(DataSet == that_modality) %>%
        filter(Classifier == "Response") %>%
        filter(Method == "Logistic Regression") %>%
        filter(Combination == "Combination") %>%
        select(Feature, Coefficient, DataSet, Run, Origin) %>% #filter(Coefficient ! = 0 ) %>%
        mutate(absCoef = abs(Coefficient)) %>%
        group_by(DataSet, Feature, Run) %>%
        summarise(mean_features = mean(absCoef), sem_features = sd(absCoef) / sqrt(length(absCoef))) %>%
        mutate(adjusted_mean = mean_features - sem_features) %>%
        arrange(DataSet, desc(adjusted_mean)) %>%
        group_by(DataSet, Run) %>%
        top_n(n = round(n() * 0.1), wt = adjusted_mean) %>%
        mutate(rank_per_run = row_number(desc(adjusted_mean))) %>%
        mutate(Bootstrap = factor(as.numeric(Run) + 1, levels = c(1:50))) %>%
        group_by(Feature) %>%
        #mutate(Feature2 = factor(Feature, levels = order_features)) %>%
        filter(Bootstrap == run) %>%
        filter(rank_per_run <= n_feat) %>%
        pull(Feature)

      stability_value <- df_features %>%
        filter(DataSet == that_modality) %>%
        filter(Classifier == "Response") %>%
        filter(Method == "Logistic Regression") %>%
        filter(Combination == "Combination") %>%
        select(Feature, Coefficient, DataSet, Run, Origin) %>% #filter(Coefficient ! = 0 ) %>%
        mutate(absCoef = abs(Coefficient)) %>%
        group_by(DataSet, Feature, Run) %>%
        summarise(mean_features = mean(absCoef), sem_features = sd(absCoef) / sqrt(length(absCoef))) %>%
        mutate(adjusted_mean = mean_features - sem_features) %>%
        arrange(DataSet, desc(adjusted_mean)) %>%
        group_by(DataSet, Run) %>%
        top_n(n = round(n() * cutoff), wt = adjusted_mean) %>%
        mutate(rank_per_run = row_number(desc(adjusted_mean))) %>%
        mutate(Bootstrap = factor(as.numeric(Run) + 1, levels = c(1:50))) %>%
        filter(Bootstrap != run) %>%
        group_by(Bootstrap) %>%
        summarise(contains_features = all(robust_features %in% Feature)) %>%
        summarise(count = sum(contains_features)) %>%
        mutate(Included = count / 49) %>%
        mutate(Excluded = (49 - count) / 49)

      stability_landscape_int <- data.frame("Number of features" = n_feat, "Threshold (top X%)" = cutoff, "Reference Run" = run, "Included" = stability_value$Included, "Excluded" = stability_value$Excluded)
      stability_landscape_iteration <- rbind(stability_landscape_iteration, stability_landscape_int) }
    auc_it <- sum(diff(stability_landscape_iteration$Threshold..top.X..) *
                    (stability_landscape_iteration$Included[-1] + stability_landscape_iteration$Included[-length(stability_landscape_iteration$Included)]) / 2)
    auc_landscape_it <- data.frame("Number of features" = n_feat, "Reference Run" = run, "AUC" = auc_it)
    auc_landscape <- rbind(auc_landscape, auc_landscape_it)
  }
}

save(auc_landscape, file = paste0('../StabilityRun_', that_modality, ".RData"))

p1 <- auc_landscape %>%
  set_colnames(c("Number of features", "Reference Run", "AUC")) %>%
  group_by(`Reference Run`) %>%
  summarise(`Mean AUC` = mean(AUC)) %>%
  ungroup() %>%
  ggplot(aes(x = 1, y = `Reference Run`, fill = `Mean AUC`)) +
  geom_tile() +
  theme_minimal() +
  geom_text(aes(label = round(`Mean AUC`, 2)), size = 3.5) +
  theme(axis.title.x = element_blank(),
        axis.text.x = element_blank(),
        axis.ticks.x = element_blank()) +
  scale_fill_gradient("AUC", low = "#fee5d9", high = "#a50f15", limits = c(0, 1)) +
  scale_x_continuous(breaks = seq(1, 1, by = 1)) +
  scale_y_continuous(breaks = seq(0, 100, by = 10)) +
  ylab(paste("Iterations"))

print(p1)
ggsave(p1, file = paste0("../", project_name, "_StabilityAnalysis_", that_modality, "_Run.pdf"), width = 2, height = 10)

refine_reference_run <- auc_landscape %>%
  set_colnames(c("Number of features", "Reference Run", "AUC")) %>%
  group_by(`Reference Run`) %>%
  summarise(`Mean AUC` = mean(AUC)) %>%
  filter(`Mean AUC` == max(`Mean AUC`)) %>%
  slice(1)

print(refine_reference_run)

p2 <- auc_landscape %>%
  set_colnames(c("Number of features", "Reference Run", "AUC")) %>%
  filter(`Reference Run` == refine_reference_run$`Reference Run`) %>%
  ggplot(aes(x = `Number of features`, y = AUC, fill = AUC)) +
  geom_bar(stat = "identity") +
  theme_minimal() +
  geom_hline(yintercept = 0.7, linetype = "dashed", color = "red") +
  scale_fill_gradient("AUC", low = "#fee5d9", high = "#a50f15", limits = c(0, 1)) +
  scale_x_continuous(breaks = seq(1, number_of_features, by = 1)) +
  scale_y_continuous(limits = c(0, 1), breaks = seq(0, 1, by = 0.25)) +
  xlab(paste("Number of features")) +
  labs(title = paste("Reference Run", refine_reference_run$`Reference Run`))
ggsave(p2, file = paste0("../", project_name, "_StabilityAnalysis_", that_modality, "_RunRef.pdf"), width = 5, height = 2)

how_many_features <- 5

feature_panel_FACS_CyTOF_Clinical <- df_features %>%
  filter(DataSet == that_modality) %>%
  filter(Classifier == "Response") %>%
  filter(Method == that_classifier) %>%
  filter(Combination == "Combination") %>%
  select(Feature, Coefficient, DataSet, Run, Origin) %>% #filter(Coefficient ! = 0 ) %>%
  mutate(absCoef = abs(Coefficient)) %>%
  group_by(DataSet, Feature, Run) %>%
  summarise(mean_features = mean(absCoef), sem_features = sd(absCoef) / sqrt(length(absCoef))) %>%
  mutate(adjusted_mean = mean_features - sem_features) %>%
  arrange(DataSet, desc(adjusted_mean)) %>%
  group_by(DataSet, Run) %>%
  top_n(n = how_many_features, wt = adjusted_mean) %>%
  mutate(rank_per_run = row_number(desc(adjusted_mean))) %>%
  mutate(Bootstrap = factor(as.numeric(Run) + 1, levels = c(1:50))) %>%
  group_by(Feature) %>%
  filter(Bootstrap == refine_reference_run$`Reference Run`) %>%
  pull(Feature)

print(feature_panel_FACS_CyTOF_Clinical)

feature_info_FACS_CyTOF_Clinical <- df_features %>%
  filter(DataSet == that_modality) %>%
  filter(Classifier == "Response") %>%
  filter(Method == that_classifier) %>%
  filter(Combination == "Combination") %>%
  select(Feature, Coefficient, DataSet, Run, Origin) %>% #filter(Coefficient ! = 0 ) %>%
  mutate(absCoef = abs(Coefficient)) %>%
  mutate(Bootstrap = factor(as.numeric(Run) + 1, levels = c(1:50))) %>%
  filter(Bootstrap == refine_reference_run$`Reference Run`) %>%
  filter(Feature %in% feature_panel_FACS_CyTOF_Clinical)

p1 <- df_features %>%
  filter(DataSet == that_modality) %>%
  filter(Classifier == "Response") %>%
  filter(Method == that_classifier) %>%
  filter(Combination == "Combination") %>%
  select(Feature, Coefficient, DataSet, Run, Origin) %>% #filter(Coefficient ! = 0 ) %>%
  mutate(absCoef = abs(Coefficient)) %>%
  mutate(Bootstrap = factor(as.numeric(Run) + 1, levels = c(1:50))) %>%
  filter(Bootstrap == refine_reference_run$`Reference Run`) %>%
  ungroup %>%
  filter(Feature %in% feature_info_FACS_CyTOF_Clinical) %>%
  mutate(Feature = factor(Feature, levels = feature_panel_FACS_CyTOF_Clinical)) %>%
  ggplot(aes(x = Feature, y = absCoef)) +
  geom_boxplot() +
  expand_limits(y = c(0, 0.3)) +
  theme_minimal() +
  ylab("Absolute Feature Coefficient") +
  xlab("Features") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))

print(p1)

ggsave(p1, file = paste0("../", project_name, "_Bestfeatures_", that_modality, ".pdf"), width = 3, height = 2)

p1 <- df_features %>%
  filter(DataSet == that_modality) %>%
  filter(Classifier == "Response") %>%
  filter(Method == that_classifier) %>%
  filter(Combination == "Combination") %>%
  select(Feature, Coefficient, DataSet, Run, Origin) %>% #filter(Coefficient ! = 0 ) %>%
  #mutate(absCoef = abs(Coefficient)) %>%
  mutate(Bootstrap = factor(as.numeric(Run) + 1, levels = c(1:50))) %>%
  filter(Bootstrap == refine_reference_run$`Reference Run`) %>%
  ungroup %>%
  filter(Feature %in% feature_panel_FACS_CyTOF_Clinical) %>%
  mutate(Feature = factor(Feature, levels = feature_panel_FACS_CyTOF_Clinical)) %>%
  ggplot(aes(x = Feature, y = Coefficient)) +
  geom_boxplot() +
  expand_limits(y = c(0, 0.3)) +
  theme_minimal() +
  ylab("Absolute Feature Coefficient") +
  xlab("Features") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))

print(p1)

```

#### Using LASS0

```{r fig.width = 15, fig.height = 10}
Analyse_Modality_overview(which_modality = that_modality, which_classifier = "LASSO")
df_features %>%
  filter(DataSet == that_modality) %>%
  filter(Classifier == "Response") %>%
  filter(Method == "LASSO") %>%
  filter(Combination == "Combination") %>%
  select(Feature, Coefficient, DataSet, Run, Origin) %>%
  filter(Coefficient != 0) %>%
  mutate(absCoef = abs(Coefficient)) %>%
  group_by(DataSet, Feature, Run) %>%
  summarise(mean_features = mean(absCoef), sem_features = sd(absCoef) / sqrt(length(absCoef))) %>%
  mutate(adjusted_mean = mean_features - sem_features) %>%
  arrange(DataSet, desc(adjusted_mean)) %>%
  mutate(rank_per_run = row_number(desc(adjusted_mean))) %>%
  mutate(Bootstrap = factor(as.numeric(Run) + 1, levels = c(1:50))) %>%
  group_by(Bootstrap) %>%
  summarise(contains_features = all(feature_panel_FACS_CyTOF_Clinical %in% Feature)) %>%
  summarise(count = sum(contains_features))

lasso_run_FACS_CyTOF_Clinical <- df_features %>%
  filter(DataSet == that_modality) %>%
  filter(Classifier == "Response") %>%
  filter(Method == "LASSO") %>%
  filter(Combination == "Combination") %>%
  select(Feature, Coefficient, DataSet, Run, Origin) %>%
  filter(Coefficient != 0) %>%
  mutate(absCoef = abs(Coefficient)) %>%
  group_by(DataSet, Feature, Run) %>%
  summarise(mean_features = mean(absCoef), sem_features = sd(absCoef) / sqrt(length(absCoef))) %>%
  mutate(adjusted_mean = mean_features - sem_features) %>%
  arrange(DataSet, desc(adjusted_mean)) %>%
  mutate(rank_per_run = row_number(desc(adjusted_mean))) %>%
  mutate(Bootstrap = factor(as.numeric(Run) + 1, levels = c(1:50))) %>%
  filter(Bootstrap == refine_reference_run$`Reference Run`) %>%
  select(DataSet, Feature) %>%
  mutate(Method = "LASSO")

intersect(unique(lasso_run_FACS_CyTOF_Clinical$Feature), feature_panel_FACS_CyTOF_Clinical)
```

#### Using ELASTICNET

```{r fig.width = 15, fig.height = 10}
Analyse_Modality_overview(which_modality = that_modality, which_classifier = "ELASTICNET")
df_features %>%
  filter(DataSet == that_modality) %>%
  filter(Classifier == "Response") %>%
  filter(Method == "ELASTICNET") %>%
  filter(Combination == "Combination") %>%
  select(Feature, Coefficient, DataSet, Run, Origin) %>%
  filter(Coefficient != 0) %>%
  mutate(absCoef = abs(Coefficient)) %>%
  group_by(DataSet, Feature, Run) %>%
  summarise(mean_features = mean(absCoef), sem_features = sd(absCoef) / sqrt(length(absCoef))) %>%
  mutate(adjusted_mean = mean_features - sem_features) %>%
  arrange(DataSet, desc(adjusted_mean)) %>%
  mutate(rank_per_run = row_number(desc(adjusted_mean))) %>%
  mutate(Bootstrap = factor(as.numeric(Run) + 1, levels = c(1:50))) %>%
  group_by(Bootstrap) %>%
  summarise(contains_features = all(feature_panel_FACS_CyTOF_Clinical %in% Feature)) %>%
  summarise(count = sum(contains_features))

el_run_FACS_CyTOF_Clinical <- df_features %>%
  filter(DataSet == that_modality) %>%
  filter(Classifier == "Response") %>%
  filter(Method == "ELASTICNET") %>%
  filter(Combination == "Combination") %>%
  select(Feature, Coefficient, DataSet, Run, Origin) %>%
  filter(Coefficient != 0) %>%
  mutate(absCoef = abs(Coefficient)) %>%
  group_by(DataSet, Feature, Run) %>%
  summarise(mean_features = mean(absCoef), sem_features = sd(absCoef) / sqrt(length(absCoef))) %>%
  mutate(adjusted_mean = mean_features - sem_features) %>%
  arrange(DataSet, desc(adjusted_mean)) %>%
  mutate(rank_per_run = row_number(desc(adjusted_mean))) %>%
  mutate(Bootstrap = factor(as.numeric(Run) + 1, levels = c(1:50))) %>%
  filter(Bootstrap == refine_reference_run$`Reference Run`) %>%
  select(DataSet, Feature) %>%
  mutate(Method = "ELASTICNET")

intersect(unique(el_run_FACS_CyTOF_Clinical$Feature), feature_panel_FACS_CyTOF_Clinical)
```

```{r}
summary_reg_FACS_CyTOF_Clinical <- rbind(el_run_FACS_CyTOF_Clinical, lasso_run_FACS_CyTOF_Clinical) %>%
  group_by(DataSet, Feature) %>%
  summarise(
    Methods = paste(unique(Method), collapse = ","),
    .groups = 'drop'
  ) %>%
  mutate(Occurrence = case_when(
    Methods == "ELASTICNET,LASSO" ~ "Both",
    Methods == "ELASTICNET" ~ "ELASTICNET only",
    Methods == "LASSO" ~ "LASSO only",
    TRUE ~ "None"
  ))
```

### FACS_CyTOFab

```{r fig.width = 15, fig.height = 10}
that_modality <- "FACS_CyTOFa4b7"
that_classifier <- "Logistic Regression"
Analyse_Modality_overview(which_modality = that_modality, which_classifier = that_classifier)
number_of_features <- df_features %>%
  filter(DataSet == that_modality) %>%
  filter(Classifier == "Response") %>%
  filter(Method == "Logistic Regression") %>%
  filter(Combination == "Combination") %>%
  select(Feature, Coefficient, DataSet, Run, Origin) %>% #filter(Coefficient ! = 0 ) %>%
  mutate(absCoef = abs(Coefficient)) %>%
  group_by(DataSet, Feature, Run) %>%
  summarise(mean_features = mean(absCoef), sem_features = sd(absCoef) / sqrt(length(absCoef))) %>%
  mutate(adjusted_mean = mean_features - sem_features) %>%
  arrange(DataSet, desc(adjusted_mean)) %>%
  group_by(DataSet, Run) %>%
  top_n(n = round(n() * 0.1), wt = adjusted_mean) %>%
  filter(Run == 0) %>%
  pull(Feature) %>%
  length()
auc_landscape <- data.frame("Number of features" = integer(), "Reference Run" = integer(), "AUC" = integer())

for (run in 1:50) {
  for (n_feat in 1:number_of_features) {
    stability_landscape_iteration <- data.frame("Number of features" = integer(), "Threshold (top X%)" = integer(), "Reference Run" = integer(), "Included" = integer(), "Excluded" = integer())
    for (cutoff in seq(0, 1.0, by = 0.1)) {
      robust_features <- df_features %>%
        filter(DataSet == that_modality) %>%
        filter(Classifier == "Response") %>%
        filter(Method == "Logistic Regression") %>%
        filter(Combination == "Combination") %>%
        select(Feature, Coefficient, DataSet, Run, Origin) %>% #filter(Coefficient ! = 0 ) %>%
        mutate(absCoef = abs(Coefficient)) %>%
        group_by(DataSet, Feature, Run) %>%
        summarise(mean_features = mean(absCoef), sem_features = sd(absCoef) / sqrt(length(absCoef))) %>%
        mutate(adjusted_mean = mean_features - sem_features) %>%
        arrange(DataSet, desc(adjusted_mean)) %>%
        group_by(DataSet, Run) %>%
        top_n(n = round(n() * 0.1), wt = adjusted_mean) %>%
        mutate(rank_per_run = row_number(desc(adjusted_mean))) %>%
        mutate(Bootstrap = factor(as.numeric(Run) + 1, levels = c(1:50))) %>%
        group_by(Feature) %>%
        filter(Bootstrap == run) %>%
        filter(rank_per_run <= n_feat) %>%
        pull(Feature)

      stability_value <- df_features %>%
        filter(DataSet == that_modality) %>%
        filter(Classifier == "Response") %>%
        filter(Method == "Logistic Regression") %>%
        filter(Combination == "Combination") %>%
        select(Feature, Coefficient, DataSet, Run, Origin) %>% #filter(Coefficient ! = 0 ) %>%
        mutate(absCoef = abs(Coefficient)) %>%
        group_by(DataSet, Feature, Run) %>%
        summarise(mean_features = mean(absCoef), sem_features = sd(absCoef) / sqrt(length(absCoef))) %>%
        mutate(adjusted_mean = mean_features - sem_features) %>%
        arrange(DataSet, desc(adjusted_mean)) %>%
        group_by(DataSet, Run) %>%
        top_n(n = round(n() * cutoff), wt = adjusted_mean) %>%
        mutate(rank_per_run = row_number(desc(adjusted_mean))) %>%
        mutate(Bootstrap = factor(as.numeric(Run) + 1, levels = c(1:50))) %>%
        filter(Bootstrap != run) %>%
        group_by(Bootstrap) %>%
        summarise(contains_features = all(robust_features %in% Feature)) %>%
        summarise(count = sum(contains_features)) %>%
        mutate(Included = count / 49) %>%
        mutate(Excluded = (49 - count) / 49)

      stability_landscape_int <- data.frame("Number of features" = n_feat, "Threshold (top X%)" = cutoff, "Reference Run" = run, "Included" = stability_value$Included, "Excluded" = stability_value$Excluded)
      stability_landscape_iteration <- rbind(stability_landscape_iteration, stability_landscape_int) }
    auc_it <- sum(diff(stability_landscape_iteration$Threshold..top.X..) *
                    (stability_landscape_iteration$Included[-1] + stability_landscape_iteration$Included[-length(stability_landscape_iteration$Included)]) / 2)
    auc_landscape_it <- data.frame("Number of features" = n_feat, "Reference Run" = run, "AUC" = auc_it)
    auc_landscape <- rbind(auc_landscape, auc_landscape_it)
  }
}

save(auc_landscape, file = paste0('../StabilityRun_', that_modality, ".RData"))

p1 <- auc_landscape %>%
  set_colnames(c("Number of features", "Reference Run", "AUC")) %>%
  group_by(`Reference Run`) %>%
  summarise(`Mean AUC` = mean(AUC)) %>%
  ungroup() %>%
  ggplot(aes(x = 1, y = `Reference Run`, fill = `Mean AUC`)) +
  geom_tile() +
  theme_minimal() +
  geom_text(aes(label = round(`Mean AUC`, 2)), size = 3.5) +
  theme(axis.title.x = element_blank(),
        axis.text.x = element_blank(),
        axis.ticks.x = element_blank()) +
  scale_fill_gradient("AUC", low = "#fee5d9", high = "#a50f15", limits = c(0, 1)) +
  scale_x_continuous(breaks = seq(1, 1, by = 1)) +
  scale_y_continuous(breaks = seq(0, 100, by = 10)) +
  ylab(paste("Iterations"))

print(p1)
ggsave(p1, file = paste0("../", project_name, "_StabilityAnalysis_", that_modality, "_Run.pdf"), width = 2, height = 10)

refine_reference_run <- auc_landscape %>%
  set_colnames(c("Number of features", "Reference Run", "AUC")) %>%
  group_by(`Reference Run`) %>%
  summarise(`Mean AUC` = mean(AUC)) %>%
  filter(`Mean AUC` == max(`Mean AUC`)) %>%
  slice(1)

print(refine_reference_run)

p2 <- auc_landscape %>%
  set_colnames(c("Number of features", "Reference Run", "AUC")) %>%
  filter(`Reference Run` == refine_reference_run$`Reference Run`) %>%
  ggplot(aes(x = `Number of features`, y = AUC, fill = AUC)) +
  geom_bar(stat = "identity") +
  theme_minimal() +
  geom_hline(yintercept = 0.7, linetype = "dashed", color = "red") +
  scale_fill_gradient("AUC", low = "#fee5d9", high = "#a50f15", limits = c(0, 1)) +
  scale_x_continuous(breaks = seq(1, number_of_features, by = 1)) +
  scale_y_continuous(limits = c(0, 1), breaks = seq(0, 1, by = 0.25)) +
  xlab(paste("Number of features")) +
  labs(title = paste("Reference Run", refine_reference_run$`Reference Run`))
ggsave(p2, file = paste0("../", project_name, "_StabilityAnalysis_", that_modality, "_RunRef.pdf"), width = 5, height = 2)

how_many_features <- 5

feature_panel_FACS_CyTOFab <- df_features %>%
  filter(DataSet == that_modality) %>%
  filter(Classifier == "Response") %>%
  filter(Method == that_classifier) %>%
  filter(Combination == "Combination") %>%
  select(Feature, Coefficient, DataSet, Run, Origin) %>% #filter(Coefficient ! = 0 ) %>%
  mutate(absCoef = abs(Coefficient)) %>%
  group_by(DataSet, Feature, Run) %>%
  summarise(mean_features = mean(absCoef), sem_features = sd(absCoef) / sqrt(length(absCoef))) %>%
  mutate(adjusted_mean = mean_features - sem_features) %>%
  arrange(DataSet, desc(adjusted_mean)) %>%
  group_by(DataSet, Run) %>%
  top_n(n = how_many_features, wt = adjusted_mean) %>%
  mutate(rank_per_run = row_number(desc(adjusted_mean))) %>%
  mutate(Bootstrap = factor(as.numeric(Run) + 1, levels = c(1:50))) %>%
  group_by(Feature) %>%
  filter(Bootstrap == refine_reference_run$`Reference Run`) %>%
  pull(Feature)

print(feature_panel_FACS_CyTOFab)

feature_info_FACS_CyTOFab <- df_features %>%
  filter(DataSet == that_modality) %>%
  filter(Classifier == "Response") %>%
  filter(Method == that_classifier) %>%
  filter(Combination == "Combination") %>%
  select(Feature, Coefficient, DataSet, Run, Origin) %>% #filter(Coefficient ! = 0 ) %>%
  mutate(absCoef = abs(Coefficient)) %>%
  mutate(Bootstrap = factor(as.numeric(Run) + 1, levels = c(1:50))) %>%
  filter(Bootstrap == refine_reference_run$`Reference Run`) %>%
  filter(Feature %in% feature_panel_FACS_CyTOFab)

p1 <- df_features %>%
  filter(DataSet == that_modality) %>%
  filter(Classifier == "Response") %>%
  filter(Method == that_classifier) %>%
  filter(Combination == "Combination") %>%
  select(Feature, Coefficient, DataSet, Run, Origin) %>% #filter(Coefficient ! = 0 ) %>%
  mutate(absCoef = abs(Coefficient)) %>%
  mutate(Bootstrap = factor(as.numeric(Run) + 1, levels = c(1:50))) %>%
  filter(Bootstrap == refine_reference_run$`Reference Run`) %>%
  ungroup %>%
  filter(Feature %in% feature_info_FACS_CyTOFab) %>%
  mutate(Feature = factor(Feature, levels = feature_panel_FACS_CyTOFab)) %>%
  ggplot(aes(x = Feature, y = absCoef)) +
  geom_boxplot() +
  expand_limits(y = c(0, 0.3)) +
  theme_minimal() +
  ylab("Absolute Feature Coefficient") +
  xlab("Features") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))

print(p1)

ggsave(p1, file = paste0("~/Desktop/", project_name, "_Bestfeatures_", that_modality, ".pdf"), width = 3, height = 2)

p1 <- df_features %>%
  filter(DataSet == that_modality) %>%
  filter(Classifier == "Response") %>%
  filter(Method == that_classifier) %>%
  filter(Combination == "Combination") %>%
  select(Feature, Coefficient, DataSet, Run, Origin) %>% #filter(Coefficient ! = 0 ) %>%
  #mutate(absCoef = abs(Coefficient)) %>%
  mutate(Bootstrap = factor(as.numeric(Run) + 1, levels = c(1:50))) %>%
  filter(Bootstrap == refine_reference_run$`Reference Run`) %>%
  ungroup %>%
  filter(Feature %in% feature_panel_FACS_CyTOFab) %>%
  mutate(Feature = factor(Feature, levels = feature_panel_FACS_CyTOFab)) %>%
  ggplot(aes(x = Feature, y = Coefficient)) +
  geom_boxplot() +
  expand_limits(y = c(0, 0.3)) +
  theme_minimal() +
  ylab("Absolute Feature Coefficient") +
  xlab("Features") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))

print(p1)

```

#### Using LASS0

```{r fig.width = 15, fig.height = 10}
Analyse_Modality_overview(which_modality = that_modality, which_classifier = "LASSO")
df_features %>%
  filter(DataSet == that_modality) %>%
  filter(Classifier == "Response") %>%
  filter(Method == "LASSO") %>%
  filter(Combination == "Combination") %>%
  select(Feature, Coefficient, DataSet, Run, Origin) %>%
  filter(Coefficient != 0) %>%
  mutate(absCoef = abs(Coefficient)) %>%
  group_by(DataSet, Feature, Run) %>%
  summarise(mean_features = mean(absCoef), sem_features = sd(absCoef) / sqrt(length(absCoef))) %>%
  mutate(adjusted_mean = mean_features - sem_features) %>%
  arrange(DataSet, desc(adjusted_mean)) %>%
  mutate(rank_per_run = row_number(desc(adjusted_mean))) %>%
  mutate(Bootstrap = factor(as.numeric(Run) + 1, levels = c(1:50))) %>%
  group_by(Bootstrap) %>%
  summarise(contains_features = all(feature_panel_FACS_CyTOFab %in% Feature)) %>%
  summarise(count = sum(contains_features))

lasso_run_FACS_CyTOFab <- df_features %>%
  filter(DataSet == that_modality) %>%
  filter(Classifier == "Response") %>%
  filter(Method == "LASSO") %>%
  filter(Combination == "Combination") %>%
  select(Feature, Coefficient, DataSet, Run, Origin) %>%
  filter(Coefficient != 0) %>%
  mutate(absCoef = abs(Coefficient)) %>%
  group_by(DataSet, Feature, Run) %>%
  summarise(mean_features = mean(absCoef), sem_features = sd(absCoef) / sqrt(length(absCoef))) %>%
  mutate(adjusted_mean = mean_features - sem_features) %>%
  arrange(DataSet, desc(adjusted_mean)) %>%
  mutate(rank_per_run = row_number(desc(adjusted_mean))) %>%
  mutate(Bootstrap = factor(as.numeric(Run) + 1, levels = c(1:50))) %>%
  filter(Bootstrap == refine_reference_run$`Reference Run`) %>%
  select(DataSet, Feature) %>%
  mutate(Method = "LASSO")

intersect(unique(lasso_run_FACS_CyTOFab$Feature), feature_panel_FACS_CyTOFab)
```

#### Using ELASTICNET

```{r fig.width = 15, fig.height = 10}
Analyse_Modality_overview(which_modality = that_modality, which_classifier = "ELASTICNET")
df_features %>%
  filter(DataSet == that_modality) %>%
  filter(Classifier == "Response") %>%
  filter(Method == "ELASTICNET") %>%
  filter(Combination == "Combination") %>%
  select(Feature, Coefficient, DataSet, Run, Origin) %>%
  filter(Coefficient != 0) %>%
  mutate(absCoef = abs(Coefficient)) %>%
  group_by(DataSet, Feature, Run) %>%
  summarise(mean_features = mean(absCoef), sem_features = sd(absCoef) / sqrt(length(absCoef))) %>%
  mutate(adjusted_mean = mean_features - sem_features) %>%
  arrange(DataSet, desc(adjusted_mean)) %>%
  mutate(rank_per_run = row_number(desc(adjusted_mean))) %>%
  mutate(Bootstrap = factor(as.numeric(Run) + 1, levels = c(1:50))) %>%
  group_by(Bootstrap) %>%
  summarise(contains_features = all(feature_panel_FACS_CyTOFab %in% Feature)) %>%
  summarise(count = sum(contains_features))

el_run_FACS_CyTOFab <- df_features %>%
  filter(DataSet == that_modality) %>%
  filter(Classifier == "Response") %>%
  filter(Method == "ELASTICNET") %>%
  filter(Combination == "Combination") %>%
  select(Feature, Coefficient, DataSet, Run, Origin) %>%
  filter(Coefficient != 0) %>%
  mutate(absCoef = abs(Coefficient)) %>%
  group_by(DataSet, Feature, Run) %>%
  summarise(mean_features = mean(absCoef), sem_features = sd(absCoef) / sqrt(length(absCoef))) %>%
  mutate(adjusted_mean = mean_features - sem_features) %>%
  arrange(DataSet, desc(adjusted_mean)) %>%
  mutate(rank_per_run = row_number(desc(adjusted_mean))) %>%
  mutate(Bootstrap = factor(as.numeric(Run) + 1, levels = c(1:50))) %>%
  filter(Bootstrap == refine_reference_run$`Reference Run`) %>%
  select(DataSet, Feature) %>%
  mutate(Method = "ELASTICNET")

intersect(unique(el_run_FACS_CyTOFab$Feature), feature_panel_FACS_CyTOFab)

summary_reg_FACS_CyTOFab <- rbind(el_run_FACS_CyTOFab, lasso_run_FACS_CyTOFab) %>%
  group_by(DataSet, Feature) %>%
  summarise(
    Methods = paste(unique(Method), collapse = ","),
    .groups = 'drop'
  ) %>%
  mutate(Occurrence = case_when(
    Methods == "ELASTICNET,LASSO" ~ "Both",
    Methods == "ELASTICNET" ~ "ELASTICNET only",
    Methods == "LASSO" ~ "LASSO only",
    TRUE ~ "None"
  ))

```

```{r}
order_of_features <- c("Ki67", "CXCR3", "IL4", "GPR15", "CD103", "IFNg","B7", "CXCL6", "TNF", "NK cells")

combined_regularizers <- rbind(summary_reg_olink_facs, summary_reg_FACS_CyTOF_Clinical,  summary_reg_OLINK_FACS_CyTOF) %>%
  mutate(Feature = gsub(pattern = "_FACS|_A4B7", replacement = "", Feature)) %>%
  mutate(Feature = gsub(pattern = "[.]", replacement = " ", Feature)) %>%
  mutate(Feature = gsub(Feature, pattern = "CD4mem_", replacement = "")) #%>%

p1 <- combined_regularizers %>%
  mutate(DataSet = factor(gsub(pattern = "_", replacement = "+", DataSet))) %>%
  mutate(DataSet = factor(gsub(pattern = "_", replacement = "\n", DataSet), levels = c("OLINK+FACS", "OLINK+FACS+CyTOF", "FACS+CyTOF+Clinical"))) %>%
  droplevels() %>%
  mutate(Feature = factor(Feature, levels = order_of_features)) %>%
  ggplot(aes(x = Feature, y = DataSet, fill = Occurrence)) +
  geom_tile(color = "white") +
  scale_fill_manual(values = c("None" = "red", "LASSO only" = "#d9d9d9", "ELASTICNET only" = "#969696", "Both" = "#252525")) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1),
        text = element_text(size = 8),
        axis.title.x = element_blank(),
        legend.position = "bottom",
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.background = element_blank(),
        axis.line = element_line(colour = "black"),
        axis.text = element_text(size = 8),
        axis.title = element_text(size = 8),
        legend.title = element_text(size = 8),
        legend.text = element_text(size = 8),
        strip.text.x = element_text(size = 8)) +
  labs(fill = "Method Occurrence")

p1
ggsave(p1, file = paste0("../Features_best_models_3models_reg.pdf"), width = 4, height = 2)
```

```{r}
cols_modality <- c("FACS" = "#c01c2c",
                   "OLINK" = "#2e3e9b",
                   "CyTOF" = "#5e2590",
                   "CyTOF a4b7" = "#ab43a8",
                   "Clinical" = "#656565")

combined_features <- rbind(feature_info_OLINK_FACS, feature_info_FACS_CyTOF_Clinical, feature_info_OLINK_FACS_CyTOF)

p1 <- combined_features %>%
  mutate(absCoef = abs(Coefficient)) %>%
  mutate(DataSet = factor(gsub(pattern = "_", replacement = "\n", DataSet))) %>%
  mutate(Feature = gsub(pattern = "_FACS|_A4B7", replacement = "", Feature)) %>%
  mutate(Feature = gsub(pattern = "[.]", replacement = " ", Feature)) %>%
  mutate(DataSet = factor(gsub(pattern = "_", replacement = "\n", DataSet), levels = c("OLINK+FACS", "OLINK+FACS+CyTOF", "FACS+CyTOF+Clinical"))) %>%
  group_by(DataSet) %>%
  mutate(z_score_group = scale(absCoef)) %>%
  droplevels() %>%
  arrange(Origin, Feature) %>%
  mutate(Feature = factor(Feature, levels = c(order_of_features))) %>%
  ggplot(aes(x = Feature, y = z_score_group)) +
  geom_hline(yintercept = 0, color = "grey") +
  geom_boxplot(width = 0.5, outlier.size = 0.4) +
  ylab("Feature Importance") +
  xlab("Features") +
  scale_fill_manual(values = cols_modality) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1),
        text = element_text(size = 8)) +
  theme(axis.title.x = element_blank(),
        legend.position = "bottom",
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.background = element_blank(),
        axis.line = element_line(colour = "black"),
        axis.text = element_text(
          size = 8),
        axis.title = element_text(size = 8),
        legend.title = element_text(size = 8),
        legend.text = element_text(size = 8),
        axis.text.x = element_text(size = 8),
        strip.text.x = element_text(size = 8))
p1

p2 <- combined_features %>%
  mutate(absCoef = abs(Coefficient)) %>%
  mutate(DataSet = factor(gsub(pattern = "_", replacement = "+", DataSet))) %>%
  mutate(Feature = gsub(pattern = "_FACS|_A4B7", replacement = "", Feature)) %>%
  mutate(Feature = gsub(pattern = "[.]", replacement = " ", Feature)) %>%
  mutate(DataSet = factor(gsub(pattern = "_", replacement = "\n", DataSet), levels = c("OLINK+FACS", "OLINK+FACS+CyTOF", "FACS+CyTOF+Clinical"))) %>%
  mutate(Feature = gsub(Feature, pattern = "CD4mem_", replacement = "")) %>%
  dplyr::count(Feature, DataSet, Origin) %>%
  mutate(n_norm = n / 10) %>%
  arrange(Origin) %>%
  droplevels() %>%
  mutate(Feature = factor(Feature, levels = order_of_features)) %>%
  ggplot(aes(x = Feature, y = DataSet)) +
  geom_point(size = 3) +
  scale_color_manual(values = cols_modality) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1),
        text = element_text(size = 8)) +
  theme(axis.title.x = element_blank(),
        legend.position = "bottom",
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.background = element_blank(),
        axis.line = element_line(colour = "black"),
        axis.text = element_text(
          size = 8),
        axis.title = element_text(size = 8),
        legend.title = element_text(size = 8),
        legend.text = element_text(size = 8),
        axis.text.x = element_text(size = 8),
        strip.text.x = element_text(size = 8))

p1 <- p1 +
  theme(axis.title.x = element_blank(),
        axis.text.x = element_blank(),
        axis.ticks.x = element_blank(),
        legend.position = "none")

p3 <- cowplot::plot_grid(p1, p2, labels = c('', ''), rows = 2, rel_heights = c(0.9, 0.8), align = 'v')
p3
ggsave(p3, file = paste0("../Features_best_models_3models.pdf"), width = 4, height = 2.5)
```

```{r}
sessionInfo()
```

### 2024-10-14 machine learning/ML_Treatment_efficacy_visualization.rmd
Add files via upload
Added ML code into desired folder.

---
title: "R Notebook"
output: html_notebook
---

```{r lib_load, message = F}
library(tidyverse)
library(magrittr)
library(rstatix)
library(reshape2)
library(ggpubr)

theme_figures <- function() {
  theme_minimal() %+replace%    #replace elements we want to change
    theme(axis.title.x = element_blank(),
          axis.ticks.x = element_blank(),
          legend.position = "bottom",
          panel.grid.major = element_blank(),
          panel.grid.minor = element_blank(),
          panel.background = element_blank(),
          axis.line = element_line(colour = "black"),
          axis.text = element_text(
            size = 8),
          axis.title = element_text(size = 8),
          legend.title = element_text(size = 8),
          legend.text = element_text(size = 8),
          strip.text.x = element_text(size = 8)) }
```
# Model performances

## K-fold cross validation approach of

```{r auc_read_in}

# Data read in
files_accurcy <- intersect(list.files(path = "../PredictiveModels/Data/", pattern = "R_treatment_*"),
                           list.files(path = "../PredictiveModels/Data/", pattern = "*pred_test.csv"))

combined_data <- list()

for (i in 1:length(files_accurcy)) {
  data <- t(read.csv(paste0("../PredictiveModels/Data/", files_accurcy[i]), row.names = 1))
  colnames(data) <- c('TrainingMeanAUC', 'TrainingStdAUC', 'TestingMeanAUC', 'TestingStdAUC')
  data <- t(round(data, 2))
  mat <- matrix(data, nrow = 2, ncol = 2)
  rownames(mat) <- c("MeanAUC", "SEMAUC")
  colnames(mat) <- c("Training", "Test")

  combined_data[[i]] <- mat
}
# Data tweaking
names(combined_data) <- gsub(pattern = "R_treatment_|_pred_test.csv", replacement = "", files_accurcy)
data_tweaking <- names(combined_data)[grepl(pattern = "treatment|response", x = names(combined_data))]
names(combined_data) <- ifelse(names(combined_data) %in% data_tweaking, names(combined_data), paste0(names(combined_data), "_response"))

df_accuracy <- melt(combined_data) %>%
  dcast(L1 + Var2 ~ Var1) %>%
  set_colnames(c('Condition', "Set", "MeanAUC", 'SEMAUC')) %>%
  mutate(Variable = gsub(Condition, pattern = "LinSVM_|PolynomialSVM_|RadialSVM_|LogReg_|LinearLASSO_|RandomForest_", replacement = "")) %>%
  mutate(Classifier = ifelse(grepl(pattern = "treatment", Variable), "Treatment", "Response")) %>%
  mutate(Method = factor(ifelse(grepl(pattern = "LogReg", Condition), "LogReg",
                                ifelse(grepl(pattern = "LASSO", Condition), "LinearLASSO",
                                       ifelse(grepl(pattern = "LinSVM", Condition), "LinSVM",
                                              ifelse(grepl(pattern = "PolynomialSVM", Condition), "PolynomialSVM",
                                                     ifelse(grepl(pattern = "RadialSVM", Condition), "RadialSVM",
                                                            ifelse(grepl(pattern = "RandomForest", Condition), "RandomForest", NA)))))), levels = c("LogReg", "LinearLASSO", "LinSVM", "PolynomialSVM", "RadialSVM", "RandomForest"))) %>%
  mutate(DataSet = gsub(Variable, pattern = "_response|_treatment", replacement = "")) %>%
  mutate(FeatureCutoff = ifelse(grepl("__0.05", DataSet), "95%",
                                ifelse(grepl("__0.1", DataSet), "90%",
                                       ifelse(grepl("__0.2$", DataSet), "80%",
                                              ifelse(grepl("__0.25", DataSet), "75%",
                                                     ifelse(grepl("__0.5", DataSet), "50%", "none"))))))

```

```{r}
selected_models <- df_accuracy %>%
  mutate(SelectionCriteria = MeanAUC - SEMAUC) %>%
  filter(DataSet %in% good_models) %>%
  filter(FeatureCutoff %in% c("none", "90%")) %>%
  filter(Method == "LogReg") %>%
  filter(Set == "Test") %>%
  arrange(SelectionCriteria) %>%
  slice_tail(n = 4) %>%
  pull(DataSet)
```

```{r}
files_accurcy <- intersect(list.files(path = "../PredictiveModels/Data/", pattern = "R_treatment_*"),
                           list.files(path = "../PredictiveModels/Data/", pattern = "*_coefficients.csv"))

combined_data <- list()

for (i in 1:length(files_accurcy)) {
  data <- read.csv(paste0("../PredictiveModels/Data/", files_accurcy[i]))
  rownames(data) <- paste0("K-fold", 1:dim(data)[1])
  data <- data[, -1]

  combined_data[[i]] <- data
}
names(combined_data) <- gsub(pattern = "R_treatment_|_coefficients.csv", replacement = "", files_accurcy)
data_tweaking <- names(combined_data)[grepl(pattern = "treatment|response", x = names(combined_data))]
names(combined_data) <- ifelse(names(combined_data) %in% data_tweaking, names(combined_data), paste0(names(combined_data), "_response"))

df_features <- melt(combined_data) %>%
  set_colnames(c('Feature', "Coefficient", "Condition")) %>%
  mutate(Variable = gsub(Condition, pattern = "LinSVM_|PolynomialSVM_|RadialSVM_|LogReg_|RandomForest_", replacement = "")) %>%
  mutate(Classifier = ifelse(grepl(pattern = "treatment", Variable), "Treatment", "Response")) %>%
  mutate(Method = factor(ifelse(grepl(pattern = "LogReg", Condition), "LogReg",
                                ifelse(grepl(pattern = "LASSO", Condition), "LinearLASSO",
                                       ifelse(grepl(pattern = "LinSVM", Condition), "LinSVM",
                                              ifelse(grepl(pattern = "PolynomialSVM", Condition), "PolynomialSVM",
                                                     ifelse(grepl(pattern = "RadialSVM", Condition), "RadialSVM",
                                                            ifelse(grepl(pattern = "RandomForest", Condition), "RandomForest", NA)))))), levels = c("LogReg", "LinearLASSO", "LinSVM", "PolynomialSVM", "RadialSVM", "RandomForest"))) %>%
  mutate(DataSet = gsub(Variable, pattern = "_response|_treatment", replacement = "")) %>%
  mutate(FeatureCutoff = ifelse(grepl("__0.05", DataSet), "95%",
                                ifelse(grepl("__0.1", DataSet), "90%",
                                       ifelse(grepl("__0.2$", DataSet), "80%",
                                              ifelse(grepl("__0.25", DataSet), "75%",
                                                     ifelse(grepl("__0.5", DataSet), "50%", "none")))))) %>%
                                                     mutate(FeatureEngineering = ifelse(FeatureCutoff == 'none', 'No', 'Yes'))

# Code Figure 4E right panel
svg("../PredictiveModels/results/F_4E_right.svg", height = 1.55, width = 2)
df_features %>%
  mutate(absCoef = abs(Coefficient)) %>%
  filter(Method == "LogReg") %>%
  filter(DataSet %in% selected_models) %>%
  filter(FeatureCutoff == "90%") %>%
  mutate(DataSet = factor(gsub(pattern = "_", replacement = "\n", DataSet))) %>%
mutate(Feature = gsub(pattern = "_FACS|_A4B7", replacement = "", Feature)) %>%
mutate(Feature = gsub(pattern = "[.]", replacement = " ", Feature)) %>%
  mutate(DataSet = gsub(pattern = "_", replacement = "\n", DataSet)) %>% filter(FinalDown == "none") %>%  group_by(Feature, DataSet) %>%
  mutate(Feature = factor(Feature, levels=c("Eosinophils","NK cells", "cMonocytes","Plasmablasts","cDCs", "DNER","CX3CL1", "CCL11","SCF", "CCL28","TNFB","ADA","TRANCE", "PD1", "CD103"))) %>%
  mutate(DataSet = factor(DataSet, levels = c("OLINK", "OLINK\nFACS", "FACS\nCyTOFab\nOLINK", "all\nCyTOFab"))) %>%
  dplyr::count(Feature, DataSet) %>%
  mutate(n_norm = n/10 ) %>%
  ggplot(aes(x= Feature, y = DataSet)) +
  geom_point() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1),
        text = element_text(size = 8)) +
  theme(axis.title.x = element_blank(),
        legend.position = "bottom",
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.background = element_blank(),
        axis.line = element_line(colour = "black"),
        axis.text = element_text(
          size = 6),
        axis.title = element_text(size = 8),
        legend.title = element_text(size = 8),
        legend.text = element_text(size = 6),
        axis.text.x=element_text(size = 6),
        strip.text.x = element_text(size = 8))
dev.off()

# Code Figure 4E left panel
svg("../PredictiveModels/results/F_4E_left.svg", height = 1.55, width = 2)
df_features %>%
  mutate(absCoef = abs(Coefficient)) %>%
  filter(Method == "LogReg") %>%
  filter(DataSet %in% selected_models) %>%
  filter(FeatureCutoff == "90%") %>%
  mutate(DataSet = factor(gsub(pattern = "_", replacement = "\n", DataSet))) %>%
  filter(FinalDown == "none") %>%
  mutate(Feature = gsub(pattern = "_FACS|_A4B7", replacement = "", Feature)) %>%
  mutate(Feature = gsub(pattern = "[.]", replacement = " ", Feature)) %>%
  mutate(DataSet = gsub(pattern = "_", replacement = "\n", DataSet))%>%
  group_by(DataSet) %>%
  mutate(z_score_group = scale(absCoef)) %>%
  ggplot(aes(x = reorder(Feature, -z_score_group, FUN = median), y = z_score_group)) +
  geom_hline(yintercept = 0, color = "grey") +
  geom_boxplot(width=0.5) +
  ylab("Z Score") +
  xlab("Features") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1),
        text = element_text(size = 8)) +
  theme(axis.title.x = element_blank(),
        legend.position = "bottom",
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.background = element_blank(),
        axis.line = element_line(colour = "black"),
        axis.text = element_text(
          size = 6),
        axis.title = element_text(size = 8),
        legend.title = element_text(size = 8),
        legend.text = element_text(size = 6),
        axis.text.x=element_text(size = 6),
        strip.text.x = element_text(size = 8))

  dev.off()
```

`

### 2024-10-14 machine learning/ML_Response_prediction_HelperScript.py

@@ -1,0 +1,288 @@
+import pandas as pd
+import numpy as np
+import matplotlib.pyplot as plt
+from sklearn.metrics import RocCurveDisplay, auc
+from numpy import mean
+from scipy import stats
+from sklearn.model_selection import permutation_test_score
+from sklearn.preprocessing import StandardScaler
+from sklearn.metrics import roc_auc_score
+from sklearn.inspection import permutation_importance
+from sklearn.linear_model import LogisticRegression
+from sklearn.model_selection import RepeatedStratifiedKFold
+import warnings
+warnings.filterwarnings('ignore')
+scaler = StandardScaler()
+
+# Plotting layout
+rc_ticks = {
+    "text.usetex": False,
+    "font.size": 8,
+    "legend.fontsize":4,
+    "legend.title_fontsize":4,
+    "xtick.labelsize":6,
+    "ytick.labelsize": 6,
+    "axes.labelsize":8,
+    "axes.titlesize":8,
+    "lines.linewidth": 0.5,
+    "axes.linewidth": 0.4,
+    "lines.markersize": 3,
+    "xtick.major.size": 2.5,
+    "xtick.major.width": 0.5,
+    "xtick.major.pad": 1,
+    "xtick.minor.size": 1.5,
+    "xtick.minor.width": 0.5,
+    "xtick.minor.pad": 1,
+    "ytick.major.size": 2.5,
+    "ytick.major.width": 0.5,
+    "ytick.major.pad": 1,
+    "ytick.minor.size": 1.5,
+    "ytick.minor.width": 0.5,
+    "ytick.minor.pad": 1,
+    "axes.labelpad": 1,
+    "axes.titlepad": 1
+}
+
+# Define the k-Fold Crossvalidation function
+# X: dataframe, modality values
+# X_scaler: dataframe, Scaled X dataframe
+# y: Vector, Outcome variable
+# classifier: sklearn-classifier, Define classifier to be used --> eg logistic regression
+# coeff_or_not: Boolean, Export feature coefficients
+# run_perm: Boolean, Calculate empirical p-values?
+# data_type: character, What is the name of the dataset currently analysed?
+# returns: feature coefficients(data), classifier, AUC statistics (y_auc_test)
+def run_k_fold_CV(X,X_scaler, y, classifier, coeff_or_not, run_perm,data_type):
+    data_set_size = len(y)
[…]

### 2024-10-14 machine learning/ML_Response_prediction_SystematicScanning.py

@@ -1,0 +1,73 @@
+exec(open("../ML_Response_prediction_HelperScript.py").read())
+import sys
+class Logger(object):
+    def __init__(self, filename="Default.log"):
+        self.terminal = sys.stdout
+        self.log = open(filename, "a")
+
+    def write(self, message):
+        self.terminal.write(message)
+        self.log.write(message)
+
+sys.stdout = Logger('../ML_outputLog.txt')
+
+# Where are modality information stored?
+clinicaliloc = range(12, 14)
+cytof_freiloc = range(22, 38)
+cytof_totaliloc = range(38, 54)
+cytof_totalAbiloc = range(55, 70)
+cytof_Abiloc = range(71, 87)
+facsiloc = range(163, 193)
+olinikiloc = range(87, 161)
+
+# Read in metadata file
+metadata = pd.read_csv("../Metadata_file.csv", sep=';', index_col=0, decimal=',')
+# Select data frame for response prediction
+only_timepoint_one = metadata[metadata["timepoint"] == 1]
+# Select data frame for treatment prediction
+only_treated = metadata[metadata["treatment"] == "VDZ"]
+
+# Run only single modalities
+run_modality(dataframe = only_timepoint_one,input_features =olinikiloc , data_modality_name = "onlyOLINK")
+
+contrast_data = only_timepoint_one.iloc[:, np.r_[clinicaliloc]]
+contrast_data['Response'] = only_timepoint_one['response']
+X, y = Run_analysis(data_frame=contrast_data)
+
+print(str("onlyClinical") + ":\n================================================")
+features = scan_over_models(X=X, y=y, data_type=str("onlyClinical"))
+
+run_modality(dataframe = only_timepoint_one,input_features =cytof_freiloc , data_modality_name = "onlyCyTOF")
+run_modality(dataframe = only_timepoint_one,input_features =cytof_Abiloc , data_modality_name = "onlyCyTOFab")
+run_modality(dataframe = only_timepoint_one,input_features =facsiloc , data_modality_name = "onlyFACS")
+
+# Run all duo modalities
+run_modality(dataframe = only_timepoint_one,input_features =(olinikiloc,clinicaliloc) ,data_modality_name = "OLINK_Clinical")
[…]
